---
title: "Hospital Anxiety and Depression Scale (HADS)"
author:
  - name: "Chiara Cotronei"
    url: https://www.research.ed.ac.uk/en/persons/chiara-cotronei
    affiliations:
      - ref: Lothian
  - name: "Alexander Rudge"
    affiliations:
      - ref: CGEM
  - name: "Nathan Constantine-Cooke"
    corresponding: true
    url: https://scholar.google.com/citations?user=2emHWR0AAAAJ&hl=en&oi=ao
    affiliations:
      - ref: CGEM
      - ref: HGU
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-expand: true
editor_options: 
  chunk_output_type: console
---

# Hospital Anxiety and Depression Scale

## Setting up in R

R packages

```{r}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = TRUE)
options(dplyr.summarise.inform = FALSE)

library(plyr) # Used for mapping values
suppressPackageStartupMessages(library(tidyverse)) # ggplot2, dplyr, and magrittr
library(readxl) # Read in Excel files
library(datefixR) # Standardise dates  
library(openxlsx)
library(tidyr)
library(ggdist)
suppressPackageStartupMessages(library(rstatix))
suppressPackageStartupMessages(library(table1))
library(knitr)
library(pander)
suppressPackageStartupMessages(library(kableExtra))
# Generate flowchart of cohort derivation
library(DiagrammeR)
library(DiagrammeRsvg)

#cox-regression
library(survival)
suppressPackageStartupMessages(library(survminer))
library(finalfit)

library(mice)
library(splines)
library(naniar)
library(ggmice)
```

## Helper functions

```{r}

summon_statistical_test <- function(data, dependent, independent) {
  
  # Chi square tests/Fisher's exact test/Wilcox/Kruskal-Wallis
  # for the dependent variable vs each independent variable
  # Dependent variables must be discrete
  
  purrr::map2_df(
    .x = independent,
    .y = dependent,
    .f = function(.x, .y) {
      # Remove NA
      data %<>%
        dplyr::filter(!is.na(!!sym(.x)), !is.na(!!sym(.y)))
      
      x <- data %>%
        dplyr::pull(.x)
      
      y <- data %>%
        dplyr::pull(.y)
      
      # Check if x is continuous (numeric) or not
      continuous_flag <- is.numeric(x)
      
      # If continuous, use a Wilcox/Mann Whitney if y has 2 levels
      # or Kruskal Wallis if >2
      n_levels_y <- levels(y) %>% length()
      
      if (continuous_flag & (n_levels_y <= 2)) {
        # Wilcoxon
        
        # Formula as score ~ group
        test_formula <- as.formula(paste0(.x, " ~ ", .y))
        
        wilcox_test(data, formula = test_formula) %>%
          # Calculate n for consistency with kruskal_test
          dplyr::mutate(n = n1 + n2) %>%
          # Don't need the adjusted p values as we adjust ourselves in a second
          dplyr::select(-tidyselect::any_of(c('.y.', 'group1', 'group2', 'n1', 'n2', 'p.adj', 'p.adj.signif'))) %>%
          dplyr::mutate(x = .x,
                        y = .y,
                        method = 'Wilcox test')
        
      } else if (continuous_flag & (n_levels_y > 2)) {
        
        # Formula as score ~ group
        test_formula <- as.formula(paste0(.x, " ~ ", .y))
        
        kruskal_test(data, formula = test_formula) %>%
          dplyr::select(-.y.) %>%
          dplyr::mutate(x = .x, y = .y)
        
      } else {
        # Otherwise use categorical tests
        # Chi square
        test_result <- chisq.test(x = x, y = y)
        
        # If any expected counts are < 5 then we should use a Fisher
        fisher_flag <- any(test_result$expected < 5)

        if (fisher_flag) {
          # If one of the counts is < 5 use Fisher exact test
          fisher_test(table(x, y), simulate.p.value = TRUE, B = 1e5) %>%
            dplyr::mutate(x = .x,
                          y = .y,
                          method = 'Fisher\'s exact test')
        } else {
          chisq_test(x = x, y = y) %>%
            dplyr::mutate(x = .x, y = .y)
        }
      }
    }
  ) %>%
    dplyr::mutate(p.adjust = p.adjust(p = p, method = 'fdr')) %>%
    dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
    dplyr::select(-tidyselect::any_of(c('p.signif'))) %>%
    dplyr::mutate(
      p.adjust.signif = dplyr::case_when(
        p.adjust < 0.0001 ~ "****",
        p.adjust < 0.001 ~ "***",
        p.adjust < 0.01  ~ "**",
        p.adjust < 0.05  ~ "*",
        .default = 'ns'
      )
    ) %>%
    # Reorder columns
    dplyr::select(
      y, x, n, statistic, df, p, p.adjust, p.adjust.signif, method
    )
  
}

# Function to create plots for baseline associations

summon_baseline_plot_discrete <- function(data, stat_tests, dependent, independent) {
  # Creating plot for baseline variables with the dependent variable
  # Discrete variables
  
  # Stat_tests is a dataframe from summon_statistical_test to get p-value info
  
  label <- stat_tests %>%
    dplyr::filter(x == independent) %>%
    {df <- .
    method <- df %>% dplyr::pull(method)
    p <- df %>% dplyr::pull(p.adjust)
    
    paste0(stringr::str_pad(method, 25, 'right'), "\n", "Adjusted p-value: ", p)
    
    }

  
  data %>%
    dplyr::filter(!is.na(!!sym(dependent)), !is.na(!!sym(independent))) %>%
    dplyr::group_by(!!rlang::sym(independent), !!rlang::sym(dependent)) %>%
    dplyr::count() %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!rlang::sym(independent)) %>%
    dplyr::mutate(p = n/sum(n)) %>%
    dplyr::ungroup() %>%
    ggplot(aes(x = !!sym(dependent), y = p, fill = !!sym(independent))) +
    geom_col(position = 'dodge') +
    annotate(
      "text",
      label = label,
      x = Inf,
      y = Inf,
      vjust = 2,
      hjust = 1.3
    ) +
    theme_minimal()
  
}


summon_baseline_plot_continuous <- function(data, stat_tests, dependent, independent) {
  # Creating plot for baseline variables with the dependent variable
  # Continuous independent variable
  
  # Stat_tests is a dataframe from summon_statistical_test to get p-value info
  label <- stat_tests %>%
    dplyr::filter(x == independent) %>%
    {df <- .
    method <- df %>% dplyr::pull(method)
    p <- df %>% dplyr::pull(p.adjust)
    
    paste0(stringr::str_pad(method, 25, 'right'), "\n", "Adjusted p-value: ", p)
    
    }
  
  # Plot
  data %>%
    dplyr::filter(
      !is.na(!!sym(dependent)),
      !is.na(!!sym(independent))
    ) %>%
    ggplot(aes(x = !!sym(independent), colour = !!sym(dependent))) +
    geom_density() +
    annotate(
      "text",
      label = label,
      x = Inf,
      y = Inf,
      vjust = 1.1,
      hjust = 1.3
    ) +
    ylab("") +
    theme_minimal()
  
}

# Function to perform coxph with mice for missing data
coxph_mice <- function(formula, data, ...){
  
  # Variables
  variables <- all.vars(formula)
  
  # Extract time and status variable names
  timevar <- variables[1]
  statusvar <- variables[2]
  
  # Select variables in the data
  data %<>%
    dplyr::select(tidyselect::all_of(variables)) %>%
    # Calculate the cumulative hazard
    dplyr::mutate(cumhaz = mice::nelsonaalen(
      data = .,
      timevar = !!sym(timevar),
      statusvar = !!sym(statusvar)
    ))
  
  # Predictor matrix - need to exclude time from the model
  pred_matrix <- mice::make.predictorMatrix(data)
  
  pred_matrix[, timevar] <- 0
  
  data_imp <- mice::mice(
    data = data,
    predictorMatrix = pred_matrix, ...)
  
  # Return the imputation
  return(data_imp)
  
}

print_chisq <- function(data) {
  data %>%
    knitr::kable(digits = c(0, 2, 4, 0, NA, NA, NA, 4, NA),
                 col.names = c("n",
                               "Statistic",
                               "p",
                               "df",
                               "Method",
                               "Dependent",
                               "Independent",
                               "Adjusted p-value",
                               "Adjusted significance")
               )
}

print_survival <- function(data) {
  data %>%
    broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    knitr::kable(digits = 3,
                 col.names = c("Covariate",
                               "Estimate",
                               "Std. Error",
                               "Statistic",
                               "p value",
                               "Lower CI",
                               "Upper CI")
                 )
}

plot_continuous_hr <- function(data, model, variable, splineterm = NULL){
  
  # Plotting continuous Hazard Ratio curves
  # Data is the same data used to fit the Cox model
  # model is a Cox model
  # variable is the variable we are plotting
  # splineterm is the spline used in the model, if there is one
  
  
  # Range of the variable to plot
  variable_range <- data %>%
    dplyr::pull(variable) %>%
    quantile(., probs = seq(0, 0.9, 0.01), na.rm = TRUE) %>%
    unname() %>%
    unique()
  
  # If there is a spline
  if (!is.null(splineterm)){
    
    # Recreate the spline
    spline <- with(data, eval(parse(text = splineterm)))
    
    # Get the spline terms for the variable values we are plotting
    spline_values <- predict(spline, newx = variable_range)
    
    # Get the reference values of the variables that were used to fit the Cox model
    X_ref <- model$means
    
    # New values to calculate the HR at 
    # Take the ref values and vary our variable of interest
    
    # Number of values
    n_values <- variable_range %>% length()
    
    X_ref <- matrix(rep(X_ref, n_values), nrow = n_values, byrow = TRUE)
    # Fix the column names
    colnames(X_ref) <- names(model$means)
    
    # Now add our new variable values
    X_new <- X_ref
    # Identify the correct columns
    column_pos <- stringr::str_detect(colnames(X_new), variable)
    # Add the values - ordering of columns and model term names should be preserved
    X_new[,column_pos] <- spline_values
    
    # Calculate the linear predictor and se
    # Method directly from predict.coxph from the survival package
    lp <- (X_new - X_ref)%*%coef(model) 
    
    lp_se <- rowSums((X_new - X_ref)%*%vcov(model)*(X_new - X_ref)) %>%
      as.numeric() %>%
      sqrt()
    
  } else {
    # If there is no spline then it is easier
    
    # Get the reference values of the variables that were used to fit the Cox model
    X_ref <- model$means
    
    # New values to calculate the HR at 
    # Take the ref values and vary our variable of interest
    
    # Number of values
    n_values <- variable_range %>% length()
    
    X_ref <- matrix(rep(X_ref, n_values), nrow = n_values, byrow = TRUE)
    # Fix the column names
    colnames(X_ref) <- names(model$means)
    
    # Now add our new variable values
    X_new <- X_ref
    
    # Add the values
    X_new[,variable] <- variable_range
    
    # Calculate the linear predictor and se
    # Method directly from predict.coxph from the survival package
    lp <- (X_new - X_ref)%*%coef(model) %>% as.numeric() 
    
    lp_se <- rowSums((X_new - X_ref)%*%vcov(model)*(X_new - X_ref)) %>% 
      as.numeric() %>%
      sqrt()
    
  }
  
  # Store results in a tibble
  data_pred <- tibble(
    !!sym(variable) := variable_range,
    p = lp,
    se = lp_se
  )
  
  # Plot
  data_pred %>%
    ggplot(aes(
      x = !!rlang::sym(variable),
      y = exp(p),
      ymin = exp(p - 1.96 * se),
      ymax = exp(p + 1.96 * se)
    )) +
    geom_point() +
    geom_line() +
    geom_ribbon(alpha = 0.2) +
    ylab("HR") + 
    theme_minimal()
  
}

# Function to perform an LRT
summon_lrt <- function(model, remove = NULL, add = NULL) {
  # Function that removes a term and performs a
  # Likelihood ratio test to determine its significance
  
  new_formula <- "~ ."
  
  if (!is.null(remove)) {
    new_formula <- paste0(new_formula, " - ", remove)
  } else {
    remove <- NA
  }
  
  if (!is.null(add)) {
    new_formula <- paste0(new_formula, " + ", add)
  } else {
    add <- NA
  }
  
  update(model, new_formula) %>%
    anova(model, ., test = "LRT") %>%
    broom::tidy() %>%
    dplyr::filter(!is.na(p.value)) %>%
    dplyr::select(-term) %>%
    dplyr::mutate(
      removed = remove,
      added = add
    ) %>%
    dplyr::mutate(test = "LRT") %>%
    dplyr::select(test, removed, added, tidyselect::everything())
}

```

## Load data

```{r}
# paths to PREdiCCt data
if (file.exists("/.dockerenv")) {
  # If running in docker
  data.path <- "data/final/20221004/"
  redcap.path <- "data/final/20231030/"
  prefix <- "data/end-of-follow-up/"
  outdir <- "data/processed"
} else {
  # Run on OS directly
  data.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20221004/"
  redcap.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20231030/"
  prefix <- "/Volumes/igmm/cvallejo-predicct/predicct/end-of-follow-up/"
  outdir <- "/Volumes/igmm/cvallejo-predicct/predicct/processed/"
  chiara <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
}

#read-in data
#hads is question 42 of the baseline questionnaire
hads <- read.xlsx(paste0(data.path, "Baseline2022/hads.xlsx"))
demographics <- read.xlsx(paste0(data.path, "Baseline2022/demographics2022.xlsx"))
IBD <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx"))
smoking <- readRDS(paste0(chiara, "smoking.rds"))
IMD <- readRDS(paste0(chiara, "IMD.rds"))
demo <- readRDS(paste0(outdir, "demo.RDS"))
flares <- readRDS(paste0(outdir, "flares/cutoff-flares.RDS"))
IBD_C <- readRDS(paste0(chiara, "IBD_C.RDS"))
IBD_FC <- readRDS(paste0(chiara, "IBD_FC.RDS"))
flares_soft <- readRDS(paste0(chiara, "flares_soft.RDS"))
flares_hard <- readRDS(paste0(chiara, "flares_hard.RDS"))
```

## Data Cleaning

```{r}
# Exclude cases with missing ParticipantId
# 60 removed - but they are all completely empty rows.
hads <- hads %>%
  filter(!is.na(ParticipantId))

demographics <- demographics %>% 
  filter(!is.na(ParticipantId))

IBD <- IBD %>% 
  filter(!is.na(ParticipantId))

# join demographics
hads %<>%
  dplyr::left_join(demographics %>%
                     dplyr::select(ParticipantNo, Sex, age, diagnosis),
                   by = "ParticipantNo")

# Convert diagnosis column to factor with labels
hads %<>%
  dplyr::mutate(
    diagnosis2 = dplyr::case_when(
      diagnosis == 1 ~ 1,
      diagnosis == 2 ~ 2,
      diagnosis == 3 ~ 2,
      diagnosis == 4 ~ 2
    )
  ) %>%
  dplyr::mutate(
    diagnosis2 = factor(
      diagnosis2,
      levels = c("1", "2"), 
      labels = c("CD", "UC/IBDU")
    )
  )

hads %<>%
  dplyr::mutate(Sex = factor(
    Sex,
    levels = c(1, 2),
    labels = c("Male", "Female")
  )
)

#exclude participants <17, 53 participants excluded, tot 1847

n_underage <- hads %>%
  dplyr::filter(age <= 17) %>%
  nrow()

hads %<>%
  dplyr::filter(age > 17)

# Rescale
hads %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = FeelTense:CanEnjoyBookTV,
      .fns = ~. - 1)
  )
  
#correcting inversely coded HADS
hads %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = c("FeelTense", 
                 "FrightenedFeelingSomethingAwful", 
                 "WorryingThoughts", 
                 "FeelCheerful", 
                 "FeelSlowedDown",
                 "LostInterestAppearance",
                 "FeelRestless",
                 "SuddenFeelingsPanic"),
      .fns = ~ 3 - .
    )
  )

#anxiety HADS composite score
hads %<>% 
  dplyr::mutate(anxiety_hads = FeelTense + 
           FrightenedFeelingSomethingAwful + 
           WorryingThoughts +
           SitAtEase +
           FrightenedFeelingButterflies +
           FeelRestless +
           SuddenFeelingsPanic)

#depression HADS composite score
hads %<>% 
  dplyr::mutate(depression_hads = EnjoyThings +
           CanLaugh +
           FeelCheerful +
           FeelSlowedDown +
           LostInterestAppearance +
           LookForward +
           CanEnjoyBookTV)

#remove breakdown hads columns
hads %<>%
  dplyr::select(-c(FeelTense:CanEnjoyBookTV))


#create age groups
hads %<>%
  dplyr::mutate(AgeGroup = cut(
    age,
    breaks = c(18, 24, 34, 44, 54, 65, Inf),
    labels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+"),
    include.lowest = TRUE
  ))

# FC data
hads <- hads %>% 
  dplyr::left_join(
    demo %>% dplyr::select(ParticipantNo, FC, cat), 
    by = "ParticipantNo")

# Add smoking
hads %<>% 
  dplyr::left_join(
    smoking %>% dplyr::select(ParticipantNo, Smoke), 
    by = "ParticipantNo")

# Smoke as a factor
hads %<>%
  dplyr::mutate(Smoke = forcats::as_factor(Smoke)) %>%
  dplyr::mutate(Smoke = forcats::fct_relevel(Smoke, 'Never', 'Previous', 'Current'))

# Add IMD
hads %<>% 
  dplyr::left_join(
    IMD, 
    by = "ParticipantNo") 

# IMD as a factor
hads %<>%
  dplyr::mutate(
    IMD = as.factor(IMD)
  )

# Join flare data
hads <- hads %>% 
  dplyr::left_join(
    IBD %>% select(ParticipantNo, FlaresInPastYear), 
    by = "ParticipantNo"
    )

#create flare groups
hads <- hads %>%
  dplyr::mutate(
    flare_group = ifelse(FlaresInPastYear == 0, "No flares", "1 or More Flares")
    )

# make flares into levels
hads %<>%
  dplyr::mutate(flare_group = factor(flare_group, levels = c("No flares", "1 or More Flares")))

# IBD control scores
hads %<>%
  dplyr::left_join(
    IBD_C %>% dplyr::select(
      ParticipantNo,
      control_score,
      OverallControl,
      control_8,
      control_grouped,
      vas_control
    ),
    by = "ParticipantNo"
  )

#change control_grouped and vas_control from character into factors
hads %<>% 
  dplyr::mutate(control_grouped = factor(control_grouped))

hads %<>% 
  dplyr::mutate(vas_control = factor(vas_control))

# Pivot longer
hads <- hads %>%
  tidyr::pivot_longer(
    cols = c(anxiety_hads, depression_hads),
    names_to = "hads_type", values_to = "hads_score"
  ) 

#score groups
hads <- hads %>% 
  mutate(
    score_group = cut(
      hads_score, 
      breaks = c(0, 7, 21), 
      labels = c("0-7", "8-21"), 
      include.lowest = TRUE)
  )

# Age in decades
hads %<>%
  dplyr::mutate(
    age_decade = age/10
  )


# FC maximum detectable value is 1250
hads %<>%
  dplyr::mutate(
    FC = dplyr::case_when(
      FC > 1250 ~ 1250,
      .default = FC
    )
  )

# Log transform FC due to extreme positive skew
hads %<>%
  dplyr::mutate(
    FC = log(FC)
  )

# Filter missing hads scores - 7 anxiety, 6 depression due to not answering all the questions
# hads %>%
#   dplyr::group_by(hads_type) %>%
#   dplyr::summarise(sum(is.na(score_group)))

hads %<>%
  dplyr::filter(!is.na(score_group))

```

## Missingness

```{r}
# Visualise missingness
# Only need to do it for one of anxiety and depression due to repeated participants

data_vis_missingness <- hads %>%
  # Select anxiety hads
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  # Select relevant variables
  dplyr::select(
    diagnosis,
    age,
    Sex,
    IMD,
    FC,
    flare_group,
    OverallControl,
    Smoke,
    SiteNo,
    score_group
  )
  
# Number of missing values per variable
data_vis_missingness %>%
  gg_miss_var()

# Pairwise missingness
data_vis_missingness %>%
  ggmice::plot_pattern(rotate = TRUE)

```

## Baseline Analysis

### Anxiety

#### Full Cohort

```{r}
# Select anxiety data
data_baseline_anxiety <- hads %>%
  dplyr::filter(hads_type == 'anxiety_hads')

# Independent variables
independent = c('diagnosis2',
                'age',
                'Sex',
                'FC',
                'Smoke')

# Dependent variable is score_group

# Statistical tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline_anxiety, 
  dependent = 'score_group', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline_anxiety,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'diagnosis2'
)

summon_baseline_plot_discrete(
  data = data_baseline_anxiety,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Sex'
)

summon_baseline_plot_discrete(
  data = data_baseline_anxiety,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Smoke'
)

summon_baseline_plot_continuous(
  data = data_baseline_anxiety,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline_anxiety,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'FC'
)

```

#### Ulcerative Colitis/IBD Unclassified

```{r}
# Select anxiety data and UC patients
data_baseline_anxiety_uc <- hads %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Independent variables
independent = c('age',
                'Sex',
                'FC',
                'Smoke')

# Dependent variable is score_group

# Statistical tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline_anxiety_uc, 
  dependent = 'score_group', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline_anxiety_uc,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Sex'
)

summon_baseline_plot_discrete(
  data = data_baseline_anxiety_uc,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Smoke'
)

summon_baseline_plot_continuous(
  data = data_baseline_anxiety_uc,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline_anxiety_uc,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'FC'
)

```

#### Crohn's Disease

```{r}
# Select anxiety data and CD patients
data_baseline_anxiety_cd <- hads %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  dplyr::filter(diagnosis2 == 'CD')

# Dependent variable is score_group

# Statistical tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline_anxiety_cd, 
  dependent = 'score_group', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline_anxiety_cd,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Sex'
)

summon_baseline_plot_discrete(
  data = data_baseline_anxiety_cd,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Smoke'
)

summon_baseline_plot_continuous(
  data = data_baseline_anxiety_cd,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline_anxiety_cd,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'FC'
)

```

### Depression

#### Full Cohort

```{r}
# Select depression data
data_baseline_depression <- hads %>%
  dplyr::filter(hads_type == 'depression_hads')

# Independent variables
independent = c('diagnosis2',
                'age',
                'Sex',
                'FC',
                'Smoke')

# Dependent variable is score_group

# Statistical tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline_depression, 
  dependent = 'score_group', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline_depression,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Sex'
)

summon_baseline_plot_discrete(
  data = data_baseline_depression,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Smoke'
)

summon_baseline_plot_continuous(
  data = data_baseline_depression,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline_depression,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'FC'
)

```

#### Ulcerative Colitis/IBD Unclassified

```{r}
# Select depression data and UC patients
data_baseline_depression_uc <- hads %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Independent variables
independent = c('age',
                'Sex',
                'FC',
                'Smoke')

# Dependent variable is score_group

# Statistical tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline_depression_uc, 
  dependent = 'score_group', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline_depression_uc,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Sex'
)

summon_baseline_plot_continuous(
  data = data_baseline_depression_uc,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline_depression_uc,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'FC'
)

```

#### Crohn's Disease

```{r}
# Select depression data and CD patients
data_baseline_depression_cd <- hads %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::filter(diagnosis2 == 'CD')

# Dependent variable is score_group

# Statistical tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline_depression_cd, 
  dependent = 'score_group', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline_depression_cd,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'Sex'
)


summon_baseline_plot_continuous(
  data = data_baseline_depression_cd,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline_depression_cd,
  stat_tests = statistical_tests,
  dependent = 'score_group',
  independent = 'FC'
)

```

## Survival Analysis

### Data Cleaning

```{r}
# Survival data
# Join HADS with flare data
# Patient reported
data_survival_soft <- hads %>% dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Anxiety soft
data_survival_anxiety_soft <- data_survival_soft %>%
  dplyr::filter(hads_type == 'anxiety_hads')

# UC
data_survival_anxiety_soft_uc <- data_survival_anxiety_soft %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# CD
data_survival_anxiety_soft_cd <- data_survival_anxiety_soft %>%
  dplyr::filter(diagnosis2 == 'CD')


# Depression soft
data_survival_depression_soft <- data_survival_soft %>%
  dplyr::filter(hads_type == 'depression_hads')

# UC
data_survival_depression_soft_uc <- data_survival_depression_soft %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# CD
data_survival_depression_soft_cd <- data_survival_depression_soft %>%
  dplyr::filter(diagnosis2 == 'CD')


# Hard flares
data_survival_hard <- hads %>% dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

# Anxiety soft
data_survival_anxiety_hard <- data_survival_hard %>%
  dplyr::filter(hads_type == 'anxiety_hads')

# UC
data_survival_anxiety_hard_uc <- data_survival_anxiety_hard %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# CD
data_survival_anxiety_hard_cd <- data_survival_anxiety_hard %>%
  dplyr::filter(diagnosis2 == 'CD')


# Depression soft
data_survival_depression_hard <- data_survival_hard %>%
  dplyr::filter(hads_type == 'depression_hads')

# UC
data_survival_depression_hard_uc <- data_survival_depression_hard %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# CD
data_survival_depression_hard_cd <- data_survival_depression_hard %>%
  dplyr::filter(diagnosis2 == 'CD')

```

### Shapes of continuous variables

#### Full Cohort

##### Patient reported flare

```{r}

# Cox with continuous variables
# Age
cox_age_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_anxiety_soft
)

# Spline for age
summon_lrt(cox_age_soft, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_anxiety_soft
)

# Spline for FC
summon_lrt(cox_fc_soft, remove = 'FC', add = 'ns(FC, df = 2)')
# No



```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_anxiety_soft, 
  model = cox_age_soft, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_anxiety_soft, 
  model = cox_fc_soft, 
  variable = 'FC')

```

##### Objective flare

```{r}

# Age
cox_age_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_anxiety_hard
)

# Spline for age
summon_lrt(cox_age_hard, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_anxiety_hard
)

# Spline for FC
summon_lrt(cox_fc_hard, remove = 'FC', add = 'ns(FC, df = 2)')
# No

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_anxiety_hard, 
  model = cox_age_hard, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_anxiety_hard, 
  model = cox_fc_hard, 
  variable = 'FC')

```

#### Ulcerative Colitis

##### Patient reported flare

```{r}
# Age
cox_age_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_uc
)

# Spline for age
summon_lrt(cox_age_soft_uc, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_uc
)

# Spline for FC
summon_lrt(cox_fc_soft_uc, remove = 'FC', add = 'ns(FC, df = 2)')
# No

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_anxiety_soft_uc, 
  model = cox_age_soft_uc, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_anxiety_soft_uc, 
  model = cox_fc_soft_uc, 
  variable = 'FC')

```

##### Objective flare

```{r}

# Age
cox_age_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

# Spline for age
summon_lrt(cox_age_hard_uc, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# Yes

# Refit
cox_age_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(age_decade, df = 2) +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

# Test for df = 3
summon_lrt(cox_age_hard_uc, remove = 'ns(age_decade, df = 2)', add = 'ns(age_decade, df = 3)')
# Yes - very close to 0.05

# Refit
cox_age_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(age_decade, df = 3) +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

# Test for df = 4
summon_lrt(cox_age_hard_uc, remove = 'ns(age_decade, df = 3)', add = 'ns(age_decade, df = 4)')
# Yes - again very close to 0.05

# Refit
cox_age_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(age_decade, df = 4) +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

# Test for df = 5
summon_lrt(cox_age_hard_uc, remove = 'ns(age_decade, df = 4)', add = 'ns(age_decade, df = 5)')
# 5 df not needed, stay with 4

# FC
cox_fc_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

# Spline for FC
summon_lrt(cox_fc_hard_uc, remove = 'FC', add = 'ns(FC, df = 2)')
# Yes, very close to 0.05

# Refit
cox_fc_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(FC, df = 2) +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

# Test for df = 3
summon_lrt(cox_fc_hard_uc, remove = 'ns(FC, df = 2)', add = 'ns(FC, df = 3)')
# No, stay with df = 2

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_anxiety_hard_uc, 
  model = cox_age_hard_uc, 
  variable = 'age_decade',
  splineterm = 'ns(age_decade, df = 4)')

# FC
plot_continuous_hr(
  data = data_survival_anxiety_hard_uc, 
  model = cox_fc_hard_uc, 
  variable = 'FC',
  splineterm = 'ns(FC, df = 2)')
```

#### Crohn's Disease

##### Patient reported flare

```{r}

# Age
cox_age_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_cd
)

# Spline for age
summon_lrt(cox_age_soft_cd, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_cd
)

# Spline for FC
summon_lrt(cox_fc_soft_cd, remove = 'FC', add = 'ns(FC, df = 2)')
# No

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_anxiety_soft_cd, 
  model = cox_age_soft_cd, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_anxiety_soft_cd, 
  model = cox_fc_soft_cd, 
  variable = 'FC')

```

##### Objective flare

```{r}

# Age
cox_age_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_cd
)

# Spline for age
summon_lrt(cox_age_hard_cd, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_cd
)

# Spline for FC
summon_lrt(cox_fc_hard_cd, remove = 'FC', add = 'ns(FC, df = 2)')
# No

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_anxiety_hard_cd, 
  model = cox_age_hard_cd, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_anxiety_hard_cd, 
  model = cox_fc_hard_cd, 
  variable = 'FC')

```

Summary:

Full cohort, soft, , hard, no.

UC, soft, . Hard, age df = 4, FC df = 2.

CD, soft, . Hard, no splines.

### Anxiety

#### Full Cohort

##### Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_anxiety_soft, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Anxiety Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_soft
)

cox_anxiety_soft %>%
  print_survival()

```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_soft_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_soft,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_soft_pool <- with(
  cox_anxiety_soft_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      age_decade +
      FC +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_soft_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_anxiety_hard, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Anxiety Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_hard
)

cox_anxiety_hard %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_hard_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_hard,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_hard_pool <- with(
  cox_anxiety_hard_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      age_decade +
      FC +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_hard_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```

#### Ulcerative Colitis

##### Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_anxiety_soft_uc, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Anxiety Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_uc
)

cox_anxiety_soft_uc

```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_soft_uc_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_uc,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_soft_uc_pool <- with(cox_anxiety_soft_uc_mice,
     coxph(
       Surv(time, DiseaseFlareYN) ~
         score_group +
         IMD +
         Sex +
         age_decade +
         FC +
         Smoke +
         frailty(SiteNo)
     )) %>%
  mice::pool()

cox_anxiety_soft_uc_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_anxiety_hard_uc, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Anxiety Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    ns(age_decade, df = 4) +
    ns(FC, df = 2) +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

cox_anxiety_hard_uc %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
# Note: mice does not support splines so I cannot impute using a spline.
cox_anxiety_hard_uc_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_hard_uc_pool <- with(cox_anxiety_hard_uc_mice,
     coxph(
       Surv(time, DiseaseFlareYN) ~
         score_group +
         IMD +
         Sex +
         ns(age_decade, df = 4) +
         ns(FC, df = 2) +
         Smoke +
         frailty(SiteNo)
     )) %>%
  mice::pool()

cox_anxiety_hard_uc_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```

#### Crohn's Disease

##### Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_anxiety_soft_cd, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Anxiety Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_cd
)

cox_anxiety_soft_cd %>%
  print_survival()

```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_soft_cd_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_cd,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_soft_cd_pool <- with(cox_anxiety_soft_cd_mice,
     coxph(
       Surv(time, DiseaseFlareYN) ~
         score_group +
         IMD +
         Sex +
         age_decade +
         FC +
         Smoke +
         frailty(SiteNo)
     )) %>%
  mice::pool()

cox_anxiety_soft_cd_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_anxiety_hard_cd, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Anxiety Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_cd
)

cox_anxiety_hard_cd %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_hard_cd_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_cd,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_hard_cd_pool <- with(
  cox_anxiety_hard_cd_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      age_decade +
      FC +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_hard_cd_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```

### Depression

#### Full Cohort

##### Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_depression_soft, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Depression Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_depression_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_soft
)

cox_depression_soft %>%
  print_survival()

```

```{r}
# Cox model with MICE
# Run the mice
cox_depression_soft_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_soft,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_depression_soft_pool <- with(
  cox_depression_soft_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      age_decade +
      FC +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_depression_soft_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_depression_hard, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Depression Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_depression_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_hard
)

cox_depression_hard %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
cox_depression_hard_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_hard,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_depression_hard_pool <- with(
  cox_depression_hard_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      age_decade +
      FC +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_depression_hard_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```

#### Ulcerative Colitis

##### Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_depression_soft_uc, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Depression Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_depression_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_soft_uc
)

cox_depression_soft_uc %>%
  print_survival()

```

```{r}
# Cox model with MICE
# Run the mice
cox_depression_soft_uc_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_soft_uc,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_depression_soft_uc_pool <- with(
  cox_depression_soft_uc_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      age_decade +
      FC +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_depression_soft_uc_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_depression_hard_uc, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Depression Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_depression_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    ns(age_decade, df = 4) +
    ns(FC, df = 2) +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_hard_uc
)

cox_depression_hard_uc %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
# Note: mice does not support splines so I cannot impute using a spline.
cox_depression_hard_uc_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_hard_uc,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_depression_hard_uc_pool <- with(
  cox_depression_hard_uc_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      ns(age_decade, df = 4) +
      ns(FC, df = 2) +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_depression_hard_uc_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```

#### Crohn's Disease

##### Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_depression_soft_cd, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Depression Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_depression_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_soft_cd
)

cox_depression_soft_cd %>%
  print_survival()

```

```{r}
# Cox model with MICE
# Run the mice
cox_depression_soft_cd_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_soft_cd,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_depression_soft_cd_pool <- with(
  cox_depression_soft_cd_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      age_decade +
      FC +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_depression_soft_cd_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_depression_hard_cd, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'HADS Depression Score',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_depression_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_hard_cd
)

cox_depression_hard_cd %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
cox_depression_hard_cd_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_survival_depression_hard_cd,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_depression_hard_cd_pool <- with(
  cox_depression_hard_cd_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group +
      IMD +
      Sex +
      age_decade +
      FC +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_depression_hard_cd_pool %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```



## Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>

Licensed by <a href="https://creativecommons.org/licenses/by/4.0/">CC BY</a> unless otherwise stated.
