---
title: "Does change in disease activity over time affect outcomes?"
format:
  html:
    code-fold: true
    code-tools: true
toc: true
toc-depth: 4
editor_options:
  chunk_output_type: console
---

## Introduction

This analysis addresses whether change in disease activity over time is
associated with subsequent flare outcomes, using IBD-Control-VAS
(`OverallControl`) as the measure of disease activity. Unlike `control_8`
(the 8-item IBD-Control sub-scale), which is only available as a single
baseline value per participant, `OverallControl` is collected both at
baseline (`Baseline2022/IBD.xlsx`) and monthly during follow-up
(`monthlyQ.xlsx`), making it the only IBD-Control measure that can be
modelled as a genuinely time-varying covariate.

`OverallControl` is added as a **time-dependent confounder** to each of the
main analysis's extended Cox models (`Longitudinal analysis/{HADS,PHQ,PSQI,
Exercise,Fatigue,Life Events}/`), rather than modelled as a standalone
exposure. Every risk factor already covered by the main analysis is
re-fitted here with time-dependent `OverallControl` added alongside its own
time-dependent exposure, for Crohn's Disease and Ulcerative Colitis/IBDU
separately, for both patient-reported and objective flare - the same
structure and covariate set as the reference models (`IMD + Sex +
age_decade + FC + Smoke + frailty(SiteNo)`), with `OverallControl_per10`
added to each formula. `OverallControl` enters per 10-point change (of the
0-100 scale) for an interpretable hazard ratio.

```{r}
#| message: false
library(survival)
source("poisson_helpers.R")  # -> population_cohort, covariates, demo_tbl,
                              # monthly, chiara, data.path
```

## Helper functions

```{r}
# Takes already-tidied, already-filtered data (not the raw model) -
# run_cox_cell() below tidies each fit once and filters to the risk
# factor's own term(s), so confounders (including OverallControl, which is
# fitted for adjustment only) never reach this table.
print_survival <- function(tidy_data) {
  tidy_data %>%
    dplyr::select(term, estimate, std.error, statistic, conf.low, conf.high, p.value) %>%
    dplyr::mutate(dplyr::across(.cols = where(is.numeric), .fns = ~ signif(.x, 3))) %>%
    dplyr::mutate(p.value = as.character(p.value)) %>%
    knitr::kable(
      col.names = c("Covariate", "Hazard Ratio", "Std. Error", "Statistic", "Lower 95% CI", "Upper 95% CI", "P-value"),
      format.args = list(drop0trailing = TRUE)
    )
}

covariates_dx <- covariates %>%
  dplyr::left_join(population_cohort %>% dplyr::select(ParticipantNo, diagnosis2), by = "ParticipantNo")
```

## IBD-Control-VAS over time - baseline and monthly follow-up

```{r}
overall_control_baseline <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId), !is.na(OverallControl)) %>%
  dplyr::transmute(ParticipantNo, month = 0, OverallControl)

overall_control_followup <- monthly %>%
  dplyr::filter(!is.na(OverallControl)) %>%
  dplyr::transmute(ParticipantNo, month = Q_month, OverallControl)

overall_control_long <- dplyr::bind_rows(overall_control_baseline, overall_control_followup) %>%
  dplyr::distinct(ParticipantNo, month, .keep_all = TRUE)
```

## Trajectory figure

Mean `OverallControl` (with 95% CI) at each assessed month, by diagnosis
group. 0-100 scale, higher = better control.

```{r}
#| fig-width: 9
#| fig-height: 5
overall_control_long %>%
  dplyr::left_join(population_cohort %>% dplyr::select(ParticipantNo, diagnosis2), by = "ParticipantNo") %>%
  dplyr::filter(!is.na(diagnosis2), month <= 24) %>%
  dplyr::group_by(diagnosis2, month) %>%
  dplyr::summarise(
    mean_control = mean(OverallControl, na.rm = TRUE),
    se = sd(OverallControl, na.rm = TRUE) / sqrt(dplyr::n()),
    n = dplyr::n(),
    .groups = "drop"
  ) %>%
  dplyr::mutate(ci_low = mean_control - 1.96 * se, ci_high = mean_control + 1.96 * se) %>%
  ggplot2::ggplot(ggplot2::aes(x = month, y = mean_control, colour = diagnosis2, fill = diagnosis2)) +
  ggplot2::geom_ribbon(ggplot2::aes(ymin = ci_low, ymax = ci_high), alpha = 0.15, colour = NA) +
  ggplot2::geom_line(linewidth = 0.8) +
  ggplot2::geom_point(size = 1.5) +
  ggplot2::scale_colour_brewer(palette = "Set1", name = "Diagnosis") +
  ggplot2::scale_fill_brewer(palette = "Set1", guide = "none") +
  ggplot2::coord_cartesian(ylim = c(0, 100)) +
  ggplot2::labs(
    title = "IBD-Control-VAS over time, by diagnosis group",
    x = "Month", y = "Mean OverallControl (0-100, higher = better control)"
  ) +
  ggplot2::theme_minimal()
```

## Base survival endpoint

`flares_soft.RDS`/`flares_hard.RDS` provide the first-flare survival
endpoint (`time`, `DiseaseFlareYN`), the same source used for the main
analysis's Cox models. Built before sourcing any risk factor's
`data_cleaning.R` below, since some of those scripts also define
`flares_soft`/`flares_hard` as a side effect - harmless (identical source
files, reloaded), but this keeps the base data unambiguous.

```{r}
flares_soft <- readRDS(paste0(chiara, "flares_soft.RDS"))
flares_hard <- readRDS(paste0(chiara, "flares_hard.RDS"))

cox_base_soft <- covariates_dx %>%
  dplyr::inner_join(flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time), by = "ParticipantNo") %>%
  dplyr::transmute(
    ParticipantNo, SiteNo, diagnosis2, Sex, age_decade, FC, Smoke, IMD,
    time = softflare_time, DiseaseFlareYN = softflare
  ) %>%
  dplyr::filter(time > 0)

cox_base_hard <- covariates_dx %>%
  dplyr::inner_join(flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time), by = "ParticipantNo") %>%
  dplyr::transmute(
    ParticipantNo, SiteNo, diagnosis2, Sex, age_decade, FC, Smoke, IMD,
    time = hardflare_time, DiseaseFlareYN = hardflare
  ) %>%
  dplyr::filter(time > 0)

overall_control_td <- overall_control_long %>%
  dplyr::arrange(ParticipantNo, month) %>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(OverallControl, .direction = "down") %>%
  dplyr::ungroup() %>%
  dplyr::transmute(ParticipantNo, day = month * 365 / 12, OverallControl_per10 = OverallControl / 10)
```

## Model-fitting and reporting helpers

Each risk factor's own time-dependent score keeps a generic `score_group`
name (matching the recurrent-event analogue in `Poisson Analysis -
Time-varying covariates.qmd`), so one set of helpers covers every exposure
below.

```{r}
# (ParticipantNo, month, score_group) -> the day-based, LOCF-filled format
# tdc() needs.
exposure_to_day <- function(exposure_long) {
  exposure_long %>%
    dplyr::distinct(ParticipantNo, month, .keep_all = TRUE) %>%
    dplyr::arrange(ParticipantNo, month) %>%
    dplyr::group_by(ParticipantNo) %>%
    tidyr::fill(score_group, .direction = "down") %>%
    dplyr::ungroup() %>%
    dplyr::transmute(ParticipantNo, day = as.numeric(month) * 365 / 12, score_group)
}

# Start-stop Cox dataset for one flare type, adding both the risk factor's
# own time-dependent score and OverallControl_per10 as a second
# time-dependent confounder.
build_cox_td <- function(cox_base, exposure_day) {
  tmerge(cox_base, cox_base, id = ParticipantNo, endpoint = event(time, DiseaseFlareYN)) %>%
    tmerge(data2 = exposure_day, id = ParticipantNo, score_group = tdc(day, score_group)) %>%
    tmerge(data2 = overall_control_td, id = ParticipantNo, OverallControl_per10 = tdc(day, OverallControl_per10))
}

cox_hazard_ratios <- tibble::tibble()

run_cox_cell <- function(data_merged, dx, exposure_name, outcome_label, diagnosis_label) {
  d <- data_merged %>% dplyr::filter(diagnosis2 == dx)
  fit <- coxph(
    Surv(tstart, tstop, endpoint) ~
      score_group + OverallControl_per10 + IMD + Sex + age_decade + FC + Smoke + frailty(SiteNo),
    data = d
  )
  cat("n =", dplyr::n_distinct(d$ParticipantNo), "\n\n")

  # OverallControl_per10 and the other covariates are fitted for adjustment
  # only - only the risk factor's own term(s) are reported or plotted.
  tidy_exposure_only <- fit %>%
    broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "score_group"))

  cox_hazard_ratios <<- dplyr::bind_rows(
    cox_hazard_ratios,
    tidy_exposure_only %>% dplyr::mutate(exposure = exposure_name, outcome = outcome_label, diagnosis = diagnosis_label)
  )

  tidy_exposure_only %>% print_survival()
}
```

## HADS (anxiety and depression)

Same item recoding and composite-score formulas as
`Longitudinal analysis/HADS/HADS.qmd`.

```{r}
#| message: false
hads_items <- c(
  "FeelTense", "FrightenedFeelingSomethingAwful", "WorryingThoughts", "SitAtEase",
  "FrightenedFeelingButterflies", "FeelRestless", "SuddenFeelingsPanic",
  "EnjoyThings", "CanLaugh", "FeelCheerful", "FeelSlowedDown",
  "LostInterestAppearance", "LookForward", "CanEnjoyBookTV"
)
inverse_coded_items <- c(
  "FeelTense", "FrightenedFeelingSomethingAwful", "WorryingThoughts", "FeelCheerful",
  "FeelSlowedDown", "LostInterestAppearance", "FeelRestless", "SuddenFeelingsPanic"
)

score_hads <- function(df) {
  df %>%
    dplyr::mutate(dplyr::across(tidyselect::all_of(hads_items), ~ . - 1)) %>%
    dplyr::mutate(dplyr::across(tidyselect::all_of(inverse_coded_items), ~ 3 - .)) %>%
    dplyr::mutate(
      anxiety_hads = FeelTense + FrightenedFeelingSomethingAwful + WorryingThoughts +
        SitAtEase + FrightenedFeelingButterflies + FeelRestless + SuddenFeelingsPanic,
      depression_hads = EnjoyThings + CanLaugh + FeelCheerful + FeelSlowedDown +
        LostInterestAppearance + LookForward + CanEnjoyBookTV
    )
}

hads_baseline_raw <- readxl::read_xlsx(paste0(data.path, "Baseline2022/hads.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId))
hads_followup_raw <- readxl::read_xlsx(paste0(data.path, "Followup/hads.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId))

hads_long <- dplyr::bind_rows(
  score_hads(hads_baseline_raw) %>% dplyr::transmute(ParticipantNo, month = 0, anxiety_hads, depression_hads),
  score_hads(hads_followup_raw) %>% dplyr::transmute(ParticipantNo, month = Q_month, anxiety_hads, depression_hads)
)

anxiety_long <- hads_long %>%
  dplyr::transmute(
    ParticipantNo, month,
    score_group = cut(anxiety_hads, breaks = c(0, 7, 21), labels = c("0-7", "8-21"), include.lowest = TRUE)
  )
depression_long <- hads_long %>%
  dplyr::transmute(
    ParticipantNo, month,
    score_group = cut(depression_hads, breaks = c(0, 7, 21), labels = c("0-7", "8-21"), include.lowest = TRUE)
  )

anxiety_day <- exposure_to_day(anxiety_long)
depression_day <- exposure_to_day(depression_long)

data_soft_anxiety <- build_cox_td(cox_base_soft, anxiety_day)
data_hard_anxiety <- build_cox_td(cox_base_hard, anxiety_day)
data_soft_depression <- build_cox_td(cox_base_soft, depression_day)
data_hard_depression <- build_cox_td(cox_base_hard, depression_day)
```

### Anxiety

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_anxiety, "CD", "Anxiety", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_anxiety, "CD", "Anxiety", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_anxiety, "UC/IBDU", "Anxiety", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_anxiety, "UC/IBDU", "Anxiety", "Objective flare", "Ulcerative Colitis/IBDU")
```

### Depression

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_depression, "CD", "Depression", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_depression, "CD", "Depression", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_depression, "UC/IBDU", "Depression", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_depression, "UC/IBDU", "Depression", "Objective flare", "Ulcerative Colitis/IBDU")
```

## PHQ (somatisation)

Sourced from `Longitudinal analysis/PHQ/data_cleaning.R`. `phq_long` is
that file's own long-format output.

```{r}
#| message: false
source("../Longitudinal analysis/PHQ/data_cleaning.R")

phq_score_long <- phq_long %>%
  dplyr::transmute(ParticipantNo, month, score_group = somatisation)
phq_day <- exposure_to_day(phq_score_long)

data_soft_phq <- build_cox_td(cox_base_soft, phq_day)
data_hard_phq <- build_cox_td(cox_base_hard, phq_day)
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_phq, "CD", "Somatisation", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_phq, "CD", "Somatisation", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_phq, "UC/IBDU", "Somatisation", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_phq, "UC/IBDU", "Somatisation", "Objective flare", "Ulcerative Colitis/IBDU")
```

## PSQI (sleep disturbance)

Sourced from `Longitudinal analysis/PSQI/data_cleaning.R`. `psqi` is that
file's own long-format output.

```{r}
#| message: false
source("../Longitudinal analysis/PSQI/data_cleaning.R")

psqi_score_long <- psqi %>%
  dplyr::transmute(ParticipantNo, month, score_group = SleepDisturbance)
psqi_day <- exposure_to_day(psqi_score_long)

data_soft_psqi <- build_cox_td(cox_base_soft, psqi_day)
data_hard_psqi <- build_cox_td(cox_base_hard, psqi_day)
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_psqi, "CD", "Sleep disturbance", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_psqi, "CD", "Sleep disturbance", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_psqi, "UC/IBDU", "Sleep disturbance", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_psqi, "UC/IBDU", "Sleep disturbance", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Exercise

Sourced from `Longitudinal analysis/Exercise/data_cleaning.R`. `exercise`
is that file's own long-format output (`MinimumExercise`: meets WHO minimum
activity guidelines).

```{r}
#| message: false
source("../Longitudinal analysis/Exercise/data_cleaning.R")

exercise_score_long <- exercise %>%
  dplyr::transmute(ParticipantNo, month, score_group = MinimumExercise)
exercise_day <- exposure_to_day(exercise_score_long)

data_soft_exercise <- build_cox_td(cox_base_soft, exercise_day)
data_hard_exercise <- build_cox_td(cox_base_hard, exercise_day)
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_exercise, "CD", "Exercise", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_exercise, "CD", "Exercise", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_exercise, "UC/IBDU", "Exercise", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_exercise, "UC/IBDU", "Exercise", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Fatigue

Sourced from `Longitudinal analysis/Fatigue/data_cleaning.R`. Unlike the
other exposures, fatigue (`OftenLackEnergy`) is assessed every month via
the monthly follow-up questionnaire, not just at 0/12/24 -
`exposure_to_day()` handles this the same way regardless.

```{r}
#| message: false
source("../Longitudinal analysis/Fatigue/data_cleaning.R")

fatigue_score_long <- fatigue %>%
  dplyr::transmute(ParticipantNo, month, score_group = OftenLackEnergy)
fatigue_day <- exposure_to_day(fatigue_score_long)

data_soft_fatigue <- build_cox_td(cox_base_soft, fatigue_day)
data_hard_fatigue <- build_cox_td(cox_base_hard, fatigue_day)
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_fatigue, "CD", "Fatigue", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_fatigue, "CD", "Fatigue", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_fatigue, "UC/IBDU", "Fatigue", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_fatigue, "UC/IBDU", "Fatigue", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Life Events

Sourced from `Longitudinal analysis/Life Events/data_cleaning.R`.
`AnyLifeEvents` is already binary (No/Yes), no cut-point needed.

```{r}
#| message: false
source("../Longitudinal analysis/Life Events/data_cleaning.R")

lifeevents_score_long <- lifeevents %>%
  dplyr::transmute(ParticipantNo, month, score_group = AnyLifeEvents)
lifeevents_day <- exposure_to_day(lifeevents_score_long)

data_soft_lifeevents <- build_cox_td(cox_base_soft, lifeevents_day)
data_hard_lifeevents <- build_cox_td(cox_base_hard, lifeevents_day)
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_lifeevents, "CD", "Any life event", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_lifeevents, "CD", "Any life event", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cox_cell(data_soft_lifeevents, "UC/IBDU", "Any life event", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cox_cell(data_hard_lifeevents, "UC/IBDU", "Any life event", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Combined forest plot

Each risk factor's own hazard ratio, adjusted for the main analysis's usual
confounders (`IMD`, `Sex`, `age_decade`, `FC`, `Smoke`, `frailty(SiteNo)`)
plus time-dependent `OverallControl`. `cox_hazard_ratios` only ever
contains each risk factor's own term(s) - confounders, including
`OverallControl_per10`, are fitted for adjustment only and never enter this
table (see `run_cox_cell()` in Setup).

```{r}
#| fig-width: 10
#| fig-height: 10
cox_hazard_ratios %>%
  dplyr::mutate(
    label = paste0(exposure, ": ", term),
    significant = conf.low > 1 | conf.high < 1,
    diagnosis = factor(diagnosis, levels = c("Crohn's Disease", "Ulcerative Colitis/IBDU"))
  ) %>%
  ggplot2::ggplot(ggplot2::aes(x = estimate, y = label, colour = significant)) +
  ggplot2::geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = conf.low, xmax = conf.high)) +
  ggplot2::scale_x_log10() +
  ggplot2::scale_colour_manual(values = c(`TRUE` = "#0F8B8D", `FALSE` = "grey60"), guide = "none") +
  ggplot2::facet_grid(diagnosis ~ outcome) +
  ggplot2::labs(x = "Adjusted hazard ratio (95% CI, log scale)", y = NULL) +
  ggplot2::theme_minimal()
```

# Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>
