---
title: "Recurrent event analysis (Poisson / negative binomial)"
author:
  - name: "Alexander Rudge"
  - name: "Chiara Cotronei"
  - name: "Nathan Constantine-Cooke"
    corresponding: true
    url: https://scholar.google.com/citations?user=2emHWR0AAAAJ&hl=en&oi=ao
format:
  html:
    code-fold: true
    code-tools: true
toc: true
toc-depth: 3
editor_options:
  chunk_output_type: console
---

## Introduction

Poisson/negative-binomial equivalent of the Cox models in `Primary
Analysis/`: We thi models the number
of flare events during follow-up (recurrent events) (see
`Events per individual.R` and `build_event_counts.R` for how those counts
and the endpoint definitions are constructed).

Six exposures are tested, as in thecorresponding `Primary Analysis` file
(HADS, PHQ, PSQI, Exercise, Fatigue, Life Events).

Every exposure is broken down into Full Cohort/Crohn's Disease/Ulcerative
Colitis-IBDU, each further split by patient-reported (soft) vs objective
(hard) flare (the same structure `Primary Analysis`).
Each of those cells fits an unadjusted model (exposure +
`(1 | SiteNo)`) and an adjusted one (+ `Sex + age_decade + Smoke + IMD +
FC`), as both Poisson and negative binomial,
written out explicitly per cell rather than looped. A combined forest plot across
all six exposures follows the individual sections.

Requires the `lme4` and `broom.mixed` packages.

```{r}
#| message: false
source("poisson_helpers.R")

# Only age is needed here (for each exposure's own age-eligibility filter,
# matching Primary Analysis) - Sex/diagnosis2/etc. already come from
# event_counts_soft/hard via poisson_helpers.R.
demographics <- read.xlsx(paste0(data.path, "Baseline2022/demographics2022.xlsx")) %>%
  dplyr::select(ParticipantNo, age)
```

Accumulates every model's tidied rate ratios as we go, for the combined
forest plot at the end.

```{r}
rate_ratios <- tibble::tibble()
```

## HADS (anxiety and depression)

```{r}
hads <- read.xlsx(paste0(data.path, "Baseline2022/hads.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId)) %>%
  dplyr::left_join(demographics, by = "ParticipantNo") %>%
  dplyr::filter(age > 17) %>%
  # Substract 1 from every value (scores are coded 1-4)
  dplyr::mutate(dplyr::across(FeelTense:CanEnjoyBookTV, ~ . - 1)) %>%
  # Reverse encoding
  dplyr::mutate(dplyr::across(
    c(FeelTense, FrightenedFeelingSomethingAwful, WorryingThoughts, FeelCheerful,
      FeelSlowedDown, LostInterestAppearance, FeelRestless, SuddenFeelingsPanic),
    ~ 3 - .
  )) %>%
  # Works out composite score
  dplyr::mutate(
    anxiety_hads = FeelTense + FrightenedFeelingSomethingAwful + WorryingThoughts +
      SitAtEase + FrightenedFeelingButterflies + FeelRestless + SuddenFeelingsPanic,
    depression_hads = EnjoyThings + CanLaugh + FeelCheerful + FeelSlowedDown +
      LostInterestAppearance + LookForward + CanEnjoyBookTV
  )

group_hads <- function(score) {
  cut(score, breaks = c(0, 7, 10, 21), labels = c("0-7", "8-10", "11-21"), include.lowest = TRUE)
}

hads_exposure <- hads %>%
  dplyr::transmute(
    ParticipantNo,
    anxiety_score_group = group_hads(anxiety_hads),
    depression_score_group = group_hads(depression_hads)
  )

event_counts_soft <- event_counts_soft %>% dplyr::left_join(hads_exposure, by = "ParticipantNo")
event_counts_hard <- event_counts_hard %>% dplyr::left_join(hads_exposure, by = "ParticipantNo")
```

### Anxiety

#### Full Cohort

##### Patient-reported flare

```{r}
data_anxiety_soft_full <- event_counts_soft %>%
  dplyr::filter(!is.na(anxiety_score_group))

poisson_anxiety_soft_full_unadj <- lme4::glmer(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_soft_full, family = poisson, control = glmer_control
)
nb_anxiety_soft_full_unadj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_soft_full, control = glmer_control
)
poisson_anxiety_soft_full_adj <- lme4::glmer(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_soft_full, family = poisson, control = glmer_control
)
nb_anxiety_soft_full_adj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_soft_full, control = glmer_control
)

cat("n =", nrow(data_anxiety_soft_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_soft_full_unadj), 1), "/", round(AIC(nb_anxiety_soft_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_soft_full_adj), 1), "/", round(AIC(nb_anxiety_soft_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_anxiety_soft_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_anxiety_soft_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

##### Objective flare

```{r}
data_anxiety_hard_full <- event_counts_hard %>%
  dplyr::filter(!is.na(anxiety_score_group))

poisson_anxiety_hard_full_unadj <- lme4::glmer(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_hard_full, family = poisson, control = glmer_control
)
nb_anxiety_hard_full_unadj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_hard_full, control = glmer_control
)
poisson_anxiety_hard_full_adj <- lme4::glmer(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_hard_full, family = poisson, control = glmer_control
)
nb_anxiety_hard_full_adj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_hard_full, control = glmer_control
)

cat("n =", nrow(data_anxiety_hard_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_hard_full_unadj), 1), "/", round(AIC(nb_anxiety_hard_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_hard_full_adj), 1), "/", round(AIC(nb_anxiety_hard_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_anxiety_hard_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_anxiety_hard_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
data_anxiety_soft_cd <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(anxiety_score_group))

poisson_anxiety_soft_cd_unadj <- lme4::glmer(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_soft_cd, family = poisson, control = glmer_control
)
nb_anxiety_soft_cd_unadj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_soft_cd, control = glmer_control
)
poisson_anxiety_soft_cd_adj <- lme4::glmer(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_soft_cd, family = poisson, control = glmer_control
)
nb_anxiety_soft_cd_adj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_soft_cd, control = glmer_control
)

cat("n =", nrow(data_anxiety_soft_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_soft_cd_unadj), 1), "/", round(AIC(nb_anxiety_soft_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_soft_cd_adj), 1), "/", round(AIC(nb_anxiety_soft_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_anxiety_soft_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_anxiety_soft_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

##### Objective flare

```{r}
data_anxiety_hard_cd <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(anxiety_score_group))

poisson_anxiety_hard_cd_unadj <- lme4::glmer(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_hard_cd, family = poisson, control = glmer_control
)
nb_anxiety_hard_cd_unadj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_hard_cd, control = glmer_control
)
poisson_anxiety_hard_cd_adj <- lme4::glmer(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_hard_cd, family = poisson, control = glmer_control
)
nb_anxiety_hard_cd_adj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_hard_cd, control = glmer_control
)

cat("n =", nrow(data_anxiety_hard_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_hard_cd_unadj), 1), "/", round(AIC(nb_anxiety_hard_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_hard_cd_adj), 1), "/", round(AIC(nb_anxiety_hard_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_anxiety_hard_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_anxiety_hard_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
data_anxiety_soft_uc <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(anxiety_score_group))

poisson_anxiety_soft_uc_unadj <- lme4::glmer(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_soft_uc, family = poisson, control = glmer_control
)
nb_anxiety_soft_uc_unadj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_soft_uc, control = glmer_control
)
poisson_anxiety_soft_uc_adj <- lme4::glmer(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_soft_uc, family = poisson, control = glmer_control
)
nb_anxiety_soft_uc_adj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_soft_uc, control = glmer_control
)

cat("n =", nrow(data_anxiety_soft_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_soft_uc_unadj), 1), "/", round(AIC(nb_anxiety_soft_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_soft_uc_adj), 1), "/", round(AIC(nb_anxiety_soft_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_anxiety_soft_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_anxiety_soft_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

##### Objective flare

```{r}
data_anxiety_hard_uc <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(anxiety_score_group))

poisson_anxiety_hard_uc_unadj <- lme4::glmer(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_hard_uc, family = poisson, control = glmer_control
)
nb_anxiety_hard_uc_unadj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + (1 | SiteNo),
  data = data_anxiety_hard_uc, control = glmer_control
)
poisson_anxiety_hard_uc_adj <- lme4::glmer(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_hard_uc, family = poisson, control = glmer_control
)
nb_anxiety_hard_uc_adj <- lme4::glmer.nb(
  n_events ~ anxiety_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_anxiety_hard_uc, control = glmer_control
)

cat("n =", nrow(data_anxiety_hard_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_hard_uc_unadj), 1), "/", round(AIC(nb_anxiety_hard_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_anxiety_hard_uc_adj), 1), "/", round(AIC(nb_anxiety_hard_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_anxiety_hard_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_anxiety_hard_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "anxiety_score_group")) %>%
    dplyr::mutate(exposure = "Anxiety", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

### Depression

#### Full Cohort

##### Patient-reported flare

```{r}
data_depression_soft_full <- event_counts_soft %>%
  dplyr::filter(!is.na(depression_score_group))

poisson_depression_soft_full_unadj <- lme4::glmer(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_soft_full, family = poisson, control = glmer_control
)
nb_depression_soft_full_unadj <- lme4::glmer.nb(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_soft_full, control = glmer_control
)
poisson_depression_soft_full_adj <- lme4::glmer(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_soft_full, family = poisson, control = glmer_control
)
nb_depression_soft_full_adj <- lme4::glmer.nb(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_soft_full, control = glmer_control
)

cat("n =", nrow(data_depression_soft_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_depression_soft_full_unadj), 1), "/", round(AIC(nb_depression_soft_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_depression_soft_full_adj), 1), "/", round(AIC(nb_depression_soft_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_depression_soft_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_depression_soft_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

##### Objective flare

```{r}
data_depression_hard_full <- event_counts_hard %>%
  dplyr::filter(!is.na(depression_score_group))

poisson_depression_hard_full_unadj <- lme4::glmer(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_hard_full, family = poisson, control = glmer_control
)
nb_depression_hard_full_unadj <- lme4::glmer.nb(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_hard_full, control = glmer_control
)
poisson_depression_hard_full_adj <- lme4::glmer(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_hard_full, family = poisson, control = glmer_control
)
nb_depression_hard_full_adj <- lme4::glmer.nb(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_hard_full, control = glmer_control
)

cat("n =", nrow(data_depression_hard_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_depression_hard_full_unadj), 1), "/", round(AIC(nb_depression_hard_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_depression_hard_full_adj), 1), "/", round(AIC(nb_depression_hard_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_depression_hard_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_depression_hard_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
data_depression_soft_cd <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(depression_score_group))

poisson_depression_soft_cd_unadj <- lme4::glmer(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_soft_cd, family = poisson, control = glmer_control
)
nb_depression_soft_cd_unadj <- lme4::glmer.nb(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_soft_cd, control = glmer_control
)
poisson_depression_soft_cd_adj <- lme4::glmer(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_soft_cd, family = poisson, control = glmer_control
)
nb_depression_soft_cd_adj <- lme4::glmer.nb(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_soft_cd, control = glmer_control
)

cat("n =", nrow(data_depression_soft_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_depression_soft_cd_unadj), 1), "/", round(AIC(nb_depression_soft_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_depression_soft_cd_adj), 1), "/", round(AIC(nb_depression_soft_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_depression_soft_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_depression_soft_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

##### Objective flare

```{r}
data_depression_hard_cd <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(depression_score_group))

poisson_depression_hard_cd_unadj <- lme4::glmer(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_hard_cd, family = poisson, control = glmer_control
)
nb_depression_hard_cd_unadj <- lme4::glmer.nb(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_hard_cd, control = glmer_control
)
poisson_depression_hard_cd_adj <- lme4::glmer(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_hard_cd, family = poisson, control = glmer_control
)
nb_depression_hard_cd_adj <- lme4::glmer.nb(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_hard_cd, control = glmer_control
)

cat("n =", nrow(data_depression_hard_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_depression_hard_cd_unadj), 1), "/", round(AIC(nb_depression_hard_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_depression_hard_cd_adj), 1), "/", round(AIC(nb_depression_hard_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_depression_hard_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_depression_hard_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
data_depression_soft_uc <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(depression_score_group))

poisson_depression_soft_uc_unadj <- lme4::glmer(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_soft_uc, family = poisson, control = glmer_control
)
nb_depression_soft_uc_unadj <- lme4::glmer.nb(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_soft_uc, control = glmer_control
)
poisson_depression_soft_uc_adj <- lme4::glmer(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_soft_uc, family = poisson, control = glmer_control
)
nb_depression_soft_uc_adj <- lme4::glmer.nb(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_soft_uc, control = glmer_control
)

cat("n =", nrow(data_depression_soft_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_depression_soft_uc_unadj), 1), "/", round(AIC(nb_depression_soft_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_depression_soft_uc_adj), 1), "/", round(AIC(nb_depression_soft_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_depression_soft_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_depression_soft_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

##### Objective flare

```{r}
data_depression_hard_uc <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(depression_score_group))

poisson_depression_hard_uc_unadj <- lme4::glmer(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_hard_uc, family = poisson, control = glmer_control
)
nb_depression_hard_uc_unadj <- lme4::glmer.nb(
  n_events ~ depression_score_group + (1 | SiteNo),
  data = data_depression_hard_uc, control = glmer_control
)
poisson_depression_hard_uc_adj <- lme4::glmer(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_hard_uc, family = poisson, control = glmer_control
)
nb_depression_hard_uc_adj <- lme4::glmer.nb(
  n_events ~ depression_score_group + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_depression_hard_uc, control = glmer_control
)

cat("n =", nrow(data_depression_hard_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_depression_hard_uc_unadj), 1), "/", round(AIC(nb_depression_hard_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_depression_hard_uc_adj), 1), "/", round(AIC(nb_depression_hard_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_depression_hard_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_depression_hard_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "depression_score_group")) %>%
    dplyr::mutate(exposure = "Depression", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

## PHQ (somatisation)

```{r}
phq <- read.xlsx(paste0(data.path, "Baseline2022/phq.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId)) %>%
  dplyr::left_join(demographics, by = "ParticipantNo") %>%
  dplyr::filter(age >= 18) %>%
  dplyr::mutate(dplyr::across(StomachPain:TroubleSleeping, ~ . - 1)) %>%
  tidyr::replace_na(list(MenstrualCramps = 0)) %>%
  dplyr::mutate(
    TotalPHQ = StomachPain + BackPain + PainInArms + MenstrualCramps + Headaches +
      ChestPain + Dizziness + FaintingSpells + HeartPound + ShortnessBreath +
      SexualIntercourse + ConstipationDiarrhoea + NauseaIndigestion + FeelingTired + TroubleSleeping
  ) %>%
  dplyr::filter(!is.na(TotalPHQ)) %>%
  dplyr::mutate(somatisation = cut(TotalPHQ, breaks = c(0, 4, 9, 30), labels = c("None", "Mild", "ModSev"), include.lowest = TRUE))

phq_exposure <- phq %>% dplyr::select(ParticipantNo, somatisation)
event_counts_soft <- event_counts_soft %>% dplyr::left_join(phq_exposure, by = "ParticipantNo")
event_counts_hard <- event_counts_hard %>% dplyr::left_join(phq_exposure, by = "ParticipantNo")

```

### Full Cohort

#### Patient-reported flare

```{r}
data_somatisation_soft_full <- event_counts_soft %>%
  dplyr::filter(!is.na(somatisation))

poisson_somatisation_soft_full_unadj <- lme4::glmer(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_soft_full, family = poisson, control = glmer_control
)
nb_somatisation_soft_full_unadj <- lme4::glmer.nb(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_soft_full, control = glmer_control
)
poisson_somatisation_soft_full_adj <- lme4::glmer(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_soft_full, family = poisson, control = glmer_control
)
nb_somatisation_soft_full_adj <- lme4::glmer.nb(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_soft_full, control = glmer_control
)

cat("n =", nrow(data_somatisation_soft_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_soft_full_unadj), 1), "/", round(AIC(nb_somatisation_soft_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_soft_full_adj), 1), "/", round(AIC(nb_somatisation_soft_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_somatisation_soft_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_somatisation_soft_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_somatisation_hard_full <- event_counts_hard %>%
  dplyr::filter(!is.na(somatisation))

poisson_somatisation_hard_full_unadj <- lme4::glmer(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_hard_full, family = poisson, control = glmer_control
)
nb_somatisation_hard_full_unadj <- lme4::glmer.nb(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_hard_full, control = glmer_control
)
poisson_somatisation_hard_full_adj <- lme4::glmer(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_hard_full, family = poisson, control = glmer_control
)
nb_somatisation_hard_full_adj <- lme4::glmer.nb(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_hard_full, control = glmer_control
)

cat("n =", nrow(data_somatisation_hard_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_hard_full_unadj), 1), "/", round(AIC(nb_somatisation_hard_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_hard_full_adj), 1), "/", round(AIC(nb_somatisation_hard_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_somatisation_hard_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_somatisation_hard_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

### Crohn's Disease

#### Patient-reported flare

```{r}
data_somatisation_soft_cd <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(somatisation))

poisson_somatisation_soft_cd_unadj <- lme4::glmer(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_soft_cd, family = poisson, control = glmer_control
)
nb_somatisation_soft_cd_unadj <- lme4::glmer.nb(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_soft_cd, control = glmer_control
)
poisson_somatisation_soft_cd_adj <- lme4::glmer(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_soft_cd, family = poisson, control = glmer_control
)
nb_somatisation_soft_cd_adj <- lme4::glmer.nb(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_soft_cd, control = glmer_control
)

cat("n =", nrow(data_somatisation_soft_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_soft_cd_unadj), 1), "/", round(AIC(nb_somatisation_soft_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_soft_cd_adj), 1), "/", round(AIC(nb_somatisation_soft_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_somatisation_soft_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_somatisation_soft_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_somatisation_hard_cd <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(somatisation))

poisson_somatisation_hard_cd_unadj <- lme4::glmer(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_hard_cd, family = poisson, control = glmer_control
)
nb_somatisation_hard_cd_unadj <- lme4::glmer.nb(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_hard_cd, control = glmer_control
)
poisson_somatisation_hard_cd_adj <- lme4::glmer(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_hard_cd, family = poisson, control = glmer_control
)
nb_somatisation_hard_cd_adj <- lme4::glmer.nb(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_hard_cd, control = glmer_control
)

cat("n =", nrow(data_somatisation_hard_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_hard_cd_unadj), 1), "/", round(AIC(nb_somatisation_hard_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_hard_cd_adj), 1), "/", round(AIC(nb_somatisation_hard_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_somatisation_hard_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_somatisation_hard_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

### Ulcerative Colitis/IBDU

#### Patient-reported flare

```{r}
data_somatisation_soft_uc <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(somatisation))

poisson_somatisation_soft_uc_unadj <- lme4::glmer(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_soft_uc, family = poisson, control = glmer_control
)
nb_somatisation_soft_uc_unadj <- lme4::glmer.nb(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_soft_uc, control = glmer_control
)
poisson_somatisation_soft_uc_adj <- lme4::glmer(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_soft_uc, family = poisson, control = glmer_control
)
nb_somatisation_soft_uc_adj <- lme4::glmer.nb(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_soft_uc, control = glmer_control
)

cat("n =", nrow(data_somatisation_soft_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_soft_uc_unadj), 1), "/", round(AIC(nb_somatisation_soft_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_soft_uc_adj), 1), "/", round(AIC(nb_somatisation_soft_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_somatisation_soft_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_somatisation_soft_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_somatisation_hard_uc <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(somatisation))

poisson_somatisation_hard_uc_unadj <- lme4::glmer(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_hard_uc, family = poisson, control = glmer_control
)
nb_somatisation_hard_uc_unadj <- lme4::glmer.nb(
  n_events ~ somatisation + (1 | SiteNo),
  data = data_somatisation_hard_uc, control = glmer_control
)
poisson_somatisation_hard_uc_adj <- lme4::glmer(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_hard_uc, family = poisson, control = glmer_control
)
nb_somatisation_hard_uc_adj <- lme4::glmer.nb(
  n_events ~ somatisation + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_somatisation_hard_uc, control = glmer_control
)

cat("n =", nrow(data_somatisation_hard_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_hard_uc_unadj), 1), "/", round(AIC(nb_somatisation_hard_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_somatisation_hard_uc_adj), 1), "/", round(AIC(nb_somatisation_hard_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_somatisation_hard_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_somatisation_hard_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "somatisation")) %>%
    dplyr::mutate(exposure = "Somatisation", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

## PSQI (sleep disturbance)

```{r}
psqi <- read.xlsx(paste0(data.path, "Baseline2022/psqi.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantNo)) %>%
  dplyr::left_join(demographics, by = "ParticipantNo") %>%
  dplyr::filter(age >= 18) %>%
  dplyr::filter(!dplyr::if_any(.cols = c(7, 10, 12, 14:23, 25:28), .fns = is.na)) %>%
  dplyr::mutate(dplyr::across(
    c(CantGetToSleep, WakeUpMiddleNight, UseBathroom, CannotBreath, CoughOrSnore,
      FeelTooCold, FeelTooWarm, BadDreams, HavePain, OtherTroubleSleeping,
      SleepQuality, SleepMedicines, StayingAwake, EnoughEnthusiasm),
    ~ . - 1
  )) %>%
  dplyr::rename(comp1 = SleepQuality) %>%
  dplyr::mutate(FallAsleepMns = dplyr::case_when(
    FallAsleepMns <= 14 ~ 0, FallAsleepMns <= 30 ~ 1, FallAsleepMns <= 60 ~ 2, TRUE ~ 3
  )) %>%
  dplyr::mutate(comp2 = dplyr::case_when(
    (FallAsleepMns + CantGetToSleep) %in% c(0) ~ 0,
    (FallAsleepMns + CantGetToSleep) %in% c(1, 2) ~ 1,
    (FallAsleepMns + CantGetToSleep) %in% c(3, 4) ~ 2,
    (FallAsleepMns + CantGetToSleep) %in% c(5, 6) ~ 3
  )) %>%
  tidyr::replace_na(list(SleepEachNightMns = 0)) %>%
  dplyr::mutate(
    TotalSleepMinutes = SleepEachNightHrs * 60 + SleepEachNightMns,
    comp3 = dplyr::case_when(
      TotalSleepMinutes < 300 ~ 3,
      TotalSleepMinutes >= 300 & TotalSleepMinutes <= 359 ~ 2,
      TotalSleepMinutes >= 360 & TotalSleepMinutes <= 419 ~ 1,
      TotalSleepMinutes >= 420 ~ 0
    )
  ) %>%
  # Correcting non-24h clock entries (verbatim from Primary Analysis/PSQI.qmd,
  # including its hardcoded per-participant exceptions/corrections below).
  dplyr::mutate(
    BedTimeHrs = dplyr::case_when(
      ParticipantNo %in% c(110415, 110463) ~ BedTimeHrs,
      BedTimeHrs == 7 ~ 19, BedTimeHrs == 8 ~ 20, BedTimeHrs == 9 ~ 21,
      BedTimeHrs == 10 ~ 22, BedTimeHrs == 11 ~ 23, BedTimeHrs == 12 ~ 00,
      BedTimeHrs == 13 ~ 1, BedTimeHrs == 0 ~ 00, BedTimeHrs == 1 ~ 01,
      BedTimeHrs == 2 ~ 02, BedTimeHrs == 3 ~ 03, BedTimeHrs == 4 ~ 04,
      TRUE ~ BedTimeHrs
    )
  ) %>%
  tidyr::replace_na(list(BedTimeMns = 0, WokenUpMns = 0)) %>%
  # Participant 170004: obvious typo (5h sleep from the reported times) - excluded.
  dplyr::filter(ParticipantNo != 170004) %>%
  # Participants 110351/120042: obvious typos in WokenUpHrs - corrected.
  dplyr::mutate(WokenUpHrs = dplyr::case_when(
    ParticipantNo == 110351 ~ 6, ParticipantNo == 120042 ~ 9, .default = WokenUpHrs
  )) %>%
  dplyr::mutate(
    BedTimeDT = lubridate::make_datetime(hour = BedTimeHrs, min = BedTimeMns),
    WokenUpDT = lubridate::make_datetime(day = 2, hour = WokenUpHrs, min = WokenUpMns)
  ) %>%
  dplyr::mutate(
    BedTimeDT = dplyr::case_when(
      BedTimeHrs >= 0 & BedTimeHrs <= 8 ~ BedTimeDT + lubridate::days(1),
      TRUE ~ BedTimeDT
    )
  ) %>%
  dplyr::mutate(TotalBedMinutes = lubridate::interval(BedTimeDT, WokenUpDT) / lubridate::minutes()) %>%
  dplyr::mutate(SleepEff = TotalSleepMinutes / TotalBedMinutes * 100) %>%
  dplyr::mutate(comp4 = dplyr::case_when(
    SleepEff < 65 ~ 3, SleepEff < 75 ~ 2, SleepEff < 85 ~ 1, SleepEff >= 85 ~ 0
  )) %>%
  dplyr::mutate(TotalTrouble = rowSums(dplyr::select(., WakeUpMiddleNight:OtherTroubleSleeping))) %>%
  dplyr::mutate(comp5 = dplyr::case_when(
    TotalTrouble == 0 ~ 0, TotalTrouble <= 9 ~ 1, TotalTrouble <= 18 ~ 2, TotalTrouble <= 27 ~ 3
  )) %>%
  dplyr::rename(comp6 = SleepMedicines) %>%
  dplyr::mutate(comp7 = dplyr::case_when(
    (StayingAwake + EnoughEnthusiasm) %in% c(0) ~ 0,
    (StayingAwake + EnoughEnthusiasm) %in% c(1, 2) ~ 1,
    (StayingAwake + EnoughEnthusiasm) %in% c(3, 4) ~ 2,
    (StayingAwake + EnoughEnthusiasm) %in% c(5, 6) ~ 3
  )) %>%
  dplyr::mutate(TotalScore = comp1 + comp2 + comp3 + comp4 + comp5 + comp6 + comp7) %>%
  # score > 5 is "poor sleep"
  dplyr::mutate(SleepDisturbance = factor(dplyr::if_else(TotalScore <= 5, "No", "Yes"), levels = c("No", "Yes")))

psqi_exposure <- psqi %>% dplyr::select(ParticipantNo, SleepDisturbance)
event_counts_soft <- event_counts_soft %>% dplyr::left_join(psqi_exposure, by = "ParticipantNo")
event_counts_hard <- event_counts_hard %>% dplyr::left_join(psqi_exposure, by = "ParticipantNo")

```

### Full Cohort

#### Patient-reported flare

```{r}
data_sleep_soft_full <- event_counts_soft %>%
  dplyr::filter(!is.na(SleepDisturbance))

poisson_sleep_soft_full_unadj <- lme4::glmer(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_soft_full, family = poisson, control = glmer_control
)
nb_sleep_soft_full_unadj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_soft_full, control = glmer_control
)
poisson_sleep_soft_full_adj <- lme4::glmer(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_soft_full, family = poisson, control = glmer_control
)
nb_sleep_soft_full_adj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_soft_full, control = glmer_control
)

cat("n =", nrow(data_sleep_soft_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_sleep_soft_full_unadj), 1), "/", round(AIC(nb_sleep_soft_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_sleep_soft_full_adj), 1), "/", round(AIC(nb_sleep_soft_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_sleep_soft_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_sleep_soft_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_sleep_hard_full <- event_counts_hard %>%
  dplyr::filter(!is.na(SleepDisturbance))

poisson_sleep_hard_full_unadj <- lme4::glmer(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_hard_full, family = poisson, control = glmer_control
)
nb_sleep_hard_full_unadj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_hard_full, control = glmer_control
)
poisson_sleep_hard_full_adj <- lme4::glmer(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_hard_full, family = poisson, control = glmer_control
)
nb_sleep_hard_full_adj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_hard_full, control = glmer_control
)

cat("n =", nrow(data_sleep_hard_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_sleep_hard_full_unadj), 1), "/", round(AIC(nb_sleep_hard_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_sleep_hard_full_adj), 1), "/", round(AIC(nb_sleep_hard_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_sleep_hard_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_sleep_hard_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

### Crohn's Disease

#### Patient-reported flare

```{r}
data_sleep_soft_cd <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(SleepDisturbance))

poisson_sleep_soft_cd_unadj <- lme4::glmer(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_soft_cd, family = poisson, control = glmer_control
)
nb_sleep_soft_cd_unadj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_soft_cd, control = glmer_control
)
poisson_sleep_soft_cd_adj <- lme4::glmer(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_soft_cd, family = poisson, control = glmer_control
)
nb_sleep_soft_cd_adj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_soft_cd, control = glmer_control
)

cat("n =", nrow(data_sleep_soft_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_sleep_soft_cd_unadj), 1), "/", round(AIC(nb_sleep_soft_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_sleep_soft_cd_adj), 1), "/", round(AIC(nb_sleep_soft_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_sleep_soft_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_sleep_soft_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_sleep_hard_cd <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(SleepDisturbance))

poisson_sleep_hard_cd_unadj <- lme4::glmer(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_hard_cd, family = poisson, control = glmer_control
)
nb_sleep_hard_cd_unadj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_hard_cd, control = glmer_control
)
poisson_sleep_hard_cd_adj <- lme4::glmer(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_hard_cd, family = poisson, control = glmer_control
)
nb_sleep_hard_cd_adj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_hard_cd, control = glmer_control
)

cat("n =", nrow(data_sleep_hard_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_sleep_hard_cd_unadj), 1), "/", round(AIC(nb_sleep_hard_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_sleep_hard_cd_adj), 1), "/", round(AIC(nb_sleep_hard_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_sleep_hard_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_sleep_hard_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

### Ulcerative Colitis/IBDU

#### Patient-reported flare

```{r}
data_sleep_soft_uc <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(SleepDisturbance))

poisson_sleep_soft_uc_unadj <- lme4::glmer(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_soft_uc, family = poisson, control = glmer_control
)
nb_sleep_soft_uc_unadj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_soft_uc, control = glmer_control
)
poisson_sleep_soft_uc_adj <- lme4::glmer(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_soft_uc, family = poisson, control = glmer_control
)
nb_sleep_soft_uc_adj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_soft_uc, control = glmer_control
)

cat("n =", nrow(data_sleep_soft_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_sleep_soft_uc_unadj), 1), "/", round(AIC(nb_sleep_soft_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_sleep_soft_uc_adj), 1), "/", round(AIC(nb_sleep_soft_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_sleep_soft_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_sleep_soft_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_sleep_hard_uc <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(SleepDisturbance))

poisson_sleep_hard_uc_unadj <- lme4::glmer(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_hard_uc, family = poisson, control = glmer_control
)
nb_sleep_hard_uc_unadj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + (1 | SiteNo),
  data = data_sleep_hard_uc, control = glmer_control
)
poisson_sleep_hard_uc_adj <- lme4::glmer(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_hard_uc, family = poisson, control = glmer_control
)
nb_sleep_hard_uc_adj <- lme4::glmer.nb(
  n_events ~ SleepDisturbance + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_sleep_hard_uc, control = glmer_control
)

cat("n =", nrow(data_sleep_hard_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_sleep_hard_uc_unadj), 1), "/", round(AIC(nb_sleep_hard_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_sleep_hard_uc_adj), 1), "/", round(AIC(nb_sleep_hard_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_sleep_hard_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_sleep_hard_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "SleepDisturbance")) %>%
    dplyr::mutate(exposure = "Sleep disturbance", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

## Exercise (not meeting activity recommendations)

```{r}
exercise <- read.xlsx(paste0(data.path, "Baseline2022/physicalactivity.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId)) %>%
  dplyr::inner_join(demographics, by = "ParticipantNo") %>%
  dplyr::filter(age > 17) %>%
  dplyr::filter(!dplyr::if_all(
    .cols = c(VigorousWork, ModerateWork, WalkBicycle, VigorousSport, ModerateSport),
    .fns = is.na
  )) %>%
  # 6 participants with inconsistent answers, manually identified in
  # Primary Analysis - excluded.
  dplyr::filter(!(ParticipantId %in% c(1395, 4685, 4803, 1452, 4827, 9609))) %>%
  tidyr::replace_na(list(VigorousWork = 2, ModerateWork = 2, WalkBicycle = 2, VigorousSport = 2, ModerateSport = 2)) %>%
  # Exclude >=16h in any one activity per WHO guidance.
  dplyr::filter(
    is.na(VigorousWorkHours) | VigorousWorkHours < 16,
    is.na(ModerateWorkHours) | ModerateWorkHours < 16,
    is.na(WalkBicycleHours) | WalkBicycleHours < 16,
    is.na(VigorousSportHours) | VigorousSportHours < 16,
    is.na(ModerateSportHours) | ModerateSportHours < 16
  ) %>%
  # Correct/exclude cases reporting "yes" to an activity but 0/NA days for it
  # (verbatim per-participant fixes from Primary Analysis/Exercise.qmd).
  dplyr::mutate(ModerateWork = dplyr::case_when(ParticipantId %in% c(4700, 14446) ~ 2, TRUE ~ ModerateWork)) %>%
  dplyr::filter(!(ModerateWork == 1 & (ModerateWorkDays == 0 | is.na(ModerateWorkDays)))) %>%
  dplyr::filter(!(WalkBicycle == 1 & (WalkBicycleDays == 0 | is.na(WalkBicycleDays)))) %>%
  dplyr::mutate(VigorousSport = dplyr::case_when(ParticipantId %in% c(7012, 9555, 9559) ~ 2, TRUE ~ VigorousSport)) %>%
  dplyr::filter(!(VigorousSport == 1 & (VigorousSportDays == 0 | is.na(VigorousSportDays)))) %>%
  dplyr::mutate(ModerateSport = dplyr::case_when(
    ParticipantId %in% c(61, 186, 3550, 7024, 7048, 9735, 12079) ~ 2, TRUE ~ ModerateSport
  )) %>%
  dplyr::filter(!(ModerateSport == 1 & (ModerateSportDays == 0 | is.na(ModerateSportDays)))) %>%
  tidyr::replace_na(list(
    VigorousWorkDays = 0, VigorousWorkHours = 0, VigorousWorkMinutes = 0,
    ModerateWorkDays = 0, ModerateWorkHours = 0, ModerateWorkMinutes = 0,
    WalkBicycleDays = 0, WalkBicycleHours = 0, WalkBicycleMinutes = 0,
    VigorousSportDays = 0, VigorousSportHours = 0, VigorousSportMinutes = 0,
    ModerateSportDays = 0, ModerateSportHours = 0, ModerateSportMinutes = 0
  )) %>%
  dplyr::mutate(
    VigorousWeek = VigorousWorkDays * (VigorousWorkMinutes + (VigorousWorkHours * 60)) +
      VigorousSportDays * (VigorousSportMinutes + (VigorousSportHours * 60)),
    ModerateWeek = ModerateWorkDays * (ModerateWorkMinutes + (ModerateWorkHours * 60)) +
      ModerateSportDays * (ModerateSportMinutes + (ModerateSportHours * 60)) +
      WalkBicycleDays * (WalkBicycleMinutes + (WalkBicycleHours * 60))
  ) %>%
  dplyr::mutate(
    ExerciseMins = VigorousWeek + ModerateWeek,
    METs = (VigorousWeek * 8) + (ModerateWeek * 4)
  ) %>%
  # WHO targets: 75 min vigorous, or 150 min moderate, or 600 METs per week.
  dplyr::mutate(MinimumExercise = dplyr::case_when(
    VigorousWeek >= 75 | ModerateWeek >= 150 | METs >= 600 ~ "Yes", TRUE ~ "No"
  )) %>%
  dplyr::mutate(MinimumExercise = factor(MinimumExercise, levels = c("Yes", "No"))) %>%
  dplyr::filter(ExerciseMins < 6720)

exercise_exposure <- exercise %>% dplyr::select(ParticipantNo, MinimumExercise)
event_counts_soft <- event_counts_soft %>% dplyr::left_join(exercise_exposure, by = "ParticipantNo")
event_counts_hard <- event_counts_hard %>% dplyr::left_join(exercise_exposure, by = "ParticipantNo")

```

### Full Cohort

#### Patient-reported flare

```{r}
data_exercise_soft_full <- event_counts_soft %>%
  dplyr::filter(!is.na(MinimumExercise))

poisson_exercise_soft_full_unadj <- lme4::glmer(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_soft_full, family = poisson, control = glmer_control
)
nb_exercise_soft_full_unadj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_soft_full, control = glmer_control
)
poisson_exercise_soft_full_adj <- lme4::glmer(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_soft_full, family = poisson, control = glmer_control
)
nb_exercise_soft_full_adj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_soft_full, control = glmer_control
)

cat("n =", nrow(data_exercise_soft_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_exercise_soft_full_unadj), 1), "/", round(AIC(nb_exercise_soft_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_exercise_soft_full_adj), 1), "/", round(AIC(nb_exercise_soft_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_exercise_soft_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_exercise_soft_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_exercise_hard_full <- event_counts_hard %>%
  dplyr::filter(!is.na(MinimumExercise))

poisson_exercise_hard_full_unadj <- lme4::glmer(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_hard_full, family = poisson, control = glmer_control
)
nb_exercise_hard_full_unadj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_hard_full, control = glmer_control
)
poisson_exercise_hard_full_adj <- lme4::glmer(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_hard_full, family = poisson, control = glmer_control
)
nb_exercise_hard_full_adj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_hard_full, control = glmer_control
)

cat("n =", nrow(data_exercise_hard_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_exercise_hard_full_unadj), 1), "/", round(AIC(nb_exercise_hard_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_exercise_hard_full_adj), 1), "/", round(AIC(nb_exercise_hard_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_exercise_hard_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_exercise_hard_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

### Crohn's Disease

#### Patient-reported flare

```{r}
data_exercise_soft_cd <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(MinimumExercise))

poisson_exercise_soft_cd_unadj <- lme4::glmer(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_soft_cd, family = poisson, control = glmer_control
)
nb_exercise_soft_cd_unadj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_soft_cd, control = glmer_control
)
poisson_exercise_soft_cd_adj <- lme4::glmer(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_soft_cd, family = poisson, control = glmer_control
)
nb_exercise_soft_cd_adj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_soft_cd, control = glmer_control
)

cat("n =", nrow(data_exercise_soft_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_exercise_soft_cd_unadj), 1), "/", round(AIC(nb_exercise_soft_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_exercise_soft_cd_adj), 1), "/", round(AIC(nb_exercise_soft_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_exercise_soft_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_exercise_soft_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_exercise_hard_cd <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(MinimumExercise))

poisson_exercise_hard_cd_unadj <- lme4::glmer(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_hard_cd, family = poisson, control = glmer_control
)
nb_exercise_hard_cd_unadj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_hard_cd, control = glmer_control
)
poisson_exercise_hard_cd_adj <- lme4::glmer(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_hard_cd, family = poisson, control = glmer_control
)
nb_exercise_hard_cd_adj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_hard_cd, control = glmer_control
)

cat("n =", nrow(data_exercise_hard_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_exercise_hard_cd_unadj), 1), "/", round(AIC(nb_exercise_hard_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_exercise_hard_cd_adj), 1), "/", round(AIC(nb_exercise_hard_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_exercise_hard_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_exercise_hard_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

### Ulcerative Colitis/IBDU

#### Patient-reported flare

```{r}
data_exercise_soft_uc <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(MinimumExercise))

poisson_exercise_soft_uc_unadj <- lme4::glmer(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_soft_uc, family = poisson, control = glmer_control
)
nb_exercise_soft_uc_unadj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_soft_uc, control = glmer_control
)
poisson_exercise_soft_uc_adj <- lme4::glmer(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_soft_uc, family = poisson, control = glmer_control
)
nb_exercise_soft_uc_adj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_soft_uc, control = glmer_control
)

cat("n =", nrow(data_exercise_soft_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_exercise_soft_uc_unadj), 1), "/", round(AIC(nb_exercise_soft_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_exercise_soft_uc_adj), 1), "/", round(AIC(nb_exercise_soft_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_exercise_soft_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_exercise_soft_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_exercise_hard_uc <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(MinimumExercise))

poisson_exercise_hard_uc_unadj <- lme4::glmer(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_hard_uc, family = poisson, control = glmer_control
)
nb_exercise_hard_uc_unadj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + (1 | SiteNo),
  data = data_exercise_hard_uc, control = glmer_control
)
poisson_exercise_hard_uc_adj <- lme4::glmer(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_hard_uc, family = poisson, control = glmer_control
)
nb_exercise_hard_uc_adj <- lme4::glmer.nb(
  n_events ~ MinimumExercise + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_exercise_hard_uc, control = glmer_control
)

cat("n =", nrow(data_exercise_hard_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_exercise_hard_uc_unadj), 1), "/", round(AIC(nb_exercise_hard_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_exercise_hard_uc_adj), 1), "/", round(AIC(nb_exercise_hard_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_exercise_hard_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_exercise_hard_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "MinimumExercise")) %>%
    dplyr::mutate(exposure = "Not meeting activity recommendations", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

## Fatigue

```{r}
fatigue <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx")) %>%
  dplyr::select(ParticipantId, ParticipantNo, OftenLackEnergy) %>%
  dplyr::filter(!is.na(OftenLackEnergy)) %>%
  dplyr::mutate(OftenLackEnergy = dplyr::case_match(OftenLackEnergy, 1 ~ "Yes", 2 ~ "No", 3 ~ "Not sure")) %>%
  # Group "Not sure" with "No" due to low counts.
  dplyr::mutate(OftenLackEnergy = dplyr::case_when(OftenLackEnergy == "Not sure" ~ "No", TRUE ~ OftenLackEnergy)) %>%
  dplyr::mutate(OftenLackEnergy = forcats::fct_relevel(OftenLackEnergy, "No", "Yes")) %>%
  dplyr::left_join(demographics, by = "ParticipantNo") %>%
  dplyr::filter(age >= 18)

fatigue_exposure <- fatigue %>% dplyr::select(ParticipantNo, OftenLackEnergy)
event_counts_soft <- event_counts_soft %>% dplyr::left_join(fatigue_exposure, by = "ParticipantNo")
event_counts_hard <- event_counts_hard %>% dplyr::left_join(fatigue_exposure, by = "ParticipantNo")

```

### Full Cohort

#### Patient-reported flare

```{r}
data_fatigue_soft_full <- event_counts_soft %>%
  dplyr::filter(!is.na(OftenLackEnergy))

poisson_fatigue_soft_full_unadj <- lme4::glmer(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_soft_full, family = poisson, control = glmer_control
)
nb_fatigue_soft_full_unadj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_soft_full, control = glmer_control
)
poisson_fatigue_soft_full_adj <- lme4::glmer(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_soft_full, family = poisson, control = glmer_control
)
nb_fatigue_soft_full_adj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_soft_full, control = glmer_control
)

cat("n =", nrow(data_fatigue_soft_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_soft_full_unadj), 1), "/", round(AIC(nb_fatigue_soft_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_soft_full_adj), 1), "/", round(AIC(nb_fatigue_soft_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_fatigue_soft_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_fatigue_soft_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_fatigue_hard_full <- event_counts_hard %>%
  dplyr::filter(!is.na(OftenLackEnergy))

poisson_fatigue_hard_full_unadj <- lme4::glmer(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_hard_full, family = poisson, control = glmer_control
)
nb_fatigue_hard_full_unadj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_hard_full, control = glmer_control
)
poisson_fatigue_hard_full_adj <- lme4::glmer(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_hard_full, family = poisson, control = glmer_control
)
nb_fatigue_hard_full_adj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_hard_full, control = glmer_control
)

cat("n =", nrow(data_fatigue_hard_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_hard_full_unadj), 1), "/", round(AIC(nb_fatigue_hard_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_hard_full_adj), 1), "/", round(AIC(nb_fatigue_hard_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_fatigue_hard_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_fatigue_hard_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

### Crohn's Disease

#### Patient-reported flare

```{r}
data_fatigue_soft_cd <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(OftenLackEnergy))

poisson_fatigue_soft_cd_unadj <- lme4::glmer(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_soft_cd, family = poisson, control = glmer_control
)
nb_fatigue_soft_cd_unadj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_soft_cd, control = glmer_control
)
poisson_fatigue_soft_cd_adj <- lme4::glmer(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_soft_cd, family = poisson, control = glmer_control
)
nb_fatigue_soft_cd_adj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_soft_cd, control = glmer_control
)

cat("n =", nrow(data_fatigue_soft_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_soft_cd_unadj), 1), "/", round(AIC(nb_fatigue_soft_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_soft_cd_adj), 1), "/", round(AIC(nb_fatigue_soft_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_fatigue_soft_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_fatigue_soft_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_fatigue_hard_cd <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(OftenLackEnergy))

poisson_fatigue_hard_cd_unadj <- lme4::glmer(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_hard_cd, family = poisson, control = glmer_control
)
nb_fatigue_hard_cd_unadj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_hard_cd, control = glmer_control
)
poisson_fatigue_hard_cd_adj <- lme4::glmer(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_hard_cd, family = poisson, control = glmer_control
)
nb_fatigue_hard_cd_adj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_hard_cd, control = glmer_control
)

cat("n =", nrow(data_fatigue_hard_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_hard_cd_unadj), 1), "/", round(AIC(nb_fatigue_hard_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_hard_cd_adj), 1), "/", round(AIC(nb_fatigue_hard_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_fatigue_hard_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_fatigue_hard_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

### Ulcerative Colitis/IBDU

#### Patient-reported flare

```{r}
data_fatigue_soft_uc <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(OftenLackEnergy))

poisson_fatigue_soft_uc_unadj <- lme4::glmer(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_soft_uc, family = poisson, control = glmer_control
)
nb_fatigue_soft_uc_unadj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_soft_uc, control = glmer_control
)
poisson_fatigue_soft_uc_adj <- lme4::glmer(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_soft_uc, family = poisson, control = glmer_control
)
nb_fatigue_soft_uc_adj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_soft_uc, control = glmer_control
)

cat("n =", nrow(data_fatigue_soft_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_soft_uc_unadj), 1), "/", round(AIC(nb_fatigue_soft_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_soft_uc_adj), 1), "/", round(AIC(nb_fatigue_soft_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_fatigue_soft_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_fatigue_soft_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_fatigue_hard_uc <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(OftenLackEnergy))

poisson_fatigue_hard_uc_unadj <- lme4::glmer(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_hard_uc, family = poisson, control = glmer_control
)
nb_fatigue_hard_uc_unadj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + (1 | SiteNo),
  data = data_fatigue_hard_uc, control = glmer_control
)
poisson_fatigue_hard_uc_adj <- lme4::glmer(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_hard_uc, family = poisson, control = glmer_control
)
nb_fatigue_hard_uc_adj <- lme4::glmer.nb(
  n_events ~ OftenLackEnergy + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_fatigue_hard_uc, control = glmer_control
)

cat("n =", nrow(data_fatigue_hard_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_hard_uc_unadj), 1), "/", round(AIC(nb_fatigue_hard_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_fatigue_hard_uc_adj), 1), "/", round(AIC(nb_fatigue_hard_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_fatigue_hard_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_fatigue_hard_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "OftenLackEnergy")) %>%
    dplyr::mutate(exposure = "Fatigue", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

## Life events

```{r}
lifeevents <- read.xlsx(paste0(data.path, "Baseline2022/lifeevents.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId)) %>%
  # Excludes patients who didn't fill out the questionnaire at all.
  dplyr::filter(!is.na(AnyLifeEvents)) %>%
  dplyr::mutate(AnyLifeEvents = factor(AnyLifeEvents, levels = c(1, 2), labels = c("Yes", "No"))) %>%
  dplyr::mutate(AnyLifeEvents = forcats::fct_relevel(AnyLifeEvents, "No", "Yes")) %>%
  dplyr::left_join(demographics, by = "ParticipantNo") %>%
  dplyr::filter(age >= 18)

life_events_exposure <- lifeevents %>% dplyr::select(ParticipantNo, AnyLifeEvents)
event_counts_soft <- event_counts_soft %>% dplyr::left_join(life_events_exposure, by = "ParticipantNo")
event_counts_hard <- event_counts_hard %>% dplyr::left_join(life_events_exposure, by = "ParticipantNo")

```

### Full Cohort

#### Patient-reported flare

```{r}
data_lifeevents_soft_full <- event_counts_soft %>%
  dplyr::filter(!is.na(AnyLifeEvents))

poisson_lifeevents_soft_full_unadj <- lme4::glmer(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_soft_full, family = poisson, control = glmer_control
)
nb_lifeevents_soft_full_unadj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_soft_full, control = glmer_control
)
poisson_lifeevents_soft_full_adj <- lme4::glmer(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_soft_full, family = poisson, control = glmer_control
)
nb_lifeevents_soft_full_adj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_soft_full, control = glmer_control
)

cat("n =", nrow(data_lifeevents_soft_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_soft_full_unadj), 1), "/", round(AIC(nb_lifeevents_soft_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_soft_full_adj), 1), "/", round(AIC(nb_lifeevents_soft_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_lifeevents_soft_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_lifeevents_soft_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Patient-reported flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_lifeevents_hard_full <- event_counts_hard %>%
  dplyr::filter(!is.na(AnyLifeEvents))

poisson_lifeevents_hard_full_unadj <- lme4::glmer(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_hard_full, family = poisson, control = glmer_control
)
nb_lifeevents_hard_full_unadj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_hard_full, control = glmer_control
)
poisson_lifeevents_hard_full_adj <- lme4::glmer(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_hard_full, family = poisson, control = glmer_control
)
nb_lifeevents_hard_full_adj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_hard_full, control = glmer_control
)

cat("n =", nrow(data_lifeevents_hard_full),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_hard_full_unadj), 1), "/", round(AIC(nb_lifeevents_hard_full_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_hard_full_adj), 1), "/", round(AIC(nb_lifeevents_hard_full_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_lifeevents_hard_full_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Unadjusted"),
  broom.mixed::tidy(nb_lifeevents_hard_full_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Objective flare", diagnosis = "Full Cohort", model = "Adjusted")
)
```

### Crohn's Disease

#### Patient-reported flare

```{r}
data_lifeevents_soft_cd <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(AnyLifeEvents))

poisson_lifeevents_soft_cd_unadj <- lme4::glmer(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_soft_cd, family = poisson, control = glmer_control
)
nb_lifeevents_soft_cd_unadj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_soft_cd, control = glmer_control
)
poisson_lifeevents_soft_cd_adj <- lme4::glmer(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_soft_cd, family = poisson, control = glmer_control
)
nb_lifeevents_soft_cd_adj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_soft_cd, control = glmer_control
)

cat("n =", nrow(data_lifeevents_soft_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_soft_cd_unadj), 1), "/", round(AIC(nb_lifeevents_soft_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_soft_cd_adj), 1), "/", round(AIC(nb_lifeevents_soft_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_lifeevents_soft_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_lifeevents_soft_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Patient-reported flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_lifeevents_hard_cd <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "CD", !is.na(AnyLifeEvents))

poisson_lifeevents_hard_cd_unadj <- lme4::glmer(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_hard_cd, family = poisson, control = glmer_control
)
nb_lifeevents_hard_cd_unadj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_hard_cd, control = glmer_control
)
poisson_lifeevents_hard_cd_adj <- lme4::glmer(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_hard_cd, family = poisson, control = glmer_control
)
nb_lifeevents_hard_cd_adj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_hard_cd, control = glmer_control
)

cat("n =", nrow(data_lifeevents_hard_cd),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_hard_cd_unadj), 1), "/", round(AIC(nb_lifeevents_hard_cd_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_hard_cd_adj), 1), "/", round(AIC(nb_lifeevents_hard_cd_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_lifeevents_hard_cd_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Unadjusted"),
  broom.mixed::tidy(nb_lifeevents_hard_cd_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Objective flare", diagnosis = "Crohn's Disease", model = "Adjusted")
)
```

### Ulcerative Colitis/IBDU

#### Patient-reported flare

```{r}
data_lifeevents_soft_uc <- event_counts_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(AnyLifeEvents))

poisson_lifeevents_soft_uc_unadj <- lme4::glmer(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_soft_uc, family = poisson, control = glmer_control
)
nb_lifeevents_soft_uc_unadj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_soft_uc, control = glmer_control
)
poisson_lifeevents_soft_uc_adj <- lme4::glmer(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_soft_uc, family = poisson, control = glmer_control
)
nb_lifeevents_soft_uc_adj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_soft_uc, control = glmer_control
)

cat("n =", nrow(data_lifeevents_soft_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_soft_uc_unadj), 1), "/", round(AIC(nb_lifeevents_soft_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_soft_uc_adj), 1), "/", round(AIC(nb_lifeevents_soft_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_lifeevents_soft_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_lifeevents_soft_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Patient-reported flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

#### Objective flare

```{r}
data_lifeevents_hard_uc <- event_counts_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU", !is.na(AnyLifeEvents))

poisson_lifeevents_hard_uc_unadj <- lme4::glmer(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_hard_uc, family = poisson, control = glmer_control
)
nb_lifeevents_hard_uc_unadj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + (1 | SiteNo),
  data = data_lifeevents_hard_uc, control = glmer_control
)
poisson_lifeevents_hard_uc_adj <- lme4::glmer(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_hard_uc, family = poisson, control = glmer_control
)
nb_lifeevents_hard_uc_adj <- lme4::glmer.nb(
  n_events ~ AnyLifeEvents + Sex + age_decade + Smoke + IMD + FC + (1 | SiteNo),
  data = data_lifeevents_hard_uc, control = glmer_control
)

cat("n =", nrow(data_lifeevents_hard_uc),
    "| Unadjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_hard_uc_unadj), 1), "/", round(AIC(nb_lifeevents_hard_uc_unadj), 1),
    "| Adjusted Poisson/NB AIC:", round(AIC(poisson_lifeevents_hard_uc_adj), 1), "/", round(AIC(nb_lifeevents_hard_uc_adj), 1), "\n")

rate_ratios <- dplyr::bind_rows(
  rate_ratios,
  broom.mixed::tidy(nb_lifeevents_hard_uc_unadj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Unadjusted"),
  broom.mixed::tidy(nb_lifeevents_hard_uc_adj, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(stringr::str_starts(term, "AnyLifeEvents")) %>%
    dplyr::mutate(exposure = "Any life event", outcome = "Objective flare", diagnosis = "Ulcerative Colitis/IBDU", model = "Adjusted")
)
```

## Combined forest plot

All six exposures together, split by flare type and diagnosis group - the
same structure the manuscript uses to report the Cox findings. Blue = 95%
CI excludes 1 (the recurrent-event analogue of a significant association).

```{r}
#| fig-width: 10
#| fig-height: 10
rate_ratios %>%
  # broom.mixed::tidy() names these estimate/conf.low/conf.high - rename to
  # match the rate_ratio/lower/upper used below.
  dplyr::rename(rate_ratio = estimate, lower = conf.low, upper = conf.high) %>%
  dplyr::filter(model == "Adjusted") %>%
  dplyr::mutate(
    label = paste0(exposure, ": ", term),
    significant = lower > 1 | upper < 1,
    # Character "diagnosis" sorts alphabetically (Crohn's Disease, Full
    # Cohort, Ulcerative...) putting Full Cohort in the middle - force it
    # to the top instead.
    diagnosis = factor(diagnosis, levels = c("Full Cohort", "Crohn's Disease", "Ulcerative Colitis/IBDU"))
  ) %>%
  ggplot2::ggplot(ggplot2::aes(x = rate_ratio, y = label, colour = significant)) +
  ggplot2::geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = lower, xmax = upper)) +
  ggplot2::scale_x_log10() +
  ggplot2::scale_colour_manual(values = c(`TRUE` = "#0F8B8D", `FALSE` = "grey60"), guide = "none") +
  ggplot2::facet_grid(diagnosis ~ outcome) +
  ggplot2::labs(x = "Adjusted rate ratio (95% CI, log scale)", y = NULL) +
  ggplot2::theme_minimal()
```

### Associations where the 95% CI excludes 1 (adjusted model)

```{r}
rate_ratios %>%
  dplyr::rename(rate_ratio = estimate, lower = conf.low, upper = conf.high) %>%
  dplyr::filter(model == "Adjusted", lower > 1 | upper < 1) %>%
  dplyr::transmute(
    Exposure = exposure,
    Term = term,
    `Flare type` = outcome,
    `Diagnosis group` = diagnosis,
    `Rate ratio (95% CI)` = sprintf("%.2f (%.2f–%.2f)", rate_ratio, lower, upper)
  ) %>%
  dplyr::arrange(`Diagnosis group`, `Flare type`, Exposure) %>%
  knitr::kable()
```

# Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>

Licensed by <a href="https://creativecommons.org/licenses/by/4.0/">CC BY</a> unless otherwise stated.
