---
title: "Recurrent event analysis with time-varying covariates (Poisson / negative binomial)"
format:
  html:
    code-fold: true
    code-tools: true
toc: true
toc-depth: 4
editor_options:
  chunk_output_type: console
---

## Introduction

Recurrent-event analogue of `Longitudinal analysis/{HADS,PHQ,PSQI,Exercise,
Fatigue,Life Events}/`'s extended Cox models with a time-dependent
covariate. Those files treat each psychosocial measure as time-varying
(last-observation-carried-forward across assessments, via
`survival::tmerge()`/`tdc()`) but only model the *first* flare, via
`Surv(tstart, tstop, endpoint) ~ score_group + ... + frailty(SiteNo)`.

This file keeps the same time-varying-exposure idea but models *all* flare
episodes (from `build_event_counts.R`), via person-period Poisson/negative
binomial regression instead of Cox:

$$Y_{ij} \mid b_i, s_{k(i)} \sim \text{Poisson}(\mu_{ij}), \qquad
\log(\mu_{ij}) = \log(t_{ij}) + \beta_0 + \boldsymbol\beta_1^\top \mathbf{x}_{ij}
+ \boldsymbol\beta_2^\top \mathbf{z}_i + b_i + s_{k(i)}$$

where $Y_{ij}$ is the number of flares in participant $i$'s $j$-th interval
(bounded by their assessment months and a 730.5-day/2-year cutoff, matching
the cutoff used throughout this repo's other flare definitions), $t_{ij}$ is
that interval's length (an offset, not a covariate), $\mathbf{x}_{ij}$ is the
time-varying exposure category active during the interval, $b_i$ is a
per-participant random intercept (playing the same role
`frailty`/`cluster(ParticipantNo)` plays in the Cox version - it's what
makes this valid for *recurrent*, correlated events rather than one
independent row per person), and $s_{k(i)}$ is the per-site random intercept
used everywhere else in this repo. Six exposures are covered (HADS anxiety,
HADS depression, PHQ somatisation, PSQI sleep disturbance, Exercise, Fatigue,
Life Events), each broken down into Full Cohort/Crohn's Disease/Ulcerative
Colitis-IBDU x patient-reported/objective flare, matching `Poisson
Analysis.qmd`'s structure.

**Known simplifications, worth revisiting before treating this as final:**

- **FC held fixed at baseline**, not time-varying - matches every other
  model in this repo (Cox and static Poisson alike), not a deliberate
  choice specific to this file.
- **No withdrawal-date filtering** - same known gap as `build_event_counts.R`.
- **Sex/age/Smoke/IMD/FC/SiteNo come from `build_event_counts.R`'s shared
  `covariates` object throughout**, not from each exposure's own
  `data_cleaning.R` (which computes its own copies) - kept deliberately
  consistent with the rest of this repo's recurrent analysis rather than
  exposure-specific.
- Flare episode dates are pulled directly from `build_event_counts.R`'s
  objects - any future change to that file's endpoint definitions flows
  through here automatically.
- **PHQ/PSQI/Exercise's `data_cleaning.R` write an RDS file** to a shared
  `people/Alex/...` path as a side effect of being sourced (pre-existing
  behaviour of those scripts, not introduced here) - comment those lines out
  locally first if you don't want that write to fire.
- Model-fitting/reporting is done through small shared helper functions
  (`fit_recurrent_tdc`, `run_cell`) rather than every cell being written out
  fully independently (unlike `Poisson Analysis.qmd`'s explicit style) -
  a deliberate trade-off given the combinatorial size here (7 exposure
  scores x 2 flare types x 3 diagnosis groups = 42 cells).

## Setup

```{r}
#| message: false
source("poisson_helpers.R")  # -> event_counts_soft/hard, covariates, glmer_control,
                              # and (transitively) build_event_counts.R's
                              # episode-date objects, demo_tbl, population_cohort

# covariates deliberately excludes diagnosis2 (see build_event_counts.R's own
# comment) - add it back for the diagnosis-subgroup filtering below.
covariates_dx <- covariates %>%
  dplyr::left_join(population_cohort %>% dplyr::select(ParticipantNo, diagnosis2), by = "ParticipantNo")
```

## Flare episode dates

All soft/hard flare dates per participant (not just the first) - the same
objects `build_event_counts.R` uses to build `event_counts_soft/hard`, just
kept at the individual-episode level instead of summed to a count.

```{r}
soft_episode_dates <- dplyr::bind_rows(
  episodes_soft_portal %>% dplyr::select(ParticipantNo, episode_date),
  qflare_episodes,
  imputed_soft %>% dplyr::transmute(ParticipantNo, episode_date = flare_start_date)
)

# eos_objective_episodes is exactly "every hard flare date" (first + further,
# already censored at 730.5 days) - it's only named for its other use as the
# imputation candidate pool in build_event_counts.R.
hard_episode_dates <- eos_objective_episodes %>%
  dplyr::select(ParticipantNo, episode_date = flare_start_date)

entry_dates <- demo_tbl %>% dplyr::select(ParticipantNo, entry_date)
```

## Person-period interval builder

Builds one row per participant per interval between consecutive assessments
of the time-varying exposure (last-observation-carried-forward for any
missed assessment), censored at 730.5 days. `Y` = number of flares landing
in that interval, `t` = interval length in days (the offset). Generic across
exposures - works whether the assessment schedule is fixed (HADS/PHQ/PSQI/
Exercise/Life Events, months 0/12/24) or monthly (Fatigue).

```{r}
build_person_period <- function(cohort, episode_dates, exposure_long, entry_dates, cutoff_days = 730.5) {
  exposure_locf <- exposure_long %>%
    dplyr::semi_join(cohort, by = "ParticipantNo") %>%
    # Guards against a zero-length interval (log(0) in the offset) if a
    # participant has a duplicate row for the same assessment month.
    dplyr::distinct(ParticipantNo, month, .keep_all = TRUE) %>%
    dplyr::arrange(ParticipantNo, month) %>%
    dplyr::group_by(ParticipantNo) %>%
    tidyr::fill(score_group, .direction = "down") %>%
    dplyr::ungroup() %>%
    dplyr::filter(!is.na(score_group)) %>%  # no assessment yet to carry forward
    dplyr::mutate(interval_start = month * 365 / 12) %>%
    dplyr::filter(interval_start < cutoff_days)

  # Next assessment (or the cutoff, for the last one) closes each interval -
  # the same "carried forward until superseded" logic tdc() implements,
  # built directly instead of via tmerge.
  person_period <- exposure_locf %>%
    dplyr::group_by(ParticipantNo) %>%
    dplyr::arrange(interval_start, .by_group = TRUE) %>%
    dplyr::mutate(interval_stop = dplyr::lead(interval_start, default = cutoff_days)) %>%
    dplyr::ungroup() %>%
    dplyr::left_join(entry_dates, by = "ParticipantNo") %>%
    dplyr::mutate(t = interval_stop - interval_start)

  # Count flares landing in each interval (fan-out join: one row per
  # participant-interval-episode combination, then collapse to a count).
  events_per_interval <- person_period %>%
    dplyr::select(ParticipantNo, interval_start, interval_stop, entry_date) %>%
    dplyr::left_join(episode_dates, by = "ParticipantNo") %>%
    dplyr::mutate(
      event_day = as.numeric(as.Date(episode_date) - as.Date(entry_date)),
      in_interval = !is.na(event_day) & event_day >= interval_start & event_day < interval_stop
    ) %>%
    dplyr::group_by(ParticipantNo, interval_start, interval_stop) %>%
    dplyr::summarise(Y = sum(in_interval), .groups = "drop")

  person_period %>%
    dplyr::left_join(events_per_interval, by = c("ParticipantNo", "interval_start", "interval_stop")) %>%
    dplyr::mutate(Y = tidyr::replace_na(Y, 0))
}
```

## Model-fitting and reporting helpers

`run_cell()` fits one exposure/flare-type/diagnosis-group cell, prints the
AIC comparison and tidied NB rate ratios (auto-printed as the chunk's
result), and appends the `score_group` rows to the global `rate_ratios`
accumulator (used by the combined forest plot at the end) - one line per
cell below rather than each of the 42 cells repeating this in full.

```{r}
fit_recurrent_tdc <- function(data) {
  f <- Y ~ score_group + Sex + age_decade + Smoke + IMD + FC +
    (1 | ParticipantNo) + (1 | SiteNo) + offset(log(t))
  list(
    poisson = lme4::glmer(f, data = data, family = poisson, control = glmer_control),
    nb = lme4::glmer.nb(f, data = data, control = glmer_control)
  )
}

# A model "failed to converge" means lme4's own post-fit check flagged it
# (fit@optinfo$conv$lme4$messages non-NULL) - NOT the same as a singular fit
# (boundary (singular) fit), which is a valid degenerate solution (a
# variance component at zero), not a convergence failure.
model_converged <- function(fit) {
  is.null(fit@optinfo$conv$lme4$messages)
}

rate_ratios <- tibble::tibble()

run_cell <- function(pp, dx, exposure, outcome, diagnosis_label) {
  d <- if (dx == "Full") pp else dplyr::filter(pp, diagnosis2 == dx)
  fit <- fit_recurrent_tdc(d)
  poisson_ok <- model_converged(fit$poisson)
  nb_ok <- model_converged(fit$nb)

  cat("n =", nrow(d),
      "| Poisson AIC:", if (poisson_ok) round(AIC(fit$poisson), 1) else "NA (failed to converge)",
      "| NB AIC:", if (nb_ok) round(AIC(fit$nb), 1) else "NA (failed to converge)", "\n\n")

  # Prefers NB (this repo's usual choice); falls back to Poisson if NB
  # failed to converge but Poisson didn't; discards the term entirely if
  # NEITHER converged. model_fit tags which model actually produced the
  # reported rate ratio, so the combined forest plot can show it.
  if (nb_ok) {
    tidy <- fit$nb %>% broom.mixed::tidy(exponentiate = TRUE, conf.int = TRUE) %>% dplyr::mutate(model_fit = "NB")
  } else if (poisson_ok) {
    cat("NB failed to converge - falling back to Poisson for this cell.\n\n")
    tidy <- fit$poisson %>% broom.mixed::tidy(exponentiate = TRUE, conf.int = TRUE) %>% dplyr::mutate(model_fit = "Poisson")
  } else {
    cat("Both Poisson and NB failed to converge for this cell - rate ratio not reported.\n\n")
    return(invisible(NULL))
  }

  rate_ratios <<- dplyr::bind_rows(
    rate_ratios,
    tidy %>%
      dplyr::filter(stringr::str_starts(term, "score_group")) %>%
      dplyr::mutate(exposure = exposure, outcome = outcome, diagnosis = diagnosis_label)
  )
  tidy
}
```

## HADS (anxiety and depression)

Same item recoding and composite-score formulas as
`Longitudinal analysis/HADS/HADS.qmd` - reproduced here rather than sourced,
since that file also builds first-event survival data we don't need.

```{r}
#| message: false
hads_items <- c(
  "FeelTense", "FrightenedFeelingSomethingAwful", "WorryingThoughts", "SitAtEase",
  "FrightenedFeelingButterflies", "FeelRestless", "SuddenFeelingsPanic",
  "EnjoyThings", "CanLaugh", "FeelCheerful", "FeelSlowedDown",
  "LostInterestAppearance", "LookForward", "CanEnjoyBookTV"
)
inverse_coded_items <- c(
  "FeelTense", "FrightenedFeelingSomethingAwful", "WorryingThoughts", "FeelCheerful",
  "FeelSlowedDown", "LostInterestAppearance", "FeelRestless", "SuddenFeelingsPanic"
)

score_hads <- function(df) {
  df %>%
    dplyr::mutate(dplyr::across(tidyselect::all_of(hads_items), ~ . - 1)) %>%
    dplyr::mutate(dplyr::across(tidyselect::all_of(inverse_coded_items), ~ 3 - .)) %>%
    dplyr::mutate(
      anxiety_hads = FeelTense + FrightenedFeelingSomethingAwful + WorryingThoughts +
        SitAtEase + FrightenedFeelingButterflies + FeelRestless + SuddenFeelingsPanic,
      depression_hads = EnjoyThings + CanLaugh + FeelCheerful + FeelSlowedDown +
        LostInterestAppearance + LookForward + CanEnjoyBookTV
    )
}

hads_baseline_raw <- readxl::read_xlsx(paste0(data.path, "Baseline2022/hads.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId))
hads_followup_raw <- readxl::read_xlsx(paste0(data.path, "Followup/hads.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId))

# One row per participant per assessment month (0/12/24) - month 0 from the
# baseline file, months 12/24 (Q_month) from the follow-up file.
hads_long <- dplyr::bind_rows(
  score_hads(hads_baseline_raw) %>% dplyr::transmute(ParticipantNo, month = 0, anxiety_hads, depression_hads),
  score_hads(hads_followup_raw) %>% dplyr::transmute(ParticipantNo, month = Q_month, anxiety_hads, depression_hads)
)

# Same breakpoints as the Cox version.
anxiety_long <- hads_long %>%
  dplyr::transmute(
    ParticipantNo, month,
    score_group = cut(anxiety_hads, breaks = c(0, 7, 21), labels = c("0-7", "8-21"), include.lowest = TRUE)
  )
depression_long <- hads_long %>%
  dplyr::transmute(
    ParticipantNo, month,
    score_group = cut(depression_hads, breaks = c(0, 7, 21), labels = c("0-7", "8-21"), include.lowest = TRUE)
  )

pp_anxiety_soft <- build_person_period(population_cohort, soft_episode_dates, anxiety_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_anxiety_hard <- build_person_period(population_cohort, hard_episode_dates, anxiety_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_depression_soft <- build_person_period(population_cohort, soft_episode_dates, depression_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_depression_hard <- build_person_period(population_cohort, hard_episode_dates, depression_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

### Anxiety

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_anxiety_soft, "Full", "Anxiety", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_anxiety_hard, "Full", "Anxiety", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_anxiety_soft, "CD", "Anxiety", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_anxiety_hard, "CD", "Anxiety", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_anxiety_soft, "UC/IBDU", "Anxiety", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_anxiety_hard, "UC/IBDU", "Anxiety", "Objective flare", "Ulcerative Colitis/IBDU")
```

### Depression

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_depression_soft, "Full", "Depression", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_depression_hard, "Full", "Depression", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_depression_soft, "CD", "Depression", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_depression_hard, "CD", "Depression", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_depression_soft, "UC/IBDU", "Depression", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_depression_hard, "UC/IBDU", "Depression", "Objective flare", "Ulcerative Colitis/IBDU")
```

## PHQ (somatisation)

Sourced directly from `Longitudinal analysis/PHQ/data_cleaning.R` rather
than reproduced by hand - PHQ-15 scoring is compact but there's no benefit
to retyping it. `phq_long` is that file's own long-format output
(`ParticipantNo`, `month`, `somatisation`), used as-is. Note this also
writes an RDS to a shared path as a side effect (see Introduction).

```{r}
#| message: false
source("../Longitudinal analysis/PHQ/data_cleaning.R")

phq_score_long <- phq_long %>%
  dplyr::transmute(ParticipantNo, month, score_group = somatisation)

pp_phq_soft <- build_person_period(population_cohort, soft_episode_dates, phq_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_phq_hard <- build_person_period(population_cohort, hard_episode_dates, phq_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_phq_soft, "Full", "Somatisation", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_phq_hard, "Full", "Somatisation", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_phq_soft, "CD", "Somatisation", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_phq_hard, "CD", "Somatisation", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_phq_soft, "UC/IBDU", "Somatisation", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_phq_hard, "UC/IBDU", "Somatisation", "Objective flare", "Ulcerative Colitis/IBDU")
```

## PSQI (sleep disturbance)

Sourced from `Longitudinal analysis/PSQI/data_cleaning.R` - the 7-component
PSQI score involves bespoke per-participant data corrections (typo fixes,
non-24h clock entries) not worth re-deriving by hand. `psqi` is that file's
own long-format output. Also writes an RDS as a side effect.

```{r}
#| message: false
source("../Longitudinal analysis/PSQI/data_cleaning.R")

psqi_score_long <- psqi %>%
  dplyr::transmute(ParticipantNo, month, score_group = SleepDisturbance)

pp_psqi_soft <- build_person_period(population_cohort, soft_episode_dates, psqi_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_psqi_hard <- build_person_period(population_cohort, hard_episode_dates, psqi_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_psqi_soft, "Full", "Sleep disturbance", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_psqi_hard, "Full", "Sleep disturbance", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_psqi_soft, "CD", "Sleep disturbance", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_psqi_hard, "CD", "Sleep disturbance", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_psqi_soft, "UC/IBDU", "Sleep disturbance", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_psqi_hard, "UC/IBDU", "Sleep disturbance", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Exercise

Sourced from `Longitudinal analysis/Exercise/data_cleaning.R` - extensive
manually-corrected exclusions (implausible hours, inconsistent yes/no
answers) not worth re-deriving by hand. `exercise` is that file's own
long-format output (`MinimumExercise`: meets WHO minimum activity
guidelines). Also writes an RDS as a side effect.

```{r}
#| message: false
source("../Longitudinal analysis/Exercise/data_cleaning.R")

exercise_score_long <- exercise %>%
  dplyr::transmute(ParticipantNo, month, score_group = MinimumExercise)

pp_exercise_soft <- build_person_period(population_cohort, soft_episode_dates, exercise_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_exercise_hard <- build_person_period(population_cohort, hard_episode_dates, exercise_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_exercise_soft, "Full", "Exercise", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_exercise_hard, "Full", "Exercise", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_exercise_soft, "CD", "Exercise", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_exercise_hard, "CD", "Exercise", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_exercise_soft, "UC/IBDU", "Exercise", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_exercise_hard, "UC/IBDU", "Exercise", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Fatigue

Sourced from `Longitudinal analysis/Fatigue/data_cleaning.R`. Unlike the
other exposures, fatigue (`OftenLackEnergy`) is assessed every month via the
monthly follow-up questionnaire, not just at 0/12/24 - `build_person_period()`
handles this the same way regardless, since it derives intervals from
whatever assessment months are actually present.

```{r}
#| message: false
source("../Longitudinal analysis/Fatigue/data_cleaning.R")

fatigue_score_long <- fatigue %>%
  dplyr::transmute(ParticipantNo, month, score_group = OftenLackEnergy)

pp_fatigue_soft <- build_person_period(population_cohort, soft_episode_dates, fatigue_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_fatigue_hard <- build_person_period(population_cohort, hard_episode_dates, fatigue_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_fatigue_soft, "Full", "Fatigue", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_fatigue_hard, "Full", "Fatigue", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_fatigue_soft, "CD", "Fatigue", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_fatigue_hard, "CD", "Fatigue", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_fatigue_soft, "UC/IBDU", "Fatigue", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_fatigue_hard, "UC/IBDU", "Fatigue", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Life Events

Sourced from `Longitudinal analysis/Life Events/data_cleaning.R`. `AnyLifeEvents`
is already binary (No/Yes), no cut-point needed.

```{r}
#| message: false
source("../Longitudinal analysis/Life Events/data_cleaning.R")

lifeevents_score_long <- lifeevents %>%
  dplyr::transmute(ParticipantNo, month, score_group = AnyLifeEvents)

pp_lifeevents_soft <- build_person_period(population_cohort, soft_episode_dates, lifeevents_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_lifeevents_hard <- build_person_period(population_cohort, hard_episode_dates, lifeevents_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_lifeevents_soft, "Full", "Any life event", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_lifeevents_hard, "Full", "Any life event", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_lifeevents_soft, "CD", "Any life event", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_lifeevents_hard, "CD", "Any life event", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_lifeevents_soft, "UC/IBDU", "Any life event", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_lifeevents_hard, "UC/IBDU", "Any life event", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Combined rate ratio forest plot

All seven exposure scores together, split by flare type and diagnosis group
- the same structure `Poisson Analysis.qmd` uses. Uses the NB fit where it
converged; falls back to Poisson (shown as a triangle instead of a circle,
and labelled) for any cell where NB failed to converge but Poisson didn't -
see `run_cell()` in Setup. A cell is absent entirely if neither converged.

```{r}
#| fig-width: 10
#| fig-height: 12
rate_ratios %>%
  dplyr::mutate(
    label = paste0(exposure, ": ", term, " [", model_fit, "]"),
    significant = conf.low > 1 | conf.high < 1,
    diagnosis = factor(diagnosis, levels = c("Full Cohort", "Crohn's Disease", "Ulcerative Colitis/IBDU"))
  ) %>%
  ggplot2::ggplot(ggplot2::aes(x = estimate, y = label, colour = significant, shape = model_fit)) +
  ggplot2::geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = conf.low, xmax = conf.high)) +
  ggplot2::scale_x_log10() +
  ggplot2::scale_colour_manual(values = c(`TRUE` = "#0F8B8D", `FALSE` = "grey60"), guide = "none") +
  ggplot2::scale_shape_manual(values = c(NB = 16, Poisson = 17), name = "Model used") +
  ggplot2::facet_grid(diagnosis ~ outcome) +
  ggplot2::labs(x = "Adjusted rate ratio (95% CI, log scale)", y = NULL) +
  ggplot2::theme_minimal()
```

# Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>
