---
title: "Recurrent event analysis with time-varying covariates - sensitivity analysis (Poisson / negative binomial)"
format:
  html:
    code-fold: true
    code-tools: true
toc: true
toc-depth: 4
editor_options:
  chunk_output_type: console
---

## Introduction

Recurrent-event (Poisson/NB) analogue of `Sensitivity analysis/{HADS,PHQ,
PSQI,Exercise,Fatigue,Life Events}.qmd`'s Cox sensitivity models, the same
way `Poisson Analysis - Time-varying covariates.qmd` is the recurrent
analogue of `Longitudinal analysis/`'s primary Cox models. Structure and
person-period machinery are identical to that file (see it for the
mathematical specification) - this file only adds three covariates found in
the reference Cox sensitivity formula (e.g. `Sensitivity analysis/PSQI.qmd`,
`cox_soft`/`cox_hard`/`cox_soft_uc`/etc.):

```
Surv(time, DiseaseFlareYN) ~ <exposure> + IMD + Sex + age_decade + FC +
  Smoke + flare_group + OverallControl + control_8 + frailty(SiteNo)
```

- **`flare_group`** - patient-reported flare in the previous year
  ("No Flares" vs "1 or More Flares"), from `IBD.xlsx`'s `FlaresInPastYear`.
- **`OverallControl`** - IBD-Control-VAS (the single visual-analogue-scale
  item of the IBD-Control questionnaire).
- **`control_8`** - IBD-Control-8 (the 8-item summed score of the same
  questionnaire).

Both control scores come from `IBD_C.RDS`.

**Two things in the reference sensitivity analysis this file does NOT
replicate:**

- **Multiple imputation (`mice`).** The reference reports both a
  complete-case model and a MICE-pooled model (e.g.
  `cox_results_psqi_cc` vs `cox_results_psqi_mice`, `m = 20` imputations).
  `mice::pool()` doesn't have solid native support for `glmer`/`glmer.nb`
  objects the way it does for `coxph`, and an untested MI-pooling path
  isn't something to guess at. **This file is complete-case only** - rows
  with a missing `flare_group`/`OverallControl`/`control_8` are silently
  dropped by `glmer`'s default `na.action`, same as any other model here.
- **Per-cell spline exploration.** The reference tests `ns(OverallControl,
  df = 2)`/`ns(age_decade, df = 2)` via likelihood-ratio tests in some
  cells. This file uses linear terms throughout, matching how FC/age_decade
  are already treated everywhere else in this repo's Poisson models.

Everything else - the person-period construction, the six exposures, the
Full Cohort/CD/UC x soft/hard breakdown, FC held fixed at baseline, no
withdrawal-date filtering - carries over unchanged from
`Poisson Analysis - Time-varying covariates.qmd`.

## Setup

```{r}
#| message: false
source("poisson_helpers.R")  # -> event_counts_soft/hard, covariates, glmer_control,
                              # and (transitively) build_event_counts.R's
                              # episode-date objects, demo_tbl, population_cohort

# covariates deliberately excludes diagnosis2 (see build_event_counts.R's own
# comment) - add it back for the diagnosis-subgroup filtering below.
covariates_dx <- covariates %>%
  dplyr::left_join(population_cohort %>% dplyr::select(ParticipantNo, diagnosis2), by = "ParticipantNo")
```

## Additional sensitivity covariates

Same `IBD.xlsx`/`IBD_C.RDS` sources and `flare_group` recoding each
exposure's own `data_cleaning.R` already uses internally - built once here
and joined identically across all six exposures, rather than trusting each
`data_cleaning.R`'s own (currently-unused-by-us) copy.

```{r}
#| message: false
sensitivity_covariates <- population_cohort %>%
  dplyr::select(ParticipantNo) %>%
  dplyr::left_join(
    read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx")) %>%
      dplyr::transmute(
        ParticipantNo,
        flare_group = factor(
          dplyr::if_else(FlaresInPastYear == 0, "No Flares", "1 or More Flares"),
          levels = c("No Flares", "1 or More Flares")
        )
      ),
    by = "ParticipantNo"
  ) %>%
  dplyr::left_join(
    readRDS(paste0(chiara, "IBD_C.RDS")) %>%
      dplyr::select(ParticipantNo, OverallControl, control_8),
    by = "ParticipantNo"
  )

covariates_dx <- covariates_dx %>%
  dplyr::left_join(sensitivity_covariates, by = "ParticipantNo")
```

These models show substantially more convergence difficulty than the
primary time-varying analysis - nearly every cell produces a `bobyqa:
maximum number of function evaluations exceeded` warning together with
`Model is nearly unidentifiable: very large eigenvalue`, for both Poisson
and NB. The likely cause is collinearity between `OverallControl` and
`control_8` - two different scorings (single VAS item vs. an 8-item sum) of
the same IBD-Control questionnaire, entered as separate covariates in every
model here. This has not been formally tested (their correlation has not
been checked) and should be verified before the rate ratios reported below
are relied upon beyond the fallback-model mechanism described in
`run_cell()`.

## Flare episode dates

All soft/hard flare dates per participant (not just the first) - the same
objects `build_event_counts.R` uses to build `event_counts_soft/hard`, just
kept at the individual-episode level instead of summed to a count.

```{r}
soft_episode_dates <- dplyr::bind_rows(
  episodes_soft_portal %>% dplyr::select(ParticipantNo, episode_date),
  qflare_episodes,
  imputed_soft %>% dplyr::transmute(ParticipantNo, episode_date = flare_start_date)
)

# eos_objective_episodes is exactly "every hard flare date" (first + further,
# already censored at 730.5 days) - it's only named for its other use as the
# imputation candidate pool in build_event_counts.R.
hard_episode_dates <- eos_objective_episodes %>%
  dplyr::select(ParticipantNo, episode_date = flare_start_date)

entry_dates <- demo_tbl %>% dplyr::select(ParticipantNo, entry_date)
```

## Person-period interval builder

Unchanged from `Poisson Analysis - Time-varying covariates.qmd`.

```{r}
build_person_period <- function(cohort, episode_dates, exposure_long, entry_dates, cutoff_days = 730.5) {
  exposure_locf <- exposure_long %>%
    dplyr::semi_join(cohort, by = "ParticipantNo") %>%
    # Guards against a zero-length interval (log(0) in the offset) if a
    # participant has a duplicate row for the same assessment month.
    dplyr::distinct(ParticipantNo, month, .keep_all = TRUE) %>%
    dplyr::arrange(ParticipantNo, month) %>%
    dplyr::group_by(ParticipantNo) %>%
    tidyr::fill(score_group, .direction = "down") %>%
    dplyr::ungroup() %>%
    dplyr::filter(!is.na(score_group)) %>%  # no assessment yet to carry forward
    dplyr::mutate(interval_start = month * 365 / 12) %>%
    dplyr::filter(interval_start < cutoff_days)

  # Next assessment (or the cutoff, for the last one) closes each interval -
  # the same "carried forward until superseded" logic tdc() implements,
  # built directly instead of via tmerge.
  person_period <- exposure_locf %>%
    dplyr::group_by(ParticipantNo) %>%
    dplyr::arrange(interval_start, .by_group = TRUE) %>%
    dplyr::mutate(interval_stop = dplyr::lead(interval_start, default = cutoff_days)) %>%
    dplyr::ungroup() %>%
    dplyr::left_join(entry_dates, by = "ParticipantNo") %>%
    dplyr::mutate(t = interval_stop - interval_start)

  # Count flares landing in each interval (fan-out join: one row per
  # participant-interval-episode combination, then collapse to a count).
  events_per_interval <- person_period %>%
    dplyr::select(ParticipantNo, interval_start, interval_stop, entry_date) %>%
    dplyr::left_join(episode_dates, by = "ParticipantNo") %>%
    dplyr::mutate(
      event_day = as.numeric(as.Date(episode_date) - as.Date(entry_date)),
      in_interval = !is.na(event_day) & event_day >= interval_start & event_day < interval_stop
    ) %>%
    dplyr::group_by(ParticipantNo, interval_start, interval_stop) %>%
    dplyr::summarise(Y = sum(in_interval), .groups = "drop")

  person_period %>%
    dplyr::left_join(events_per_interval, by = c("ParticipantNo", "interval_start", "interval_stop")) %>%
    dplyr::mutate(Y = tidyr::replace_na(Y, 0))
}
```

## Model-fitting and reporting helpers

Only change from the primary file: `fit_recurrent_tdc()`'s formula adds
`flare_group + OverallControl + control_8`.

```{r}
fit_recurrent_tdc <- function(data) {
  f <- Y ~ score_group + Sex + age_decade + Smoke + IMD + FC +
    flare_group + OverallControl + control_8 +
    (1 | ParticipantNo) + (1 | SiteNo) + offset(log(t))
  list(
    poisson = lme4::glmer(f, data = data, family = poisson, control = glmer_control),
    nb = lme4::glmer.nb(f, data = data, control = glmer_control)
  )
}

# A model "failed to converge" means lme4's own post-fit check flagged it
# (fit@optinfo$conv$lme4$messages non-NULL) - NOT the same as a singular fit
# (boundary (singular) fit), which is a valid degenerate solution (a
# variance component at zero), not a convergence failure.
model_converged <- function(fit) {
  is.null(fit@optinfo$conv$lme4$messages)
}

rate_ratios <- tibble::tibble()

run_cell <- function(pp, dx, exposure, outcome, diagnosis_label) {
  d <- if (dx == "Full") pp else dplyr::filter(pp, diagnosis2 == dx)
  fit <- fit_recurrent_tdc(d)
  poisson_ok <- model_converged(fit$poisson)
  nb_ok <- model_converged(fit$nb)

  cat("n =", nrow(d),
      "| Poisson AIC:", if (poisson_ok) round(AIC(fit$poisson), 1) else "NA (failed to converge)",
      "| NB AIC:", if (nb_ok) round(AIC(fit$nb), 1) else "NA (failed to converge)", "\n\n")

  # Prefers NB (this repo's usual choice); falls back to Poisson if NB
  # failed to converge but Poisson didn't; discards the term entirely if
  # NEITHER converged. model_fit tags which model actually produced the
  # reported rate ratio, so the combined forest plot can show it.
  if (nb_ok) {
    tidy <- fit$nb %>% broom.mixed::tidy(exponentiate = TRUE, conf.int = TRUE) %>% dplyr::mutate(model_fit = "NB")
  } else if (poisson_ok) {
    cat("NB failed to converge - falling back to Poisson for this cell.\n\n")
    tidy <- fit$poisson %>% broom.mixed::tidy(exponentiate = TRUE, conf.int = TRUE) %>% dplyr::mutate(model_fit = "Poisson")
  } else {
    cat("Both Poisson and NB failed to converge for this cell - rate ratio not reported.\n\n")
    return(invisible(NULL))
  }

  rate_ratios <<- dplyr::bind_rows(
    rate_ratios,
    tidy %>%
      dplyr::filter(stringr::str_starts(term, "score_group")) %>%
      dplyr::mutate(exposure = exposure, outcome = outcome, diagnosis = diagnosis_label)
  )
  tidy
}
```

## HADS (anxiety and depression)

Same item recoding and composite-score formulas as
`Longitudinal analysis/HADS/HADS.qmd` - reproduced here rather than sourced,
since that file also builds first-event survival data we don't need.

```{r}
#| message: false
hads_items <- c(
  "FeelTense", "FrightenedFeelingSomethingAwful", "WorryingThoughts", "SitAtEase",
  "FrightenedFeelingButterflies", "FeelRestless", "SuddenFeelingsPanic",
  "EnjoyThings", "CanLaugh", "FeelCheerful", "FeelSlowedDown",
  "LostInterestAppearance", "LookForward", "CanEnjoyBookTV"
)
inverse_coded_items <- c(
  "FeelTense", "FrightenedFeelingSomethingAwful", "WorryingThoughts", "FeelCheerful",
  "FeelSlowedDown", "LostInterestAppearance", "FeelRestless", "SuddenFeelingsPanic"
)

score_hads <- function(df) {
  df %>%
    dplyr::mutate(dplyr::across(tidyselect::all_of(hads_items), ~ . - 1)) %>%
    dplyr::mutate(dplyr::across(tidyselect::all_of(inverse_coded_items), ~ 3 - .)) %>%
    dplyr::mutate(
      anxiety_hads = FeelTense + FrightenedFeelingSomethingAwful + WorryingThoughts +
        SitAtEase + FrightenedFeelingButterflies + FeelRestless + SuddenFeelingsPanic,
      depression_hads = EnjoyThings + CanLaugh + FeelCheerful + FeelSlowedDown +
        LostInterestAppearance + LookForward + CanEnjoyBookTV
    )
}

hads_baseline_raw <- readxl::read_xlsx(paste0(data.path, "Baseline2022/hads.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId))
hads_followup_raw <- readxl::read_xlsx(paste0(data.path, "Followup/hads.xlsx")) %>%
  dplyr::filter(!is.na(ParticipantId))

# One row per participant per assessment month (0/12/24) - month 0 from the
# baseline file, months 12/24 (Q_month) from the follow-up file.
hads_long <- dplyr::bind_rows(
  score_hads(hads_baseline_raw) %>% dplyr::transmute(ParticipantNo, month = 0, anxiety_hads, depression_hads),
  score_hads(hads_followup_raw) %>% dplyr::transmute(ParticipantNo, month = Q_month, anxiety_hads, depression_hads)
)

# Same breakpoints as the Cox version.
anxiety_long <- hads_long %>%
  dplyr::transmute(
    ParticipantNo, month,
    score_group = cut(anxiety_hads, breaks = c(0, 7, 21), labels = c("0-7", "8-21"), include.lowest = TRUE)
  )
depression_long <- hads_long %>%
  dplyr::transmute(
    ParticipantNo, month,
    score_group = cut(depression_hads, breaks = c(0, 7, 21), labels = c("0-7", "8-21"), include.lowest = TRUE)
  )

pp_anxiety_soft <- build_person_period(population_cohort, soft_episode_dates, anxiety_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_anxiety_hard <- build_person_period(population_cohort, hard_episode_dates, anxiety_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_depression_soft <- build_person_period(population_cohort, soft_episode_dates, depression_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_depression_hard <- build_person_period(population_cohort, hard_episode_dates, depression_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

### Anxiety

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_anxiety_soft, "Full", "Anxiety", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_anxiety_hard, "Full", "Anxiety", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_anxiety_soft, "CD", "Anxiety", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_anxiety_hard, "CD", "Anxiety", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_anxiety_soft, "UC/IBDU", "Anxiety", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_anxiety_hard, "UC/IBDU", "Anxiety", "Objective flare", "Ulcerative Colitis/IBDU")
```

### Depression

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_depression_soft, "Full", "Depression", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_depression_hard, "Full", "Depression", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_depression_soft, "CD", "Depression", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_depression_hard, "CD", "Depression", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_depression_soft, "UC/IBDU", "Depression", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_depression_hard, "UC/IBDU", "Depression", "Objective flare", "Ulcerative Colitis/IBDU")
```

## PHQ (somatisation)

Sourced directly from `Longitudinal analysis/PHQ/data_cleaning.R`. `phq_long`
is that file's own long-format output. Also writes an RDS to a shared path
as a side effect (pre-existing behaviour of that script).

```{r}
#| message: false
source("../Longitudinal analysis/PHQ/data_cleaning.R")

phq_score_long <- phq_long %>%
  dplyr::transmute(ParticipantNo, month, score_group = somatisation)

pp_phq_soft <- build_person_period(population_cohort, soft_episode_dates, phq_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_phq_hard <- build_person_period(population_cohort, hard_episode_dates, phq_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_phq_soft, "Full", "Somatisation", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_phq_hard, "Full", "Somatisation", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_phq_soft, "CD", "Somatisation", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_phq_hard, "CD", "Somatisation", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_phq_soft, "UC/IBDU", "Somatisation", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_phq_hard, "UC/IBDU", "Somatisation", "Objective flare", "Ulcerative Colitis/IBDU")
```

## PSQI (sleep disturbance)

Sourced from `Longitudinal analysis/PSQI/data_cleaning.R`. `psqi` is that
file's own long-format output. Also writes an RDS as a side effect.

```{r}
#| message: false
source("../Longitudinal analysis/PSQI/data_cleaning.R")

psqi_score_long <- psqi %>%
  dplyr::transmute(ParticipantNo, month, score_group = SleepDisturbance)

pp_psqi_soft <- build_person_period(population_cohort, soft_episode_dates, psqi_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_psqi_hard <- build_person_period(population_cohort, hard_episode_dates, psqi_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_psqi_soft, "Full", "Sleep disturbance", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_psqi_hard, "Full", "Sleep disturbance", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_psqi_soft, "CD", "Sleep disturbance", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_psqi_hard, "CD", "Sleep disturbance", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_psqi_soft, "UC/IBDU", "Sleep disturbance", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_psqi_hard, "UC/IBDU", "Sleep disturbance", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Exercise

Sourced from `Longitudinal analysis/Exercise/data_cleaning.R`. `exercise` is
that file's own long-format output (`MinimumExercise`: meets WHO minimum
activity guidelines). Also writes an RDS as a side effect.

```{r}
#| message: false
source("../Longitudinal analysis/Exercise/data_cleaning.R")

exercise_score_long <- exercise %>%
  dplyr::transmute(ParticipantNo, month, score_group = MinimumExercise)

pp_exercise_soft <- build_person_period(population_cohort, soft_episode_dates, exercise_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_exercise_hard <- build_person_period(population_cohort, hard_episode_dates, exercise_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_exercise_soft, "Full", "Exercise", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_exercise_hard, "Full", "Exercise", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_exercise_soft, "CD", "Exercise", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_exercise_hard, "CD", "Exercise", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_exercise_soft, "UC/IBDU", "Exercise", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_exercise_hard, "UC/IBDU", "Exercise", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Fatigue

Sourced from `Longitudinal analysis/Fatigue/data_cleaning.R`. Unlike the
other exposures, fatigue (`OftenLackEnergy`) is assessed every month via the
monthly follow-up questionnaire, not just at 0/12/24 - `build_person_period()`
handles this the same way regardless.

```{r}
#| message: false
source("../Longitudinal analysis/Fatigue/data_cleaning.R")

fatigue_score_long <- fatigue %>%
  dplyr::transmute(ParticipantNo, month, score_group = OftenLackEnergy)

pp_fatigue_soft <- build_person_period(population_cohort, soft_episode_dates, fatigue_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_fatigue_hard <- build_person_period(population_cohort, hard_episode_dates, fatigue_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_fatigue_soft, "Full", "Fatigue", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_fatigue_hard, "Full", "Fatigue", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_fatigue_soft, "CD", "Fatigue", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_fatigue_hard, "CD", "Fatigue", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_fatigue_soft, "UC/IBDU", "Fatigue", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_fatigue_hard, "UC/IBDU", "Fatigue", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Life Events

Sourced from `Longitudinal analysis/Life Events/data_cleaning.R`.
`AnyLifeEvents` is already binary (No/Yes), no cut-point needed.

```{r}
#| message: false
source("../Longitudinal analysis/Life Events/data_cleaning.R")

lifeevents_score_long <- lifeevents %>%
  dplyr::transmute(ParticipantNo, month, score_group = AnyLifeEvents)

pp_lifeevents_soft <- build_person_period(population_cohort, soft_episode_dates, lifeevents_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
pp_lifeevents_hard <- build_person_period(population_cohort, hard_episode_dates, lifeevents_score_long, entry_dates) %>%
  dplyr::left_join(covariates_dx, by = "ParticipantNo")
```

#### Full Cohort

##### Patient-reported flare

```{r}
run_cell(pp_lifeevents_soft, "Full", "Any life event", "Patient-reported flare", "Full Cohort")
```

##### Objective flare

```{r}
run_cell(pp_lifeevents_hard, "Full", "Any life event", "Objective flare", "Full Cohort")
```

#### Crohn's Disease

##### Patient-reported flare

```{r}
run_cell(pp_lifeevents_soft, "CD", "Any life event", "Patient-reported flare", "Crohn's Disease")
```

##### Objective flare

```{r}
run_cell(pp_lifeevents_hard, "CD", "Any life event", "Objective flare", "Crohn's Disease")
```

#### Ulcerative Colitis/IBDU

##### Patient-reported flare

```{r}
run_cell(pp_lifeevents_soft, "UC/IBDU", "Any life event", "Patient-reported flare", "Ulcerative Colitis/IBDU")
```

##### Objective flare

```{r}
run_cell(pp_lifeevents_hard, "UC/IBDU", "Any life event", "Objective flare", "Ulcerative Colitis/IBDU")
```

## Combined rate ratio forest plot

All seven exposure scores together, split by flare type and diagnosis group,
adjusted for `flare_group`/`OverallControl`/`control_8` in addition to the
base covariates. Uses the NB fit where it converged; falls back to Poisson
(shown as a triangle instead of a circle, and labelled) for any cell where
NB failed to converge but Poisson didn't - see `run_cell()` in Setup. Given
this file's much heavier convergence trouble (see the note in Setup about
`OverallControl`/`control_8` collinearity), expect more Poisson fallbacks
and more missing cells here than in the primary time-varying file.

```{r}
#| fig-width: 10
#| fig-height: 12
rate_ratios %>%
  dplyr::mutate(
    label = paste0(exposure, ": ", term, " [", model_fit, "]"),
    significant = conf.low > 1 | conf.high < 1,
    diagnosis = factor(diagnosis, levels = c("Full Cohort", "Crohn's Disease", "Ulcerative Colitis/IBDU"))
  ) %>%
  ggplot2::ggplot(ggplot2::aes(x = estimate, y = label, colour = significant, shape = model_fit)) +
  ggplot2::geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  ggplot2::geom_pointrange(ggplot2::aes(xmin = conf.low, xmax = conf.high)) +
  ggplot2::scale_x_log10() +
  ggplot2::scale_colour_manual(values = c(`TRUE` = "#0F8B8D", `FALSE` = "grey60"), guide = "none") +
  ggplot2::scale_shape_manual(values = c(NB = 16, Poisson = 17), name = "Model used") +
  ggplot2::facet_grid(diagnosis ~ outcome) +
  ggplot2::labs(x = "Adjusted rate ratio (95% CI, log scale)", y = NULL) +
  ggplot2::theme_minimal()
```

# Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>
