---
title: "Alcohol"
author:
  - name: "Chiara Cotronei"
    url: https://www.research.ed.ac.uk/en/persons/chiara-cotronei
    affiliations:
      - ref: Lothian
  - name: "Alex Rudge"
    affiliations:
      - ref: CGEM
  - name: "Nathan Constantine-Cooke"
    corresponding: true
    url: https://scholar.google.com/citations?user=2emHWR0AAAAJ&hl=en&oi=ao
    affiliations:
      - ref: CGEM
      - ref: HGU
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-expand: true
---

# Alcohol

## Setting up in R

R Packages

```{r}
#| label: load-libraries
#| echo: false
#| message: false
#| warning: false
set.seed(123)
library(plyr) # Used for mapping values
suppressPackageStartupMessages(library(tidyverse))
library(readxl) # Read in Excel files
library(datefixR) # Standardise dates
library(openxlsx)
library(tidyr)
library(ggdist)
library(rstatix)

suppressPackageStartupMessages(library(table1))
library(knitr)
library(pander)
library(kableExtra)

# Generate flowchart of cohort derivation
library(DiagrammeR)
library(DiagrammeRsvg)

# cox-regression
library(survival)
library(survminer)
library(finalfit)
```

Helper functions

```{r}
# Chi squared tests
summon_chisq_test <- function(data, dependent, independent) {
  
  # Chi square tests/Fisher's exact test for the dependent variable vs each
  # independent variable
  
  purrr::map2_df(
    .x = dependent,
    .y = independent,
    .f = function(.x, .y) {
      x <- data %>%
        dplyr::pull(.x)
      
      y <- data %>%
        dplyr::pull(.y)
      
      # Check whether we need a Fisher test
      fisher_flag <- data %>%
        # Count combinations of x and y
        dplyr::count(!!rlang::sym(.x), !!rlang::sym(.y)) %>%
        # Pull the count
        dplyr::pull(n) %>%
        # Check if any counts are < 5
        min() < 5
        
      if (fisher_flag){
        # If one of the counts is < 5 use Fisher exact test
        fisher_test(table(x, y), simulate.p.value = TRUE, B = 1e5) %>%
          dplyr::mutate(
            dependent = .x, 
            independent = .y,
            method = 'Fisher\'s exact test')
        
      } else {    
        # Otherwise use chi square
        chisq_test(x = x, y = y) %>%
          dplyr::mutate(dependent = .x, independent = .y)
      }
    }
  ) %>%
    dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
    dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
    dplyr::select(-p.signif) %>%
    dplyr::mutate(
      p.signif = dplyr::case_when(
        p.adjust < 0.0001 ~ "****",
        p.adjust < 0.001 ~ "***",
        p.adjust < 0.01  ~ "**",
        p.adjust < 0.05  ~ "*",
        .default = 'ns'
      )
    )
  
}


print_chisq <- function(data) {
  data %>%
    knitr::kable(digits = c(0, 2, 4, 0, NA, NA, NA, 4, NA),
                 col.names = c("n",
                               "Statistic",
                               "p",
                               "df",
                               "Method",
                               "Dependent",
                               "Independent",
                               "Adjusted p-value",
                               "Adjusted significance")
               )
}

print_survival <- function(data) {
  data %>%
    broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    knitr::kable(digits = 3,
                 col.names = c("Covariate",
                               "Estimate",
                               "Std. Error",
                               "Statistic",
                               "p value",
                               "Lower CI",
                               "Upper CI")
                 )
}


# Creating proportions for plotting baseline data
calc_proportion <- function(data, dependent, independent){
  
  data %>%
    dplyr::group_by(!!rlang::sym(independent), !!rlang::sym(dependent)) %>%
    dplyr::count() %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!rlang::sym(independent)) %>%
    dplyr::mutate(p = n/sum(n)) 
  
}

summon_baseline_plots <- function(data, dependent, independent) {
  
  # Chi squared tests
  chisq_results <- summon_chisq_test(
    data = data,
    dependent = dependent,
    independent = independent
  ) %>%
    dplyr::transmute(independent, p.adjust) %>%
    {setNames(as.list(.$p.adjust), .$independent)}

  
  # Creating plots for baseline variables with the dependent variable
  
  plot_list <- list()
  
  for (y in independent) {
    
    # Create a plot for each independent variable
    
    plot <- data %>%
      calc_proportion(., dependent = dependent, independent = y) %>%
      ggplot(aes(x = .data[[dependent]], y = p, fill = .data[[y]])) +
      geom_col(position = 'dodge') +
      annotate(
        "text",
        label = paste0("Adjusted p-value: ", chisq_results[[y]]),
        x = Inf,
        y = Inf,
        vjust = 2,
        hjust = 1.3
      ) +
      theme_minimal()
    
    plot_list[[y]] <- plot
    
  }
  # Return the plots
  plot_list
}
```

Load data

```{r}
# paths to PREdiCCt data
if (file.exists("/.dockerenv")) {
  # If running in docker
  data.path <- "data/final/20221004/"
  redcap.path <- "data/final/20231030/"
  prefix <- "data/end-of-follow-up/"
  outdir <- "data/processed"
  flare.dir <- "data/final/20240308/Followup/"
  chiara <- "data/people/chiara/"
} else {
  # Run on OS directly
  data.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20221004/"
  redcap.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20231030/"
  prefix <- "/Volumes/igmm/cvallejo-predicct/predicct/end-of-follow-up/"
  outdir <- "/Volumes/igmm/cvallejo-predicct/predicct/processed/"
  flare.dir <- paste0("/Volumes/igmm/cvallejo-predicct/predicct/",
                      "final/20240308/Followup/")
  chiara <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
}

# read-in data from multiple sources
physicalactivity <- read.xlsx(paste0(
  data.path,
  "Baseline2022/physicalactivity.xlsx"
))

demographics <- read.xlsx(paste0(
  data.path,
  "Baseline2022/demographics2022.xlsx"
))

IBD <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx"))

ffq <- read.xlsx(
  paste0(
    prefix,
    "predicct ffq_nutrientfood groupDQI all foods_data (n1092)Nov2022.",
    "xlsx"
  )
)

all_flares <- read.xlsx(paste0(flare.dir, "all-flares.xlsx"))
demo <- readRDS(paste0(outdir, "demo.RDS"))
flares <- readRDS(paste0(outdir, "flares/cutoff-flares.RDS"))
smoking <- readRDS(paste0(chiara, "smoking.rds"))
demo_uc <- readRDS(paste0(outdir, "demo-uc.RDS"))
demo_cd <- readRDS(paste0(outdir, "demo-cd.RDS"))
flares_soft <- readRDS(paste0(chiara, "flares_soft.RDS"))
flares_hard <- readRDS(paste0(chiara, "flares_hard.RDS"))
```

## Data Cleaning

```{r}
# Alcohol data
ETOH <- ffq %>%
  dplyr::select(participantno,
         Alcoholg,
         Alcohol_percEng,
         Alcohol_units,
         Alcohol_units_wk,
         alcohol_units_cat)

ETOH <- ETOH %>%
  dplyr::rename(ParticipantNo = participantno)

ETOH %<>%
  dplyr::mutate(ParticipantNo = as.character(ParticipantNo))

ETOH <- ETOH %>%
  left_join(demographics %>%
              select(ParticipantNo, Sex, age, SiteNo, diagnosis) ,
            by = "ParticipantNo")

# Recode diagnosis
ETOH %<>%
  dplyr::mutate(
    diagnosis2 = dplyr::case_when(
      diagnosis == 1 ~ 1,
      diagnosis == 2 ~ 2,
      diagnosis == 3 ~ 2,
      diagnosis == 4 ~ 2
    )
  ) %>%
  dplyr::mutate(diagnosis2 = factor(
    diagnosis2,
    levels = c("1", "2"),
    labels = c("CD", "UC/IBDU")
  ))

# exclude <18 yo = 6 participants excluded
ETOH <- ETOH %>%
  dplyr::filter(age >= 18)

# make sex into level
ETOH %<>%
  dplyr::mutate(Sex = factor(
    Sex, 
    levels = c(1, 2), 
    labels = c("Male", "Female"))
  )

ETOH %<>%
  dplyr::mutate(AgeGroup = cut(
    age,
    breaks = c(18, 24, 34, 44, 54, 65, Inf),
    labels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+"),
    include.lowest = TRUE
))


ETOH <- ETOH %>%
  left_join(
    IBD %>% select(ParticipantNo, FlaresInPastYear),
    by = 'ParticipantNo')

# 31 have missing flare from last year data
# ETOH %>%
#   filter(is.na(FlaresInPastYear)) %>%
#   nrow()

ETOH <- ETOH %>%
  filter(!is.na(FlaresInPastYear))

# create flare groups
ETOH <- ETOH %>%
  mutate(flare_group = ifelse(FlaresInPastYear == 0,
                              "No Flares",
                              "1 or More Flares")
         )

# make flares into levels
ETOH %<>%
  dplyr::mutate(flare_group = factor(flare_group, 
                                     levels = c("No Flares", "1 or More Flares")
                                     )
  )

# add FC
ETOH <- ETOH %>%
  left_join(
    demo %>% select(ParticipantNo, FC, cat, Biologic, IMD),
    by = 'ParticipantNo')

# n.b.78 fc cat missing
#ETOH %>%
#  filter(is.na(cat)) %>%
#  nrow()

ETOH <- ETOH %>%
  filter(!is.na(cat))


# create new ETOH groups
ETOH <- ETOH %>%
  mutate(
    weekly_units = case_when(
      Alcohol_units_wk <= 0.1 ~ "0-0.1",
      Alcohol_units_wk > 0.1 & Alcohol_units_wk <= 14 ~ "0.1-14",
      Alcohol_units_wk > 14 ~ ">14"
    )
  )

ETOH %<>%
  dplyr::mutate(weekly_units = factor(
    weekly_units,
    levels = c("0-0.1", "0.1-14", ">14")
))

# create alcohol intake level where:
# 1 = none/very rarely;
# 2 = within recommended limits;
# 3 = above recommended limits

ETOH <- ETOH %>%
  mutate(
    AlcoholIntake = case_when(
      Alcohol_units_wk <= 0.1 ~ "1",
      Alcohol_units_wk > 0.1 & Alcohol_units_wk <= 14 ~ "2",
      Alcohol_units_wk > 14 ~ "3"
    )
  )

# Any missing units data?
# ETOH %>%
#   filter(is.na(weekly_units)) %>%
#   nrow()

ETOH <- ETOH %>%
  filter(!is.na(weekly_units))

```

## Baseline Analysis

### Full Cohort

```{r}
# Rename ETOH as data_baseline
data_baseline <- ETOH

# Independent variables
independent = c('diagnosis2',
                'AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is weekly_units

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline, 
  dependent = 'weekly_units', 
  independent = independent) %>%
  print_chisq()

# Plots
baseline_plots <- summon_baseline_plots(
  data = data_baseline, 
  dependent = 'weekly_units',
  independent = independent
  )

baseline_plots$diagnosis2
baseline_plots$AgeGroup
baseline_plots$Sex
baseline_plots$flare_group
baseline_plots$cat

```

### Ulcerative Colitis

```{r}
# UC patients only
data_baseline_uc <- ETOH %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is weekly_units

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_uc, 
  dependent = 'weekly_units', 
  independent = independent) %>%
  print_chisq() 

# Plots
baseline_plots_uc <- summon_baseline_plots(
  data = data_baseline_uc, 
  dependent = 'weekly_units',
  independent = independent
  )

baseline_plots_uc$AgeGroup
baseline_plots_uc$Sex
baseline_plots_uc$flare_group
baseline_plots_uc$cat 
```

### Crohn's Disease

```{r}
# CD patients only
data_baseline_cd <- ETOH %>%
  dplyr::filter(diagnosis2 == 'CD')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is weekly_units

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_cd, 
  dependent = 'weekly_units', 
  independent = independent) %>%
  print_chisq() 

# Plots
baseline_plots_cd <- summon_baseline_plots(
  data = data_baseline_cd, 
  dependent = 'weekly_units',
  independent = independent
  )

baseline_plots_cd$AgeGroup
baseline_plots_cd$Sex
baseline_plots_cd$flare_group
baseline_plots_cd$cat
```

## Survival Analysis

### Full Cohort

```{r}
# Survival data
# Patient reported flares
data_survival_soft <- data_baseline %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard <- data_baseline %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
#| warning: false
# Soft
# Kaplan Meier
data_survival_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ weekly_units, data = .) %>%
  ggsurvplot(data = data_survival_soft,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    weekly_units +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft
)

cox_soft %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ weekly_units, data = .) %>%
  ggsurvplot(data = data_survival_hard,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    weekly_units +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard
)

cox_hard %>%
  print_survival()
```

### Ulcerative Colitis

```{r}
# Survival data
# Patient reported flares
data_survival_soft_uc <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard_uc <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Soft
# Kaplan Meier
data_survival_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ weekly_units, data = .) %>%
  ggsurvplot(data = data_survival_soft_uc,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    weekly_units +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft_uc
)

cox_soft_uc %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ weekly_units, data = .) %>%
  ggsurvplot(data = data_survival_hard_uc,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    weekly_units +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

cox_hard_uc %>%
  print_survival()
```

### Crohn's Disease

```{r}
# Survival data
# Patient reported flares
data_survival_soft_cd <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard_cd <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Soft
# Kaplan Meier
data_survival_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ weekly_units, data = .) %>%
  ggsurvplot(data = data_survival_soft_cd,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    weekly_units +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft_cd
)

cox_soft_cd %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ weekly_units, data = .) %>%
  ggsurvplot(data = data_survival_hard_cd,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    weekly_units +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard_cd
)

cox_hard_cd %>%
  print_survival()
```

## Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>

Licensed by <a href="https://creativecommons.org/licenses/by/4.0/">CC BY</a> unless otherwise stated.
