---
title: "Exercise"
author:
  - name: "Chiara Cotronei"
    url: https://www.research.ed.ac.uk/en/persons/chiara-cotronei
    affiliations:
      - ref: Lothian
  - name: "Alexander Rudge"
    affiliations:
      - ref: CGEM
  - name: "Nathan Constantine-Cooke"
    corresponding: true
    url: https://scholar.google.com/citations?user=2emHWR0AAAAJ&hl=en&oi=ao
    affiliations:
      - ref: CGEM
      - ref: HGU
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-expand: true
---

# Exercise

## Setting up in R

```{r}
#| label: load-libraries
#| echo: false
#| message: false
#| warning: false
set.seed(123)
library(plyr) # Used for mapping values
suppressPackageStartupMessages(library(tidyverse))
library(readxl) # Read in Excel files
library(datefixR) # Standardise dates
library(openxlsx)
library(tidyr)
library(ggdist)
library(rstatix)

suppressPackageStartupMessages(library(table1))
library(knitr)
library(pander)
library(kableExtra)

# Generate flowchart of cohort derivation
library(DiagrammeR)
library(DiagrammeRsvg)

# cox-regression
library(survival)
library(survminer)
library(finalfit)

```

## Helper functions
```{R}

summon_statistical_test <- function(data, dependent, independent) {
  
  # Chi square tests/Fisher's exact test/Wilcox or Mann Whitney
  # for the dependent variable vs each independent variable
  # Dependent variables must be discrete
  
  purrr::map2_df(
    .x = independent,
    .y = dependent,
    .f = function(.x, .y) {
      
      # Remove NA
      data %<>%
        dplyr::filter(
          !is.na(.x),
          !is.na(.y)
        )
      
      x <- data %>%
        dplyr::pull(.x)
      
      y <- data %>%
        dplyr::pull(.y)
      
      # Check if x is continuous (numeric) or not
      continuous_flag <- is.numeric(x)
      
      # If continuous, use a Wilcox/Mann Whitney
      if (continuous_flag) {
        # Formula as score ~ group
        test_formula <- as.formula(paste0(.x, " ~ ", .y))
        
        wilcox_test(data, formula = test_formula) %>%
          # Don't need the adjusted p values as we adjust ourselves in a second
          dplyr::select(-tidyselect::any_of(c('.y.', 'p.adj', 'p.adj.signif'))) %>%
          dplyr::mutate(independent = .x,
                        dependent = .y,
                        method = 'Wilcox test')
        
      } else { # Otherwise use categorical tests
        
        # Check whether we need a Fisher test
        fisher_flag <- data %>%
          # Count combinations of x and y
          dplyr::count(!!rlang::sym(.x), !!rlang::sym(.y)) %>%
          # Pull the count
          dplyr::pull(n) %>%
          # Check if any counts are < 5
          min() < 5
        
        if (fisher_flag) {
          # If one of the counts is < 5 use Fisher exact test
          fisher_test(table(x, y), simulate.p.value = TRUE, B = 1e5) %>%
            dplyr::mutate(independent = .x,
                          dependent = .y,
                          method = 'Fisher\'s exact test')
          
        } else {
          # Otherwise use chi square
          chisq_test(x = x, y = y) %>%
            dplyr::mutate(independent = .x, dependent = .y)
        }
      }
    }
  ) %>%
    dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
    dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
    dplyr::select(-p.signif) %>%
    dplyr::mutate(
      p.adjust.signif = dplyr::case_when(
        p.adjust < 0.0001 ~ "****",
        p.adjust < 0.001 ~ "***",
        p.adjust < 0.01  ~ "**",
        p.adjust < 0.05  ~ "*",
        .default = 'ns'
      )
    ) %>%
    # Calculate n for Wilcox tests
    dplyr::mutate(n = dplyr::coalesce(n, n1 + n2)) %>%
    # Rename group 1 and group2
    dplyr::rename(
      independent_group1 = group1, 
      independent_group2 = group2,
      n_group1 = n1,
      n_group2 = n2) %>%
    # Reorder columns
    dplyr::select(
      dependent, independent, method, independent_group1, independent_group2, n, n_group1, n_group2, statistic, df, p, p.adjust, p.adjust.signif
    )
  
}


# Function to create plots for baseline associations

summon_baseline_plot_discrete <- function(data, stat_tests, dependent, independent) {
  # Creating plot for baseline variables with the dependent variable
  # Discrete variables
  
  # Stat_tests is a dataframe from summon_statistical_test to get p-value info
  
  # Rename independent for filtering
  indep <- independent
  
  label <- stat_tests %>%
    dplyr::filter(independent == indep) %>%
    dplyr::pull(p.adjust) %>%
    {paste0("Adjusted p-value: ", .)}
  
  data %>%
    dplyr::filter(!is.na(!!sym(dependent)), !is.na(!!sym(independent))) %>%
    dplyr::group_by(!!rlang::sym(independent), !!rlang::sym(dependent)) %>%
    dplyr::count() %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!rlang::sym(independent)) %>%
    dplyr::mutate(p = n/sum(n)) %>%
    dplyr::ungroup() %>%
    ggplot(aes(x = !!sym(dependent), y = p, fill = !!sym(independent))) +
    geom_col(position = 'dodge') +
    annotate(
      "text",
      label = label,
      x = Inf,
      y = Inf,
      vjust = 2,
      hjust = 1.3
    ) +
    theme_minimal()
  
}


summon_baseline_plot_continuous <- function(data, stat_tests, dependent, independent) {
  # Creating plot for baseline variables with the dependent variable
  # Continuous independent variable
  
  # Stat_tests is a dataframe from summon_statistical_test to get p-value info
  
  # Label for the plot needs multiple tests on it
  # Rename independent for filtering
  indep <- independent

  label <- stat_tests %>%
    dplyr::filter(independent == indep) %>%
    dplyr::mutate(label = paste0(independent_group1, " vs ", independent_group2, ": ", p.adjust)) %>%
    dplyr::pull(label) %>%
    paste0(., collapse="\n") %>%
    {paste0("Adjusted p-values:\n", .)}
  
  # Plot
  data %>%
    dplyr::filter(
      !is.na(!!sym(dependent)),
      !is.na(!!sym(independent))
    ) %>%
    ggplot(aes(x = !!sym(independent), colour = !!sym(dependent))) +
    geom_density() +
    annotate(
      "text",
      label = label,
      x = Inf,
      y = Inf,
      vjust = 1.1,
      hjust = 1.3
    ) +
    ylab("") +
    theme_minimal()
  
}

# Function to perform coxph with mice for missing data
coxph_mice <- function(formula, data, ...){
  
  # Variables
  variables <- all.vars(formula)
  
  # Extract time and status variable names
  timevar <- variables[1]
  statusvar <- variables[2]
  
  # Select variables in the data
  data %<>%
    dplyr::select(tidyselect::all_of(variables)) %>%
    # Calculate the cumulative hazard
    dplyr::mutate(cumhaz = mice::nelsonaalen(
      data = .,
      timevar = !!sym(timevar),
      statusvar = !!sym(statusvar)
    ))
  
  # Predictor matrix - need to exclude time from the model
  pred_matrix <- mice::make.predictorMatrix(data)
  
  pred_matrix[, timevar] <- 0
  
  data_imp <- mice::mice(
    data = data,
    predictorMatrix = pred_matrix, ...)
  
  # Return the imputation
  return(data_imp)
  
}

print_chisq <- function(data) {
  data %>%
    knitr::kable(digits = c(0, 2, 4, 0, NA, NA, NA, 4, NA),
                 col.names = c("n",
                               "Statistic",
                               "p",
                               "df",
                               "Method",
                               "Dependent",
                               "Independent",
                               "Adjusted p-value",
                               "Adjusted significance")
               )
}

print_survival <- function(data) {
  data %>%
    broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    knitr::kable(digits = 3,
                 col.names = c("Covariate",
                               "Estimate",
                               "Std. Error",
                               "Statistic",
                               "p value",
                               "Lower CI",
                               "Upper CI")
                 )
}

plot_continuous_hr <- function(data, model, variable, splineterm = NULL){
  
  # Plotting continuous Hazard Ratio curves
  # Data is the same data used to fit the Cox model
  # model is a Cox model
  # variable is the variable we are plotting
  # splineterm is the spline used in the model, if there is one
  
  
  # Range of the variable to plot
  variable_range <- data %>%
    dplyr::pull(variable) %>%
    quantile(., probs = seq(0, 0.9, 0.01), na.rm = TRUE) %>%
    unname() %>%
    unique()
  
  # If there is a spline
  if (!is.null(splineterm)){
    
    # Recreate the spline
    spline <- with(data, eval(parse(text = splineterm)))
    
    # Get the spline terms for the variable values we are plotting
    spline_values <- predict(spline, newx = variable_range)
    
    # Get the reference values of the variables that were used to fit the Cox model
    X_ref <- model$means
    
    # New values to calculate the HR at 
    # Take the ref values and vary our variable of interest
    
    # Number of values
    n_values <- variable_range %>% length()
    
    X_ref <- matrix(rep(X_ref, n_values), nrow = n_values, byrow = TRUE)
    # Fix the column names
    colnames(X_ref) <- names(model$means)
    
    # Now add our new variable values
    X_new <- X_ref
    # Identify the correct columns
    column_pos <- stringr::str_detect(colnames(X_new), variable)
    # Add the values - ordering of columns and model term names should be preserved
    X_new[,column_pos] <- spline_values
    
    # Calculate the linear predictor and se
    # Method directly from predict.coxph from the survival package
    lp <- (X_new - X_ref)%*%coef(model) 
    
    lp_se <- rowSums((X_new - X_ref)%*%vcov(model)*(X_new - X_ref)) %>%
      as.numeric() %>%
      sqrt()
    
  } else {
    # If there is no spline then it is easier
    
    # Get the reference values of the variables that were used to fit the Cox model
    X_ref <- model$means
    
    # New values to calculate the HR at 
    # Take the ref values and vary our variable of interest
    
    # Number of values
    n_values <- variable_range %>% length()
    
    X_ref <- matrix(rep(X_ref, n_values), nrow = n_values, byrow = TRUE)
    # Fix the column names
    colnames(X_ref) <- names(model$means)
    
    # Now add our new variable values
    X_new <- X_ref
    
    # Add the values
    X_new[,variable] <- variable_range
    
    # Calculate the linear predictor and se
    # Method directly from predict.coxph from the survival package
    lp <- (X_new - X_ref)%*%coef(model) %>% as.numeric() 
    
    lp_se <- rowSums((X_new - X_ref)%*%vcov(model)*(X_new - X_ref)) %>% 
      as.numeric() %>%
      sqrt()
    
  }
  
  # Store results in a tibble
  data_pred <- tibble(
    !!sym(variable) := variable_range,
    p = lp,
    se = lp_se
  )
  
  # Plot
  data_pred %>%
    ggplot(aes(
      x = !!rlang::sym(variable),
      y = exp(p),
      ymin = exp(p - 1.96 * se),
      ymax = exp(p + 1.96 * se)
    )) +
    geom_point() +
    geom_line() +
    geom_ribbon(alpha = 0.2) +
    ylab("HR") + 
    theme_minimal()
  
}

# Function to perform an LRT
summon_lrt <- function(model, remove = NULL, add = NULL) {
  # Function that removes a term and performs a
  # Likelihood ratio test to determine its significance
  
  new_formula <- "~ ."
  
  if (!is.null(remove)) {
    new_formula <- paste0(new_formula, " - ", remove)
  } else {
    remove <- NA
  }
  
  if (!is.null(add)) {
    new_formula <- paste0(new_formula, " + ", add)
  } else {
    add <- NA
  }
  
  update(model, new_formula) %>%
    anova(model, ., test = "LRT") %>%
    broom::tidy() %>%
    dplyr::filter(!is.na(p.value)) %>%
    dplyr::select(-term) %>%
    dplyr::mutate(
      removed = remove,
      added = add
    ) %>%
    dplyr::mutate(test = "LRT") %>%
    dplyr::select(test, removed, added, tidyselect::everything())
}

```

## Load data

```{r}
# paths to PREdiCCt data
if (file.exists("/.dockerenv")) {
  # If running in docker
  data.path <- "data/final/20221004/"
  redcap.path <- "data/final/20231030/"
  prefix <- "data/end-of-follow-up/"
  outdir <- "data/processed"
  flare.dir <- "data/final/20240308/Followup/"
  chiara <- "data/people/chiara/"
} else {
  # Run on OS directly
  data.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20221004/"
  redcap.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20231030/"
  prefix <- "/Volumes/igmm/cvallejo-predicct/predicct/end-of-follow-up/"
  outdir <- "/Volumes/igmm/cvallejo-predicct/predicct/processed/"
  flare.dir <- paste0("/Volumes/igmm/cvallejo-predicct/predicct/",
                      "final/20240308/Followup/")
  chiara <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
}

# read-in data from multiple sources
physicalactivity <- read.xlsx(paste0(
  data.path,
  "Baseline2022/physicalactivity.xlsx"
))

demographics <- read.xlsx(paste0(
  data.path,
  "Baseline2022/demographics2022.xlsx"
))

IBD <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx"))

ffq <- read.xlsx(
  paste0(
    prefix,
    "predicct ffq_nutrientfood groupDQI all foods_data (n1092)Nov2022.",
    "xlsx"
  )
)

all_flares <- read.xlsx(paste0(flare.dir, "all-flares.xlsx"))
demo <- readRDS(paste0(outdir, "demo.RDS"))
flares <- readRDS(paste0(outdir, "flares/cutoff-flares.RDS"))
smoking <- readRDS(paste0(chiara, "smoking.rds"))
demo_uc <- readRDS(paste0(outdir, "demo-uc.RDS"))
demo_cd <- readRDS(paste0(outdir, "demo-cd.RDS"))
flares_soft <- readRDS(paste0(chiara, "flares_soft.RDS"))
flares_hard <- readRDS(paste0(chiara, "flares_hard.RDS"))
IBD_C <- readRDS(paste0(chiara, "IBD_C.RDS"))
```

## Data Cleaning

```{r}
#| echo: false
#| message: false
# 50 missing ParticipantId (blank rows - exclude these, from 1960 to 1910)
# PE.missing <- physicalactivity %>%
#   filter(is.na(ParticipantId)) %>%
#   nrow()

exercise <- physicalactivity %>%
  filter(!is.na(ParticipantId))

# add age and sex data from demographics
exercise <- exercise %>%
  dplyr::inner_join(
    demographics %>% select(ParticipantNo, Sex, age, diagnosis),
    by = "ParticipantNo"
  )

# Recode diagnosis
exercise %<>%
  dplyr::mutate(
    diagnosis2 = dplyr::case_when(
      diagnosis == 1 ~ 1,
      diagnosis == 2 ~ 2,
      diagnosis == 3 ~ 2,
      diagnosis == 4 ~ 2
    )
  ) %>%
  dplyr::mutate(diagnosis2 = factor(
    diagnosis2,
    levels = c("1", "2"),
    labels = c("CD", "UC/IBDU")
  ))

# exclude <18 yo = 53 participants (1910 -> 1857)
exercise %<>% dplyr::filter(age > 17)

# make sex into factor
exercise %<>% dplyr::mutate(Sex = factor(
  Sex,
  levels = c(1, 2),
  labels = c("Male", "Female"))
)

exercise %<>% dplyr::mutate(AgeGroup = cut(
  age,
  breaks = c(18, 24, 34, 44, 54, 65, Inf),
  labels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+"),
  include.lowest = TRUE))

exercise <- exercise %>%
  dplyr::left_join(
    IBD %>% dplyr::select(ParticipantNo, FlaresInPastYear)
    )

# 7 patients removed with missing data on flares
exercise <- exercise %>%
  filter(!is.na(FlaresInPastYear))

# create flare groups
exercise <- exercise %>%
  mutate(flare_group = ifelse(FlaresInPastYear == 0,
    "No Flares",
    "1 or More Flares"
  ))

# Make flares into levels
exercise %<>% 
  dplyr::mutate(flare_group = factor(
    flare_group,
    levels = c(
      "No Flares",
      "1 or More Flares"
  )
))

# Add FC data 
exercise <- exercise %>%
  dplyr::left_join(
    demo %>% select(ParticipantNo, FC, cat, Smoke, IMD),
    by = 'ParticipantNo'
    ) 

# Tidy - find and exclude if NA in VigorousWork, ModerateWork, WalkBicycle,
# VigrorousSport, and ModerateSport - Nobody removed
exercise %<>%
  filter(!dplyr::if_all(
    .cols = c(
      VigorousWork,
      ModerateWork,
      WalkBicycle,
      VigorousSport,
      ModerateSport
    ),
    .fns = is.na
  ))

# look at rows where there are NA values in key questions, exclude those with
# inconsistent answers (no valid response) - manually done, 6 values excluded
# exercise %>%
#   filter(
#     dplyr::if_any(
#       .cols = 
#     c(
#       VigorousWork,
#       ModerateWork,
#       WalkBicycle,
#       VigorousSport,
#       ModerateSport
#     ),
#     .fns = is.na))

exercise <- exercise %>%
  filter(!(ParticipantId %in% c(1395, 4685, 4803, 1452, 4827, 9609)))

# For yes/no questions, remaining NA values can be replaced with 2
exercise %<>%
  tidyr::replace_na(
    list(
      VigorousWork = 2,
      ModerateWork = 2,
      WalkBicycle = 2,
      VigorousSport = 2,
      ModerateSport = 2
    )
  )

# Exclude 10 participants who report 16 or more hours in any one activity,
# as per WHO guidelines
exercise <- exercise %>%
  filter(
    is.na(VigorousWorkHours) | VigorousWorkHours < 16,
    is.na(ModerateWorkHours) | ModerateWorkHours < 16,
    is.na(WalkBicycleHours) | WalkBicycleHours < 16,
    is.na(VigorousSportHours) | VigorousSportHours < 16,
    is.na(ModerateSportHours) | ModerateSportHours < 16
  )


# check that there are no cases of reporting 'no' to the question whilst
# providing answers in the time boxes

# exercise %>%
#   filter(VigorousWork == 2 &
#            (VigorousWorkHours > 0 | VigorousWorkMinutes > 0)) %>%
#   nrow()
# 
# exercise %>%
#   filter(ModerateWork == 2 &
#            (ModerateWorkHours > 0 | ModerateWorkMinutes > 0)) %>%
#   nrow()
# 
# exercise %>%
#   filter(WalkBicycle == 2 &
#            (WalkBicycleHours > 0 | WalkBicycleMinutes > 0)) %>%
#   nrow()
# 
# exercise %>%
#   filter(VigorousSport == 2 &
#            (VigorousSportHours > 0 | VigorousSportMinutes > 0)) %>%
#   nrow()
# 
# exercise %>%
#   filter(ModerateSport == 2 &
#            (ModerateSportHours > 0 | ModerateSportMinutes > 0)) %>%
#   nrow()

# No cases, all previously resolved.

# Similarly, fix cases of reporting 1 to the questions but for zero/NA hours.
# ModerateWork
MW_wrong <- exercise %>%
  filter(ModerateWork == 1 & (ModerateWorkDays == 0 | is.na(ModerateWorkDays)))

# 2 can be corrected recoding 1->2 ParticipantId 4700, 14446
exercise <- exercise %>%
  mutate(ModerateWork = dplyr::case_when(
    ParticipantId %in% c(4700, 14446) ~ 2,
    TRUE ~ ModerateWork
  ))

# 8 excluded as report 0 days but have answers in other domains
# Excluding subjects who say they do moderate work but for 0 days
# (they do have an hour though)

exercise <- exercise %>%
  filter(!(ModerateWork == 1 &
    (ModerateWorkDays == 0 | is.na(ModerateWorkDays))
  ))

# WalkBicycle
WB_wrong <- exercise %>%
  filter(WalkBicycle == 1 & (WalkBicycleDays == 0 | is.na(WalkBicycleDays)))

# 2 participants say they do but for 0 days
exercise <- exercise %>%
  filter(!(WalkBicycle == 1 & (WalkBicycleDays == 0 | is.na(WalkBicycleDays))))



# VigorousSport
VS_wrong <- exercise %>% filter(VigorousSport == 1 &
                                  (VigorousSportDays == 0 |
                                     is.na(VigorousSportDays)))

# Correct 3 participants who responded 1 but said for 0 days/hours/mins
exercise <- exercise %>%
  mutate(VigorousSport = dplyr::case_when(ParticipantId %in% c(7012, 9555, 9559) ~ 2,
                                          TRUE ~ VigorousSport))

# exclude 2 participants who did not include number of days
exercise <- exercise %>% filter(!(VigorousSport == 1 &
    (VigorousSportDays == 0 |
      is.na(VigorousSportDays))))

# ModerateSport
MS_wrong <- exercise %>%
  filter(ModerateSport == 1 & (ModerateSportDays == 0 | is.na(ModerateSportDays)))

exercise <- exercise %>%
  mutate(ModerateSport = dplyr::case_when(
    ParticipantId %in%
      c(61, 186, 3550, 7024, 7048, 9735, 12079) ~
      2,
    TRUE ~ ModerateSport
  ))

# exclude 6 participants who did not include number of days
exercise <- exercise %>% filter(!(ModerateSport == 1 &
                                    (ModerateSportDays == 0 | is.na(ModerateSportDays))
                                  ))

# check if anyone recorded days >7 in any days column - no cases
# check if any minutes >60 - no cases
# exercise %>%
#   filter(VigorousWorkDays > 7 | VigorousWorkMinutes > 60) %>%
#   nrow()
# exercise %>%
#   filter(ModerateWorkDays > 7 | ModerateWorkMinutes > 60) %>%
#   nrow()
# exercise %>%
#   filter(WalkBicycleDays > 7 | WalkBicycleMinutes > 60) %>%
#   nrow()
# exercise %>%
#   filter(VigorousSportDays > 7 | VigorousSportMinutes > 60) %>%
#   nrow()
# exercise %>%
#   filter(ModerateSportDays > 7 | ModerateSportMinutes > 60) %>%
#   nrow()

# Replace NA with zero in days/hours/minutes columns
exercise <- exercise %>% tidyr::replace_na(
  list(
    VigorousWorkDays = 0,
    VigorousWorkHours = 0,
    VigorousWorkMinutes = 0,
    ModerateWorkDays = 0,
    ModerateWorkHours = 0,
    ModerateWorkMinutes = 0,
    WalkBicycleDays = 0,
    WalkBicycleHours = 0,
    WalkBicycleMinutes = 0,
    VigorousSportDays = 0,
    VigorousSportHours = 0,
    VigorousSportMinutes = 0,
    ModerateSportDays = 0,
    ModerateSportHours = 0,
    ModerateSportMinutes = 0
  )
)

# Calculate total mins of vigorous and moderate exercise per week
exercise <- exercise %>%
  mutate(
    VigorousWeek =
      VigorousWorkDays * (VigorousWorkMinutes + (VigorousWorkHours * 60)) +
      VigorousSportDays * (VigorousSportMinutes + (VigorousSportHours * 60)),
    ModerateWeek = 
      ModerateWorkDays * (ModerateWorkMinutes + (ModerateWorkHours * 60)) +
      ModerateSportDays * (ModerateSportMinutes + (ModerateSportHours * 60)) +
      WalkBicycleDays * (WalkBicycleMinutes + (WalkBicycleHours * 60))
  )

# total mins + METs
exercise <- exercise %>% 
  mutate(
    ExerciseMins = VigorousWeek + ModerateWeek,
    METs = (VigorousWeek * 8) + (ModerateWeek * 4)
  )

# Targets for exercise (WHO) are 75 minutes vigorous activity or 150 minutes
# moderate activity or 600 METs per week
exercise <- exercise %>%
  mutate(
    MinimumExercise = dplyr::case_when(
      VigorousWeek >= 75 | ModerateWeek >= 150 | METs >= 600 ~ "Yes",
      TRUE ~ "No"
    ),
    LackExercise = dplyr::case_when(
      VigorousWeek < 75 & ModerateWeek < 150 & METs < 600 ~ "Yes",
      TRUE ~ "No"
    )
  )

# MinimumExercise as a factor
exercise %<>%
  dplyr::mutate(MinimumExercise = factor(MinimumExercise,
                                         levels = c('Yes', 'No')))


# exclude if minute exercise/week >6720 - 10 cases
# exercise %>% dplyr::filter(ExerciseMins > 6720) %>% nrow()

exercise <- exercise %>% dplyr::filter(ExerciseMins < 6720)


# IBD control scores
exercise %<>%
  dplyr::left_join(
    IBD_C %>% dplyr::select(
      ParticipantNo,
      control_score,
      OverallControl,
      control_8,
      control_grouped,
      vas_control
    ),
    by = "ParticipantNo"
  )

#change control_grouped and vas_control from character into factors
exercise %<>% 
  dplyr::mutate(control_grouped = factor(control_grouped))

exercise %<>% 
  dplyr::mutate(vas_control = factor(vas_control))


# Age in decades
exercise %<>%
  dplyr::mutate(
    age_decade = age/10
  )


# FC maximum detectable value is 1250
exercise %<>%
  dplyr::mutate(
    FC = dplyr::case_when(
      FC > 1250 ~ 1250,
      .default = FC
    )
  )

# Log transform FC due to extreme positive skew
exercise %<>%
  dplyr::mutate(
    FC = log(FC)
  )

```

## Baseline Analysis

#### Full Cohort

```{r}

data_baseline <- exercise

# Independent variables
independent = c('diagnosis2',
                'age',
                'Sex',
                'flare_group',
                'FC',
                'Smoke',
                'OverallControl')

# Dependent variable is MinimumExercise

# Statistical tests for each independent variable with MinimumExercise
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline, 
  dependent = 'MinimumExercise', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'diagnosis2'
)

summon_baseline_plot_discrete(
  data = data_baseline,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'Sex'
)

summon_baseline_plot_discrete(
  data = data_baseline,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'flare_group'
)

summon_baseline_plot_discrete(
  data = data_baseline,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'Smoke'
)

summon_baseline_plot_continuous(
  data = data_baseline,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'FC'
)

summon_baseline_plot_continuous(
  data = data_baseline,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'OverallControl'
)
```

#### Ulcerative Colitis/IBD Unclassified

```{r}
# Select anxiety data and UC patients
data_baseline_uc <- exercise %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Independent variables
independent = c('age',
                'Sex',
                'flare_group',
                'FC',
                'Smoke',
                'OverallControl')

# Dependent variable is MinimumExercise

# Statistical tests for each independent variable with MinimumExercise
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline_uc, 
  dependent = 'MinimumExercise', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline_uc,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'Sex'
)

summon_baseline_plot_discrete(
  data = data_baseline_uc,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'flare_group'
)

summon_baseline_plot_discrete(
  data = data_baseline_uc,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'Smoke'
)

summon_baseline_plot_continuous(
  data = data_baseline_uc,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline_uc,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'FC'
)

summon_baseline_plot_continuous(
  data = data_baseline_uc,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'OverallControl'
)
```

#### Crohn's Disease

```{r}
# Select anxiety data and CD patients
data_baseline_cd <- exercise %>%
  dplyr::filter(diagnosis2 == 'CD')

# Dependent variable is MinimumExercise

# Statistical tests for each independent variable with MinimumExercise
# With p-values adjusted for multiple testing using holm correction.
statistical_tests <- summon_statistical_test(
  data = data_baseline_cd, 
  dependent = 'MinimumExercise', 
  independent = independent)

statistical_tests

# Plots
summon_baseline_plot_discrete(
  data = data_baseline_cd,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'Sex'
)

summon_baseline_plot_discrete(
  data = data_baseline_cd,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'flare_group'
)

summon_baseline_plot_discrete(
  data = data_baseline_cd,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'Smoke'
)

summon_baseline_plot_continuous(
  data = data_baseline_cd,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'age'
)

summon_baseline_plot_continuous(
  data = data_baseline_cd,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'FC'
)

summon_baseline_plot_continuous(
  data = data_baseline_cd,
  stat_tests = statistical_tests,
  dependent = 'MinimumExercise',
  independent = 'OverallControl'
)
```


## Survival Analysis

### Data Cleaning

```{r}
# Survival data
# Join HADS with flare data
# Patient reported
data_survival_soft <- exercise %>% dplyr::inner_join(
  flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
  by = 'ParticipantNo'
) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# UC
data_survival_soft_uc <- data_survival_soft %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# CD
data_survival_soft_cd <- data_survival_soft %>%
  dplyr::filter(diagnosis2 == 'CD')


# Hard flares
data_survival_hard <- exercise %>% dplyr::inner_join(
  flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
  by = 'ParticipantNo'
) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

# UC
data_survival_hard_uc <- data_survival_hard %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# CD
data_survival_hard_cd <- data_survival_hard %>%
  dplyr::filter(diagnosis2 == 'CD')

```

### Shapes of continuous variables

#### Full Cohort

##### Patient reported flare

```{r}

# Cox with continuous variables
# Age
cox_age_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_soft
)

# Spline for age
summon_lrt(cox_age_soft, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_soft
)

# Spline for FC
summon_lrt(cox_fc_soft, remove = 'FC', add = 'ns(FC, df = 2)')
# No

# OverallControl
cox_control_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OverallControl +
    frailty(SiteNo),
  data = data_survival_soft
)

# Spline for control
summon_lrt(cox_control_soft, remove = 'OverallControl', add = 'ns(OverallControl, df = 2)')
# Yes

# OverallControl
cox_control_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(OverallControl, df = 2) +
    frailty(SiteNo),
  data = data_survival_soft
)

# Spline for control
summon_lrt(cox_control_soft, remove = 'ns(OverallControl, df = 2)', add = 'ns(OverallControl, df = 3)')
# No, 2 is sufficient

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_soft, 
  model = cox_age_soft, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_soft, 
  model = cox_fc_soft, 
  variable = 'FC')

# Control
plot_continuous_hr(
  data = data_survival_soft, 
  model = cox_control_soft, 
  variable = 'OverallControl',
  splineterm = 'ns(OverallControl, df = 2)')
```

##### Objective flare

```{r}

# Age
cox_age_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_hard
)

# Spline for age
summon_lrt(cox_age_hard, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_hard
)

# Spline for FC
summon_lrt(cox_fc_hard, remove = 'FC', add = 'ns(FC, df = 2)')
# No

# OverallControl
cox_control_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OverallControl +
    frailty(SiteNo),
  data = data_survival_hard
)

# Spline for control
summon_lrt(cox_control_hard, remove = 'OverallControl', add = 'ns(OverallControl, df = 2)')
# No
```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_hard, 
  model = cox_age_hard, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_hard, 
  model = cox_fc_hard, 
  variable = 'FC')

# Control
plot_continuous_hr(
  data = data_survival_soft, 
  model = cox_control_hard, 
  variable = 'OverallControl')
```

#### Ulcerative Colitis

##### Patient reported flare

```{r}
# Age
cox_age_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_soft_uc
)

# Spline for age
summon_lrt(cox_age_soft_uc, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_soft_uc
)

# Spline for FC
summon_lrt(cox_fc_soft_uc, remove = 'FC', add = 'ns(FC, df = 2)')
# No

# OverallControl
cox_control_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OverallControl +
    frailty(SiteNo),
  data = data_survival_soft_uc
)

# Spline for control
summon_lrt(cox_control_soft_uc, remove = 'OverallControl', add = 'ns(OverallControl, df = 2)')
# Yes

# OverallControl
cox_control_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(OverallControl, df = 2) +
    frailty(SiteNo),
  data = data_survival_soft_uc
)

summon_lrt(cox_control_soft_uc, remove = 'ns(OverallControl, df = 2)', add = 'ns(OverallControl, df = 3)')
# No, df = 2 is sufficient

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_soft_uc, 
  model = cox_age_soft_uc, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_soft_uc, 
  model = cox_fc_soft_uc, 
  variable = 'FC')

# Control
plot_continuous_hr(
  data = data_survival_soft_uc, 
  model = cox_control_soft_uc, 
  variable = 'OverallControl',
  splineterm = 'ns(OverallControl, df = 2)')
```

##### Objective flare

```{r}

# Age
cox_age_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

# Spline for age
summon_lrt(cox_age_hard_uc, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# Yes

# Refit
cox_age_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(age_decade, df = 2) +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

# Test for df = 3
summon_lrt(cox_age_hard_uc, remove = 'ns(age_decade, df = 2)', add = 'ns(age_decade, df = 3)')
# No - df = 2 is sufficient



# FC
cox_fc_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

# Spline for FC
summon_lrt(cox_fc_hard_uc, remove = 'FC', add = 'ns(FC, df = 2)')
# Yes

# Refit
cox_fc_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(FC, df = 2) +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

# Test for df = 3
summon_lrt(cox_fc_hard_uc, remove = 'ns(FC, df = 2)', add = 'ns(FC, df = 3)')
# No, stay with df = 2


# OverallControl
cox_control_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OverallControl +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

# Spline for control
summon_lrt(cox_control_hard_uc, remove = 'OverallControl', add = 'ns(OverallControl, df = 2)')
# No

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_hard_uc, 
  model = cox_age_hard_uc, 
  variable = 'age_decade',
  splineterm = 'ns(age_decade, df = 4)')

# FC
plot_continuous_hr(
  data = data_survival_hard_uc, 
  model = cox_fc_hard_uc, 
  variable = 'FC',
  splineterm = 'ns(FC, df = 2)')

# Control
plot_continuous_hr(
  data = data_survival_hard_uc, 
  model = cox_control_hard_uc, 
  variable = 'OverallControl')
```

#### Crohn's Disease

##### Patient reported flare

```{r}

# Age
cox_age_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_soft_cd
)

# Spline for age
summon_lrt(cox_age_soft_cd, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_soft_cd
)

# Spline for FC
summon_lrt(cox_fc_soft_cd, remove = 'FC', add = 'ns(FC, df = 2)')
# No

# OverallControl
cox_control_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OverallControl +
    frailty(SiteNo),
  data = data_survival_soft_cd
)

# Spline for control
summon_lrt(cox_control_soft_cd, remove = 'OverallControl', add = 'ns(OverallControl, df = 2)')
# Yes

# OverallControl
cox_control_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    ns(OverallControl, df = 2) +
    frailty(SiteNo),
  data = data_survival_soft_cd
)

summon_lrt(cox_control_soft_cd, remove = 'ns(OverallControl, df = 2)', add = 'ns(OverallControl, df = 3)')
# No, df = 2 is sufficient

```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_soft_cd, 
  model = cox_age_soft_cd, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_soft_cd, 
  model = cox_fc_soft_cd, 
  variable = 'FC')

# Control
plot_continuous_hr(
  data = data_survival_soft_cd, 
  model = cox_control_soft_cd, 
  variable = 'OverallControl',
  splineterm = 'ns(OverallControl, df = 2)')
```

##### Objective flare

```{r}

# Age
cox_age_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    age_decade +
    frailty(SiteNo),
  data = data_survival_hard_cd
)

# Spline for age
summon_lrt(cox_age_hard_cd, remove = 'age_decade', add = 'ns(age_decade, df = 2)')
# No

# FC
cox_fc_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    FC +
    frailty(SiteNo),
  data = data_survival_hard_cd
)

# Spline for FC
summon_lrt(cox_fc_hard_cd, remove = 'FC', add = 'ns(FC, df = 2)')
# No

# OverallControl
cox_control_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OverallControl +
    frailty(SiteNo),
  data = data_survival_hard_cd
)

# Spline for control
summon_lrt(cox_control_hard_cd, remove = 'OverallControl', add = 'ns(OverallControl, df = 2)')
# No
```

```{r}
# Plot shapes
# Age
plot_continuous_hr(
  data = data_survival_hard_cd, 
  model = cox_age_hard_cd, 
  variable = 'age_decade')

# FC
plot_continuous_hr(
  data = data_survival_hard_cd, 
  model = cox_fc_hard_cd, 
  variable = 'FC')

# Control
plot_continuous_hr(
  data = data_survival_hard_cd, 
  model = cox_control_hard_cd, 
  variable = 'OverallControl')
```

Summary:

Full cohort, soft, control df = 2, hard, no.

UC, soft, control df = 2. Hard, age df = 2, FC df = 2.

CD, soft, control df = 2. Hard, no splines.


### Models

#### Full Cohort

##### Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MinimumExercise, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_soft, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'Minimum Exercise',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    ns(OverallControl, df = 2) +
    Smoke +
    frailty(SiteNo),
  data = data_survival_soft
)

cox_anxiety_soft %>%
  print_survival()

```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_soft_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_soft,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_soft <- with(
  cox_anxiety_soft_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      MinimumExercise +
      IMD +
      Sex +
      age_decade +
      FC +
      flare_group +
      ns(OverallControl, df = 2) +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_soft %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MinimumExercise, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_hard, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'Minimum Exercise',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_hard
)

cox_anxiety_hard %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_hard_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_hard,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_hard <- with(
  cox_anxiety_hard_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      MinimumExercise +
      IMD +
      Sex +
      age_decade +
      FC +
      flare_group +
      OverallControl +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_hard %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```

#### Ulcerative Colitis

##### Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MinimumExercise, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_soft_uc, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'Minimum Exercise',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    ns(OverallControl, df = 2) +
    Smoke +
    frailty(SiteNo),
  data = data_survival_soft_uc
)

cox_anxiety_soft_uc %>%
  print_survival()

```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_soft_uc_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_soft_uc,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_soft_uc <- with(
  cox_anxiety_soft_uc_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      MinimumExercise +
      IMD +
      Sex +
      age_decade +
      FC +
      flare_group +
      ns(OverallControl, df = 2) +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_soft_uc %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MinimumExercise, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_hard_uc, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'Minimum Exercise',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    ns(age_decade, df = 2) +
    ns(FC, df = 2) +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

cox_anxiety_hard_uc %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
# Note: mice does not support splines so I cannot impute using a spline.
cox_anxiety_hard_uc_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_hard_uc,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_hard_uc <- with(
  cox_anxiety_hard_uc_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      MinimumExercise +
      IMD +
      Sex +
      ns(age_decade, df = 2) +
      ns(FC, df = 2) +
      flare_group +
      OverallControl +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_hard_uc %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```

#### Crohn's Disease

##### Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MinimumExercise, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_soft_cd, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'Minimum Exercise',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    ns(OverallControl, df = 2) +
    Smoke +
    frailty(SiteNo),
  data = data_survival_soft_cd
)

cox_anxiety_soft_cd %>%
  print_survival()

```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_soft_cd_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_soft_cd,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_soft_cd <- with(
  cox_anxiety_soft_cd_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      MinimumExercise +
      IMD +
      Sex +
      age_decade +
      FC +
      flare_group +
      ns(OverallControl, df = 2) +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_soft_cd %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```

##### Objective flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MinimumExercise, data = .) %>%
  ggsurvplot(
    ., 
    data = data_survival_hard_cd, 
    conf.int = TRUE, 
    risk.table = TRUE,
    pval = TRUE,
    pval.method = TRUE,
    legend.title = 'Minimum Exercise',
    xlab = "Time from study recruitment (days)",
    ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_hard_cd
)

cox_anxiety_hard_cd %>%
  print_survival()
```

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_hard_cd_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    MinimumExercise +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    Smoke +
    frailty(SiteNo),
  data = data_survival_hard_cd,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_hard_cd <- with(
  cox_anxiety_hard_cd_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      MinimumExercise +
      IMD +
      Sex +
      age_decade +
      FC +
      flare_group +
      OverallControl +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool()

cox_anxiety_hard_cd %>% 
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)
```





## Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>

Licensed by <a href="https://creativecommons.org/licenses/by/4.0/">CC BY</a> unless otherwise stated.
