---
title: "HADS"
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-expand: true
---

# Hospital Anxiety and Depression Scale

## Setting up in R

R packages

```{r}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = TRUE)
options(dplyr.summarise.inform = FALSE)

library(plyr) # Used for mapping values
suppressPackageStartupMessages(library(tidyverse)) # ggplot2, dplyr, and magrittr
library(readxl) # Read in Excel files
library(datefixR) # Standardise dates  
library(openxlsx)
library(tidyr)
library(ggdist)
suppressPackageStartupMessages(library(rstatix))
suppressPackageStartupMessages(library(table1))
library(knitr)
library(pander)
suppressPackageStartupMessages(library(kableExtra))
# Generate flowchart of cohort derivation
library(DiagrammeR)
library(DiagrammeRsvg)

#cox-regression
library(survival)
suppressPackageStartupMessages(library(survminer))
library(finalfit)
```

Helper functions

```{r}
# Chi squared tests
summon_chisq_test <- function(data, dependent, independent) {
  
  # Chi square tests for the dependent variable vs each independent variable
  # With p-values adjusted for multiple testing using Holm method
  
  purrr::map2_df(
    .x = dependent,
    .y = independent,
    .f = function(.x, .y) {
      x <- data %>%
        dplyr::pull(.x)
      
      y <- data %>%
        dplyr::pull(.y)
      
      chisq_test(x = x, y = y) %>%
        dplyr::mutate(dependent = .x, independent = .y)
    }
  ) %>%
    dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
    dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
    dplyr::select(-p.signif) %>%
    dplyr::mutate(
      p.signif = dplyr::case_when(
        p.adjust < 0.0001 ~ "****",
        p.adjust < 0.001 ~ "***",
        p.adjust < 0.01  ~ "**",
        p.adjust < 0.05  ~ "*",
        .default = 'ns'
      )
    )
  
}

# Creating proportions for plotting baseline data
calc_proportion <- function(data, dependent, independent){
  
  data %>%
    dplyr::group_by(!!rlang::sym(independent), !!rlang::sym(dependent)) %>%
    dplyr::count() %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!rlang::sym(independent)) %>%
    dplyr::mutate(p = n/sum(n)) 
  
}

summon_baseline_plots <- function(data, dependent, independent) {
  
  # Chi squared tests
  chisq_results <- summon_chisq_test(
    data = data,
    dependent = dependent,
    independent = independent
  ) %>%
    dplyr::transmute(independent, p.adjust) %>%
    {setNames(as.list(.$p.adjust), .$independent)}

  
  # Creating plots for baseline variables with the dependent variable
  
  plot_list <- list()
  
  for (y in independent) {
    
    # Create a plot for each independent variable
    
    plot <- data %>%
      calc_proportion(., dependent = dependent, independent = y) %>%
      ggplot(aes(x = .data[[dependent]], y = p, fill = .data[[y]])) +
      geom_col(position = 'dodge') +
      annotate(
        "text",
        label = paste0("Adjusted p-value: ", chisq_results[[y]]),
        x = Inf,
        y = Inf,
        vjust = 2,
        hjust = 1.3
      )
    
    plot_list[[y]] <- plot
    
  }
  
  # Return the plots
  plot_list
  
}
```

Load data

```{r}
# paths to PREdiCCt data
if (file.exists("/.dockerenv")) {
  # If running in docker
  data.path <- "data/final/20221004/"
  redcap.path <- "data/final/20231030/"
  prefix <- "data/end-of-follow-up/"
  outdir <- "data/processed"
} else {
  # Run on OS directly
  data.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20221004/"
  redcap.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20231030/"
  prefix <- "/Volumes/igmm/cvallejo-predicct/predicct/end-of-follow-up/"
  outdir <- "/Volumes/igmm/cvallejo-predicct/predicct/processed/"
}

#read-in data
#hads is question 42 of the baseline questionnaire
hads <- read.xlsx(paste0(data.path, "Baseline2022/hads.xlsx"))
health <- read.xlsx(paste0(data.path, "Baseline2022/health.xlsx"))
demographics <- read.xlsx(paste0(data.path, "Baseline2022/demographics2022.xlsx"))
lifeevents <- read.xlsx(paste0(data.path, "Baseline2022/lifeevents.xlsx"))
IBD <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx"))
ffq <- read.xlsx(paste0(prefix, "predicct ffq_nutrientfood groupDQI all foods_data (n1092)Nov2022.xlsx"))
demo <- readRDS(paste0(outdir, "demo.RDS"))
flares <- readRDS(paste0(outdir, "flares/cutoff-flares.RDS"))
IBD_C <- readRDS("/Volumes/igmm/cvallejo-predicct/people/chiara/IBD_C.RDS")
IBD_FC <- readRDS("/Volumes/igmm/cvallejo-predicct/people/chiara/IBD_FC.RDS")
flares_hard <- readRDS("/Volumes/igmm/cvallejo-predicct/people/chiara/flares_hard.RDS")
flares_soft <- readRDS("/Volumes/igmm/cvallejo-predicct/people/chiara/flares_soft.RDS")
```

## Data Cleaning

```{r}
# Exclude cases with missing ParticipantId
hads <- hads %>%
  filter(!is.na(ParticipantId))

health <- health %>%
  filter(!is.na(ParticipantId))

demographics <- demographics %>% 
  filter(!is.na(ParticipantId))

IBD <- IBD %>% 
  filter(!is.na(ParticipantId))

# join demographics
hads_with_demographics <- hads %>%
  dplyr::left_join(demographics %>%
                     dplyr::select(ParticipantNo, Sex, age, diagnosis),
                   by = "ParticipantNo")

# Convert diagnosis column to factor with labels
hads_with_demographics %<>%
  dplyr::mutate(
    diagnosis2 = dplyr::case_when(
      diagnosis == 1 ~ 1,
      diagnosis == 2 ~ 2,
      diagnosis == 3 ~ 2,
      diagnosis == 4 ~ 2
    )
  ) %>%
  dplyr::mutate(
    diagnosis2 = factor(
      diagnosis2,
      levels = c("1", "2"), 
      labels = c("CD", "UC/IBDU")
    )
  )

hads_with_demographics %<>%
  dplyr::mutate(Sex = factor(
    Sex,
    levels = c(1, 2),
    labels = c("Male", "Female")
  )
)

#exclude participants <17, 53 participants excluded, tot 1847

n_underage <- hads_with_demographics %>%
  dplyr::filter(age <= 17) %>%
  nrow()

hads_with_demographics %<>%
  dplyr::filter(age > 17)


#delete missing data, 9 participants, total 1838

n_missing_hads <- hads_with_demographics %>%
  dplyr::filter(
    dplyr::if_any(.cols = 7:20, .fns = is.na)
  ) %>%
  nrow()

hads_with_demographics %<>%
  dplyr::filter(
    !dplyr::if_any(.cols = 7:20, .fns = is.na)
  )


# correcting HADS scores
# Original data is on a 1 to 4 scale, rescale to between 0 and 3.

# Checking that currently HADS scores are up to a value of 4
# hads_scores_4 <- hads_with_demographics %>%
#   dplyr::select(FeelTense:CanEnjoyBookTV) %>%
#   dplyr::summarise(
#     dplyr::across(.cols = everything(), .fns = max)
#   ) %>% 
#   tidyr::pivot_longer(
#     cols = everything(),
#     names_to = "Score",
#     values_to = "max"
#   )

# Recale
hads_with_demographics %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = FeelTense:CanEnjoyBookTV,
      .fns = ~. - 1)
  )

  
#correcting inversely coded HADS
hads_with_demographics %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = c("FeelTense", 
                 "FrightenedFeelingSomethingAwful", 
                 "WorryingThoughts", 
                 "FeelCheerful", 
                 "FeelSlowedDown",
                 "LostInterestAppearance",
                 "FeelRestless",
                 "SuddenFeelingsPanic"),
      .fns = ~ 3 - .
    )
  )

#anxiety HADS composite score
hads_with_demographics %<>% 
  dplyr::mutate(anxiety_hads = FeelTense + 
           FrightenedFeelingSomethingAwful + 
           WorryingThoughts +
           SitAtEase +
           FrightenedFeelingButterflies +
           FeelRestless +
           SuddenFeelingsPanic)

#depression HADS composite score
hads_with_demographics %<>% 
  dplyr::mutate(depression_hads = EnjoyThings +
           CanLaugh +
           FeelCheerful +
           FeelSlowedDown +
           LostInterestAppearance +
           LookForward +
           CanEnjoyBookTV)

#remove breakdown hads columns
hads_with_demographics %<>%
  dplyr::select(-c(FeelTense:CanEnjoyBookTV))


#create age groups
hads_with_demographics %<>%
  dplyr::mutate(AgeGroup = cut(
    age,
    breaks = c(18, 24, 34, 44, 54, 65, Inf),
    labels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+"),
    include.lowest = TRUE
  ))


#add FC data and exclude missing FC
hads_with_demographics <- hads_with_demographics %>% 
  dplyr::left_join(
    demo %>% dplyr::select(ParticipantNo, FC, cat, IMD, Smoke), 
    by = "ParticipantNo")

n_missing_fc <- hads_with_demographics %>% 
  filter(is.na(cat)) %>%
  nrow()
# 209 removed with no FC

hads_with_demographics <- hads_with_demographics %>% 
  filter(!is.na(cat))

#Join flare data
hads_with_demographics <- hads_with_demographics %>% 
  dplyr::left_join(
    IBD %>% select(ParticipantNo, FlaresInPastYear), 
    by = "ParticipantNo"
    )

#create flare groups
hads_with_demographics <- hads_with_demographics %>%
  dplyr::mutate(
    flare_group = ifelse(FlaresInPastYear == 0, "No flares", "One or more flares")
    )

#make flares into levels
hads_with_demographics %<>%
  dplyr::mutate(flare_group = factor(flare_group, levels = c("No flares","One or more flares")))

# Remove subjects with missing flare group data - 3 subjects
hads_with_demographics <- hads_with_demographics %>% 
  filter(!is.na(flare_group))

#create hads_for_analysis which has new variable 'hads_type' for anxiety and depression hads
hads_for_analysis <- hads_with_demographics %>%
  tidyr::pivot_longer(
    cols = c(anxiety_hads, depression_hads),
    names_to = "hads_type", values_to = "hads_score"
  ) 

#score groups
hads_for_analysis <- hads_for_analysis %>% 
  mutate(
    score_group = cut(
      hads_score, 
      breaks = c(0, 7, 10, 21), 
      labels = c("0-7", "8-10", "11-21"), 
      include.lowest = TRUE)
  )
```

## Baseline Analysis

### Anxiety

#### Full Cohort

```{r}
# Select anxiety data
data_baseline_anxiety <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads')

# Independent variables
independent = c('diagnosis2',
                'AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is score_group

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_anxiety, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_anxiety <- summon_baseline_plots(
  data = data_baseline_anxiety, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_anxiety$diagnosis2
baseline_plots_anxiety$AgeGroup
baseline_plots_anxiety$Sex
baseline_plots_anxiety$flare_group
baseline_plots_anxiety$cat

```

#### Ulcerative Colitis/IBD Unspecified

```{r}
# Select anxiety data and UC patients
data_baseline_anxiety_uc <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is score_group

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_anxiety_uc, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_anxiety_uc <- summon_baseline_plots(
  data = data_baseline_anxiety_uc, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_anxiety_uc$AgeGroup
baseline_plots_anxiety_uc$Sex
baseline_plots_anxiety_uc$flare_group
baseline_plots_anxiety_uc$cat
```

#### Crohn's Disease

```{r}
# Select anxiety data and CD patients
data_baseline_anxiety_cd <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  dplyr::filter(diagnosis2 == 'CD')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is score_group

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_anxiety_cd, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_anxiety_cd <- summon_baseline_plots(
  data = data_baseline_anxiety_cd, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_anxiety_cd$AgeGroup
baseline_plots_anxiety_cd$Sex
baseline_plots_anxiety_cd$flare_group
baseline_plots_anxiety_cd$cat
```

### Depression

#### Full Cohort

```{r}
# Select depression data
data_baseline_depression <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads')

# Independent variables
independent = c('diagnosis2',
                'AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is score_group

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_depression, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_depression <- summon_baseline_plots(
  data = data_baseline_depression, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_depression$diagnosis2
baseline_plots_depression$AgeGroup
baseline_plots_depression$Sex
baseline_plots_depression$flare_group
baseline_plots_depression$cat
```

#### Ulcerative Colitis/IBD Unspecified

```{r}
# Select depression data and UC patients
data_baseline_depression_uc <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is score_group

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_depression_uc, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_depression_uc <- summon_baseline_plots(
  data = data_baseline_depression_uc, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_depression_uc$AgeGroup
baseline_plots_depression_uc$Sex
baseline_plots_depression_uc$flare_group
baseline_plots_depression_uc$cat
```

#### Crohn's Disease

```{r}
# Select depression data and CD patients
data_baseline_depression_cd <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::filter(diagnosis2 == 'CD')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is score_group

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_depression_cd, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_depression_cd <- summon_baseline_plots(
  data = data_baseline_depression_cd, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_depression_cd$AgeGroup
baseline_plots_depression_cd$Sex
baseline_plots_depression_cd$flare_group
baseline_plots_depression_cd$cat
```

## Survival Analysis

### Anxiety

#### Full Cohort

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_anxiety_soft <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_anxiety_hard <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

```

Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_anxiety_soft, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_anxiety_soft
)

cox_anxiety_soft %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_anxiety_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_anxiety_hard, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_anxiety_hard
)

cox_anxiety_hard %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

#### Ulcerative Colitis

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_anxiety_soft_uc <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_anxiety_hard_uc <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
   dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_anxiety_soft_uc, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_uc
)

cox_anxiety_soft_uc %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_anxiety_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_anxiety_hard_uc, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

cox_anxiety_hard_uc %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

#### Crohn's Disease

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_anxiety_soft_cd <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_anxiety_hard_cd <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'anxiety_hads') %>%
   dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_anxiety_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_anxiety_soft_cd, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_anxiety_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_anxiety_soft_cd
)

cox_anxiety_soft_cd %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_anxiety_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_anxiety_hard_cd, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_anxiety_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_anxiety_hard_cd
)

cox_anxiety_hard_cd %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

### Depression

#### Full Cohort

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_depression_soft <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_depression_hard <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

```

Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_soft, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_soft
)

cox_depression_soft %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_depression_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_hard, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_hard
)

cox_depression_hard %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

#### Ulcerative Colitis

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_depression_soft_uc <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_depression_hard_uc <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
   dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_soft_uc, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_soft_uc
)

cox_depression_soft_uc %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_depression_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_hard_uc, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_hard_uc
)

cox_depression_hard_uc %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

#### Crohn's Disease

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_depression_soft_cd <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_depression_hard_cd <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
   dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_soft_cd, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_soft_cd
)

cox_depression_soft_cd %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_depression_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_hard_cd, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_hard_cd
)

cox_depression_hard_cd %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

## Additional Analysis

### IBD Control Scores

Associations between IBD control, and anxiety and depression.

```{r}
# Control score data
data_control <- hads_for_analysis %>%
  dplyr::left_join(
    IBD_C %>% dplyr::select(
      ParticipantNo,
      control_score,
      OverallControl,
      control_8,
      control_grouped,
      vas_control
    ),
    by = "ParticipantNo"
  )

#exclude missing control scores
data_control %<>% 
  dplyr::filter(!is.na(control_grouped))

#change control_grouped and vas_control from character into factors
data_control %<>% 
  dplyr::mutate(control_grouped = factor(control_grouped))

data_control %<>% 
  dplyr::mutate(vas_control = factor(vas_control))

# Anxiety
data_control_anxiety <- data_control %>%
   dplyr::filter(hads_type == "anxiety_hads")

# Depression
data_control_depression <- data_control %>%
   dplyr::filter(hads_type == "depression_hads")
```

#### Full Cohort

```{r}
# Chi squared tests
# Anxiety
# Independent variables
independent = c('control_grouped', 'vas_control')

# Multiple testing - we should technically take the other tests we performed during the baseline analysis into account, but relax.

summon_chisq_test(
  data = data_control_anxiety, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_control_anxiety <- summon_baseline_plots(
  data = data_control_anxiety, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_control_anxiety$control_grouped
baseline_plots_control_anxiety$vas_control

# Depression
summon_chisq_test(
  data = data_control_depression, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_control_depression <- summon_baseline_plots(
  data = data_control_depression, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_control_depression$control_grouped
baseline_plots_control_depression$vas_control
```

#### Ulcerative Colitis

```{r}
# Chi squared tests
# UC patients only
data_control_anxiety_uc <- data_control_anxiety %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

data_control_depression_uc <- data_control_depression %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')


# Anxiety
# Independent variables
independent = c('control_grouped', 'vas_control')

summon_chisq_test(
  data = data_control_anxiety_uc, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_control_anxiety_uc <- summon_baseline_plots(
  data = data_control_anxiety_uc, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_control_anxiety_uc$control_grouped
baseline_plots_control_anxiety_uc$vas_control

# Depression
summon_chisq_test(
  data = data_control_depression_uc, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_control_depression_uc <- summon_baseline_plots(
  data = data_control_depression_uc, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_control_depression_uc$control_grouped
baseline_plots_control_depression_uc$vas_control
```

#### Crohn's Disease

```{r}
# Chi squared tests
# UC patients only
data_control_anxiety_cd <- data_control_anxiety %>%
  dplyr::filter(diagnosis2 == 'CD')

data_control_depression_cd <- data_control_depression %>%
  dplyr::filter(diagnosis2 == 'CD')


# Anxiety
# Independent variables
independent = c('control_grouped', 'vas_control')

summon_chisq_test(
  data = data_control_anxiety_cd, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_control_anxiety_cd <- summon_baseline_plots(
  data = data_control_anxiety_cd, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_control_anxiety_cd$control_grouped
baseline_plots_control_anxiety_cd$vas_control

# Depression
summon_chisq_test(
  data = data_control_depression_cd, 
  dependent = 'score_group', 
  independent = independent)

# Plots
baseline_plots_control_depression_cd <- summon_baseline_plots(
  data = data_control_depression_cd, 
  dependent = 'score_group',
  independent = independent
  )

baseline_plots_control_depression_cd$control_grouped
baseline_plots_control_depression_cd$vas_control
```

### Depression Grouped

Repeating survival analysis with moderate (8-10) and severe (11-21) depression scores grouped into a single category (8-21).

#### Full Cohort

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_depression_soft <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  mutate(score_group = case_when(
    score_group %in% c("8-10", "11-21") ~ "8-21",
    TRUE ~ score_group
  )) %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_depression_hard <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  mutate(score_group = case_when(
    score_group %in% c("8-10", "11-21") ~ "8-21",
    TRUE ~ score_group
  )) %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

```

Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_soft, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_soft
)

cox_depression_soft %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_depression_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_hard, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_hard
)

cox_depression_hard %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

#### Ulcerative Colitis

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_depression_soft_uc <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_depression_hard_uc <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
   dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_soft_uc, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_soft_uc
)

cox_depression_soft_uc %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_depression_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_hard_uc, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_hard_uc
)

cox_depression_hard_uc %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

#### Crohn's Disease

```{r}
# Survival data
# Join HADS with flare data

# Patient reported flares
data_survival_depression_soft_cd <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard Flares
data_survival_depression_hard_cd <- hads_for_analysis %>%
  dplyr::filter(hads_type == 'depression_hads') %>%
   dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Patient reported flare
# Kaplan Meier
data_survival_depression_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_soft_cd, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_soft_cd
)

cox_depression_soft_cd %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```

Hard flare.

```{r}
# Hard flare
# Kaplan Meier
data_survival_depression_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ score_group, data = .) %>%
  ggsurvplot(., data = data_survival_depression_hard_cd, conf.int = TRUE, ggtheme = theme_minimal())

# Cox model
cox_depression_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_depression_hard_cd
)

cox_depression_hard_cd %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)
```
