---
title: "Quality of Life"
author:
  - name: "Chiara Cotronei"
    url: https://www.research.ed.ac.uk/en/persons/chiara-cotronei
    affiliations:
      - ref: Lothian
  - name: "Alex Rudge"
    affiliations:
      - ref: CGEM
  - name: "Nathan Constantine-Cooke"
    corresponding: true
    url: https://scholar.google.com/citations?user=2emHWR0AAAAJ&hl=en&oi=ao
    affiliations:
      - ref: CGEM
      - ref: HGU
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-expand: true
---

# Quality of Life

## Setting up in R

R Packages

```{r}
#| label: load-libraries
#| echo: false
#| message: false
#| warning: false
set.seed(123)
library(plyr) # Used for mapping values
suppressPackageStartupMessages(library(tidyverse))
library(readxl) # Read in Excel files
library(datefixR) # Standardise dates
library(openxlsx)
library(tidyr)
library(ggdist)
library(rstatix)

suppressPackageStartupMessages(library(table1))
library(knitr)
library(pander)
library(kableExtra)

# Generate flowchart of cohort derivation
library(DiagrammeR)
library(DiagrammeRsvg)

# cox-regression
library(survival)
library(survminer)
library(finalfit)
```

Helper functions

```{r}
# Chi squared tests
summon_chisq_test <- function(data, dependent, independent) {
  
  # Chi square tests/Fisher's exact test for the dependent variable vs each
  # independent variable
  
  purrr::map2_df(
    .x = dependent,
    .y = independent,
    .f = function(.x, .y) {
      x <- data %>%
        dplyr::pull(.x)
      
      y <- data %>%
        dplyr::pull(.y)
      
      # Check whether we need a Fisher test
      fisher_flag <- data %>%
        # Count combinations of x and y
        count(!!rlang::sym(.x), !!rlang::sym(.y)) %>%
        # Pull the count
        dplyr::pull(n) %>%
        # Check if any counts are < 5
        min() < 5
        
      if (fisher_flag){
        # If one of the counts is < 5 use Fisher exact test
        fisher_test(table(x, y), simulate.p.value = TRUE, B = 1e5) %>%
          dplyr::mutate(
            dependent = .x, 
            independent = .y,
            method = 'Fisher\'s exact test')
        
      } else {    
        # Otherwise use chi square
        chisq_test(x = x, y = y) %>%
          dplyr::mutate(dependent = .x, independent = .y)
      }
    }
  ) %>%
    dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
    dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
    dplyr::select(-p.signif) %>%
    dplyr::mutate(
      p.signif = dplyr::case_when(
        p.adjust < 0.0001 ~ "****",
        p.adjust < 0.001 ~ "***",
        p.adjust < 0.01  ~ "**",
        p.adjust < 0.05  ~ "*",
        .default = 'ns'
      )
    )
  
}


print_chisq <- function(data) {
  data %>%
    knitr::kable(digits = c(0, 2, 4, 0, NA, NA, NA, 4, NA),
                 col.names = c("n",
                               "Statistic",
                               "p",
                               "df",
                               "Method",
                               "Dependent",
                               "Independent",
                               "Adjusted p-value",
                               "Adjusted significance")
               )
}

print_survival <- function(data) {
  data %>%
    broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    knitr::kable(digits = 3,
                 col.names = c("Covariate",
                               "Estimate",
                               "Std. Error",
                               "Statistic",
                               "p value",
                               "Lower CI",
                               "Upper CI")
                 )
}


# Creating proportions for plotting baseline data
calc_proportion <- function(data, dependent, independent){
  
  data %>%
    dplyr::group_by(!!rlang::sym(independent), !!rlang::sym(dependent)) %>%
    dplyr::count() %>%
    dplyr::ungroup() %>%
    dplyr::group_by(!!rlang::sym(independent)) %>%
    dplyr::mutate(p = n/sum(n)) 
  
}

summon_baseline_plots <- function(data, dependent, independent) {
  
  # Chi squared tests
  chisq_results <- summon_chisq_test(
    data = data,
    dependent = dependent,
    independent = independent
  ) %>%
    dplyr::transmute(independent, p.adjust) %>%
    {setNames(as.list(.$p.adjust), .$independent)}

  
  # Creating plots for baseline variables with the dependent variable
  
  plot_list <- list()
  
  for (y in independent) {
    
    # Create a plot for each independent variable
    
    plot <- data %>%
      calc_proportion(., dependent = dependent, independent = y) %>%
      ggplot(aes(x = .data[[dependent]], y = p, fill = .data[[y]])) +
      geom_col(position = 'dodge') +
      annotate(
        "text",
        label = paste0("Adjusted p-value: ", chisq_results[[y]]),
        x = Inf,
        y = Inf,
        vjust = 2,
        hjust = 1.3
      ) +
      theme_minimal()
    
    plot_list[[y]] <- plot
    
  }
  # Return the plots
  plot_list
}
```

Load data

```{r}
# paths to PREdiCCt data
if (file.exists("/.dockerenv")) {
  # If running in docker
  data.path <- "data/final/20221004/"
  redcap.path <- "data/final/20231030/"
  prefix <- "data/end-of-follow-up/"
  outdir <- "data/processed"
  flare.dir <- "data/final/20240308/Followup/"
  chiara <- "data/people/chiara/"
} else {
  # Run on OS directly
  data.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20221004/"
  redcap.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20231030/"
  prefix <- "/Volumes/igmm/cvallejo-predicct/predicct/end-of-follow-up/"
  outdir <- "/Volumes/igmm/cvallejo-predicct/predicct/processed/"
  flare.dir <- paste0("/Volumes/igmm/cvallejo-predicct/predicct/",
                      "final/20240308/Followup/")
  save_path <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
}

lifeevents <- read.xlsx(paste0(data.path, "Baseline2022/lifeevents.xlsx"))
demographics <- read.xlsx(paste0(data.path, "Baseline2022/demographics2022.xlsx"))
demo <- readRDS(paste0(outdir, "demo.RDS"))
IBD <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx"))
QOL <- read.xlsx(paste0(data.path, "Baseline2022/health.xlsx"))
flares_hard <- readRDS(paste0(save_path, "flares_hard.RDS"))
flares_soft <- readRDS(paste0(save_path, "flares_soft.RDS"))
```

## Data Cleaning

```{r}

QOL <- QOL %>% 
  filter(!is.na(ParticipantNo))

#add age and sex data from demographics
QOL <- QOL %>% 
  left_join(
    demographics %>% select(ParticipantNo, Sex, age, diagnosis), 
    by  = "ParticipantNo")

# Recode diagnosis
QOL %<>%
  dplyr::mutate(
    diagnosis2 = dplyr::case_when(
      diagnosis == 1 ~ 1,
      diagnosis == 2 ~ 2,
      diagnosis == 3 ~ 2,
      diagnosis == 4 ~ 2
    )
  ) %>%
  dplyr::mutate(diagnosis2 = factor(
    diagnosis2,
    levels = c("1", "2"),
    labels = c("CD", "UC/IBDU")
  ))

#exclude <18 yo = 53 participants (1907 -> 1854)
QOL %<>% dplyr::filter(age >= 18)

#make sex into level
QOL %<>% dplyr::mutate(Sex = factor(
  Sex,
  levels = c(1, 2),
  labels = c("Male", "Female"))
)

#make age brackets
QOL %<>%
  dplyr::mutate(AgeGroup = cut(age,
                               breaks = c(18, 24, 34, 44, 54, 65, Inf), 
                               labels = c("18-24",
                                          "25-34",
                                          "35-44",
                                          "45-54",
                                          "55-64",
                                          "65+"),
                               include.lowest = TRUE))

QOL <- QOL %>% 
  left_join(IBD %>% select(ParticipantNo, FlaresInPastYear),
            by = "ParticipantNo")

# 6 missing flare data
#QOL %>% filter(is.na(FlaresInPastYear)) %>% nrow()

QOL <- QOL %>%
  filter(!is.na(FlaresInPastYear))

#create flare groups
QOL <- QOL %>%
  mutate(flare_group = ifelse(FlaresInPastYear == 0,
                              "No Flares", "1 or More Flares"))

#make flares into levels
QOL %<>% 
  dplyr::mutate(
    flare_group = factor(flare_group,
                         levels = c( "No Flares","1 or More Flares"))
  )

QOL <- QOL %>% 
  left_join(demo %>% select(ParticipantNo, FC, cat, Biologic, IMD),
            by = "ParticipantNo")

#209 with missing FC
#QOL %>% filter(is.na(cat)) %>% nrow()

QOL <- QOL %>% 
  filter(!is.na(cat))

# Check missing data - 53 
# QOL %>%
#   filter(
#     dplyr::if_any(.cols = c(7:18), .fns = is.na)
#   ) %>% nrow()

# Exclude missing
QOL %<>%
  filter(
    !dplyr::if_any(.cols = c(7:18), .fns = is.na)
  )


# Recoding answers to be out of 6, leaving option 4 empty.
QOL <- QOL %>%
  dplyr::mutate(
    dplyr::across(
      .cols = c(CalmAndPeaceful,
                LotOfEnergy,
                DownheartedBlue,
                SocialActivities),
      .fn = function(x) dplyr::case_when(
        x == 5 ~ 6,
        x == 4 ~ 5,
        .default = x
      )
    )
  )

#Reverse code GeneralHealth, PainInterfere, CalmAndPeaceful, DownheartedBlue
QOL %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = c(CalmAndPeaceful, DownheartedBlue),
      .fn = ~ 7 - .
    )
  )

QOL %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = c(GeneralHealth, PainInterfere),
      .fn = ~ 6 - .
    )
  )

QOL <- QOL %>% 
  rename(gh1 = GeneralHealth, 
         pf02 = ModerateActivities, 
         pf04 = ClimbingStairs, 
         rp2 = PhysicalAccomplishedLess, 
         rp3 = PhysicalLimitedKindWork, 
         re2 = EmotionalAccomplishedLess, 
         re3 = EmotionalNotCarefulAsUsual, 
         bp2 = PainInterfere, 
         mh3 = CalmAndPeaceful, 
         vt2 = LotOfEnergy, 
         mh4 = DownheartedBlue, 
         sf2 = SocialActivities)


# Code from https://github.com/lbraglia/lbscorer/blob/master/R/sf12.R 
# Comments reworded for clarity
  
# Dummy variables for pf02 (3 as reference)
QOL$pf02_1 <- as.numeric(QOL$pf02 == 1L)
QOL$pf02_2 <- as.numeric(QOL$pf02 == 2L)

# Dummy variables for pf02 (3 as reference)
QOL$pf04_1 <- as.numeric(QOL$pf04 == 1L)
QOL$pf04_2 <- as.numeric(QOL$pf04 == 2L)

# Dummy variables for rp2 (2 as reference)
QOL$rp2_1 <- as.numeric(QOL$rp2 == 1L)

# Dummy variables for rp2 (2 as reference)
QOL$rp3_1 <- as.numeric(QOL$rp3 == 1L)

# Dummy variables for bp2 (5 as reference)
QOL$bp2_1 <- as.numeric(QOL$bp2 == 1L)
QOL$bp2_2 <- as.numeric(QOL$bp2 == 2L)
QOL$bp2_3 <- as.numeric(QOL$bp2 == 3L)
QOL$bp2_4 <- as.numeric(QOL$bp2 == 4L)

# Dummy variables for gh1 (5 as reference)
QOL$gh1_1 <- as.numeric(QOL$gh1 == 1L)
QOL$gh1_2 <- as.numeric(QOL$gh1 == 2L)
QOL$gh1_3 <- as.numeric(QOL$gh1 == 3L)
QOL$gh1_4 <- as.numeric(QOL$gh1 == 4L)

# Dummy variables for vt2 (6 as reference)
QOL$vt2_1 <- as.numeric(QOL$vt2 == 1L)
QOL$vt2_2 <- as.numeric(QOL$vt2 == 2L)
QOL$vt2_3 <- as.numeric(QOL$vt2 == 3L)
QOL$vt2_4 <- as.numeric(QOL$vt2 == 4L)
QOL$vt2_5 <- as.numeric(QOL$vt2 == 5L)

# Dummy variables for sf2 (5 as reference)
QOL$sf2_1 <- as.numeric(QOL$sf2 == 1L)
QOL$sf2_2 <- as.numeric(QOL$sf2 == 2L)
QOL$sf2_3 <- as.numeric(QOL$sf2 == 3L)
QOL$sf2_4 <- as.numeric(QOL$sf2 == 4L)

# Dummy variables for re2 (2 as reference)
QOL$re2_1 <- as.numeric(QOL$re2 == 1L)

# Dummy variables for re3 (2 as reference)
QOL$re3_1 <- as.numeric(QOL$re3 == 1L)

# Dummy variables for mh3 (6 as reference)
QOL$mh3_1 <- as.numeric(QOL$mh3 == 1L)
QOL$mh3_2 <- as.numeric(QOL$mh3 == 2L)
QOL$mh3_3 <- as.numeric(QOL$mh3 == 3L)
QOL$mh3_4 <- as.numeric(QOL$mh3 == 4L)
QOL$mh3_5 <- as.numeric(QOL$mh3 == 5L)

# Dummy variables for mh4 (6 as reference)
QOL$mh4_1 <- as.numeric(QOL$mh4 == 1L)
QOL$mh4_2 <- as.numeric(QOL$mh4 == 2L)
QOL$mh4_3 <- as.numeric(QOL$mh4 == 3L)
QOL$mh4_4 <- as.numeric(QOL$mh4 == 4L)
QOL$mh4_5 <- as.numeric(QOL$mh4 == 5L)


# Calculate scores using regression weights from the GitHub

QOL <- QOL %>% 
  mutate(RAWPCS12 = (-7.23216*pf02_1) + (-3.45555*pf02_2) +
                    (-6.24397*pf04_1) + (-2.73557*pf04_2) +
                    (-4.61617*rp2_1) + 
                    (-5.51747*rp3_1) +
                    (-11.25544*bp2_1) + (-8.38063*bp2_2) +
                    (-6.50522*bp2_3) + (-3.80130*bp2_4) + (-8.37399*gh1_1) +
                    (-5.56461*gh1_2) + (-3.02396*gh1_3) + (-1.31872*gh1_4) +
                    (-2.44706*vt2_1) + (-2.02168*vt2_2) + (-1.6185*vt2_3) +
                    (-1.14387*vt2_4) + (-0.42251*vt2_5) + (-0.33682*sf2_1) +
                    (-0.94342*sf2_2) + (-0.18043*sf2_3) + (0.11038*sf2_4) +
                    (3.04365*re2_1) + (2.32091*re3_1) + (3.46638*mh3_1) +
                    (2.90426*mh3_2) + (2.37241*mh3_3) + (1.36689*mh3_4) +
                    (0.66514*mh3_5) + (4.61446*mh4_1) + (3.41593*mh4_2) +
                    (2.34247*mh4_3) + (1.28044*mh4_4) + (0.41188*mh4_5))

QOL <- QOL %>% 
  mutate(RAWMCS12 = (3.93115*pf02_1) + (1.8684*pf02_2) +
                   (2.68282*pf04_1) + (1.43103*pf04_2) + (1.4406*rp2_1) +
                   (1.66968*rp3_1) + (1.48619*bp2_1) + (1.76691*bp2_2) +
                   (1.49384*bp2_3) + (0.90384*bp2_4) + (-1.71175*gh1_1) +
                   (-0.16891*gh1_2) + (0.03482*gh1_3) + (-0.06064*gh1_4) +
                   (-6.02409*vt2_1) + (-4.88962*vt2_2) + (-3.29805*vt2_3) +
                   (-1.65178*vt2_4) + (-0.92057*vt2_5) + (-6.29724*sf2_1) +
                   (-8.26066*sf2_2) + (-5.63286*sf2_3) + (-3.13896*sf2_4) +
                   (-6.82672*re2_1) + (-5.69921*re3_1) + (-10.19085*mh3_1) +
                   (-7.92717*mh3_2) + (-6.31121*mh3_3) + (-4.09842*mh3_4) +
                   (-1.94949*mh3_5) + (-16.15395*mh4_1) + (-10.77911*mh4_2) +
                   (-8.09914*mh4_3) + (-4.59055*mh4_4) + (-1.95934*mh4_5))

# Standardisation of scores - from GitHub
# Physical component score
QOL <- QOL %>% 
  mutate(PSC12 = RAWPCS12 + 56.57706)

# Mental component score
QOL <- QOL %>% 
  mutate(MCS12 = RAWMCS12 + 60.75781)

# Binary versions of the scores
QOL %<>% 
  mutate(PSC12_binary = ifelse(PSC12 >= 50, "50+", "<50"))

QOL %<>% 
  mutate(MCS12_binary = ifelse(MCS12 >= 50, "50+", "<50"))
```

```{r}
# Our dependent variables are 2 continuous variables representing patients' 
# Physical component and Mental component scores

# Data is called QOL, rename to data_baseline
data_baseline <- QOL

# Variables PSC12 (Physical) and MCS12 (Mental)

# Density plots of each score
data_baseline %>%
  ggplot(aes(x = PSC12)) +
  geom_density() +
  theme_minimal()

data_baseline %>%
  ggplot(aes(x = MCS12)) +
  geom_density() +
  theme_minimal()
```

## Baseline Analysis

### Physical Component (numeric)

#### Full Cohort

```{r}

data_baseline <- QOL

# Hypothesis test
# 2 groups use wilcox (aka Mann Whitney), for 3 or more use kruskal.
# Could also use kruskal for all as it gives the same p value.

dplyr::bind_rows(
  data_baseline %>%
    wilcox_test(PSC12 ~ diagnosis2) %>%
    dplyr::mutate(group = "diagnosis2"),
  
  data_baseline %>%
    wilcox_test(PSC12 ~ Sex) %>%
    dplyr::mutate(group = "Sex"),
  
  data_baseline %>%
    kruskal_test(PSC12 ~ AgeGroup) %>%
    dplyr::mutate(group = "AgeGroup"),
  
  data_baseline %>%
    wilcox_test(PSC12 ~ flare_group) %>%
    dplyr::mutate(group = "flare_group"),
  
  data_baseline %>%
    kruskal_test(PSC12 ~ cat) %>%
    dplyr::mutate(group = "cat"),
) %>%
  tidyr::replace_na(list(method = "Mann-Whitney")) %>%
  dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
  dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
  dplyr::mutate(
    p.signif = dplyr::case_when(
      p.adjust < 0.0001 ~ "****",
      p.adjust < 0.001 ~ "***",
      p.adjust < 0.01  ~ "**",
      p.adjust < 0.05  ~ "*",
      .default = 'ns'
    )
  )  %>%
  knitr::kable()

# Baseline plots
# Diagnosis
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = diagnosis2)
  ) + theme_minimal()

# Sex
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = Sex)
  ) + theme_minimal()

# Age Groups
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = AgeGroup)
  ) + theme_minimal()

# Flare
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = flare_group)
  ) + theme_minimal()

# FC
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = cat)
  ) + theme_minimal()
```

#### Ulcerative Colitis

```{r}
# UC patients
data_baseline_uc <- QOL %>% 
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Hypothesis test
# 2 groups use wilcox (aka Mann Whitney), for 3 or more use kruskal.
# Could also use kruskal for all as it gives the same p value.

dplyr::bind_rows(
  
  data_baseline_uc %>%
    wilcox_test(PSC12 ~ Sex) %>%
    dplyr::mutate(group = "Sex"),
  
  data_baseline_uc %>%
    kruskal_test(PSC12 ~ AgeGroup) %>%
    dplyr::mutate(group = "AgeGroup"),
  
  data_baseline_uc %>%
    wilcox_test(PSC12 ~ flare_group) %>%
    dplyr::mutate(group = "flare_group"),
  
  data_baseline_uc %>%
    kruskal_test(PSC12 ~ cat) %>%
    dplyr::mutate(group = "cat"),
) %>%
  tidyr::replace_na(list(method = "Mann-Whitney")) %>%
  dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
  dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
  dplyr::mutate(
    p.signif = dplyr::case_when(
      p.adjust < 0.0001 ~ "****",
      p.adjust < 0.001 ~ "***",
      p.adjust < 0.01  ~ "**",
      p.adjust < 0.05  ~ "*",
      .default = 'ns'
    )
  ) %>%
  knitr::kable()

# Baseline plots

# Sex
data_baseline_uc %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = Sex)
  ) + theme_minimal()

# Age Groups
data_baseline_uc %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = AgeGroup)
  ) + theme_minimal()

# Flare
data_baseline_uc %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = flare_group)
  ) + theme_minimal()

# FC
data_baseline_uc %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = cat)
  ) + theme_minimal()
```

#### Crohn's Disease

```{r}
# CD patients
data_baseline_cd <- QOL %>% 
  dplyr::filter(diagnosis2 == 'CD')

# Hypothesis test
# 2 groups use wilcox (aka Mann Whitney), for 3 or more use kruskal.
# Could also use kruskal for all as it gives the same p value.

dplyr::bind_rows(
  
  data_baseline_cd %>%
    wilcox_test(PSC12 ~ Sex) %>%
    dplyr::mutate(group = "Sex"),
  
  data_baseline_cd %>%
    kruskal_test(PSC12 ~ AgeGroup) %>%
    dplyr::mutate(group = "AgeGroup"),
  
  data_baseline_cd %>%
    wilcox_test(PSC12 ~ flare_group) %>%
    dplyr::mutate(group = "flare_group"),
  
  data_baseline_cd %>%
    kruskal_test(PSC12 ~ cat) %>%
    dplyr::mutate(group = "cat"),
) %>%
  tidyr::replace_na(list(method = "Mann-Whitney")) %>%
  dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
  dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
  dplyr::mutate(
    p.signif = dplyr::case_when(
      p.adjust < 0.0001 ~ "****",
      p.adjust < 0.001 ~ "***",
      p.adjust < 0.01  ~ "**",
      p.adjust < 0.05  ~ "*",
      .default = 'ns'
    )
  ) %>%
  knitr::kable()

# Baseline plots

# Sex
data_baseline_cd %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = Sex)
  ) + theme_minimal()

# Age Groups
data_baseline_cd %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = AgeGroup)
  ) + theme_minimal()

# Flare
data_baseline_cd %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = flare_group)
  ) + theme_minimal()

# FC
data_baseline_cd %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = cat)
  ) + theme_minimal()

```

### Mental Component (numeric)

#### Full Cohort

```{r}

data_baseline <- QOL

# Hypothesis test
# 2 groups use wilcox (aka Mann Whitney), for 3 or more use kruskal.
# Could also use kruskal for all as it gives the same p value.

dplyr::bind_rows(
  data_baseline %>%
    wilcox_test(MCS12 ~ diagnosis2) %>%
    dplyr::mutate(group = "diagnosis2"),
  
  data_baseline %>%
    wilcox_test(MCS12 ~ Sex) %>%
    dplyr::mutate(group = "Sex"),
  
  data_baseline %>%
    kruskal_test(MCS12 ~ AgeGroup) %>%
    dplyr::mutate(group = "AgeGroup"),
  
  data_baseline %>%
    wilcox_test(MCS12 ~ flare_group) %>%
    dplyr::mutate(group = "flare_group"),
  
  data_baseline %>%
    kruskal_test(MCS12 ~ cat) %>%
    dplyr::mutate(group = "cat"),
) %>%
  tidyr::replace_na(list(method = "Mann-Whitney")) %>%
  dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
  dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
  dplyr::mutate(
    p.signif = dplyr::case_when(
      p.adjust < 0.0001 ~ "****",
      p.adjust < 0.001 ~ "***",
      p.adjust < 0.01  ~ "**",
      p.adjust < 0.05  ~ "*",
      .default = 'ns'
    )
  ) %>%
  knitr::kable()

# Baseline plots
# Diagnosis
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = diagnosis2)
  ) +
  theme_minimal()

# Sex
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = Sex)
  ) +
  theme_minimal()

# Age Groups
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = AgeGroup)
  ) +
  theme_minimal()

# Flare
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = flare_group)
  ) +
  theme_minimal()

# FC
data_baseline %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = cat)
  ) +
  theme_minimal()
```

#### Ulcerative Colitis

```{r}
# UC patients
data_baseline_uc <- QOL %>% 
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Hypothesis test
# 2 groups use wilcox (aka Mann Whitney), for 3 or more use kruskal.
# Could also use kruskal for all as it gives the same p value.

dplyr::bind_rows(
  
  data_baseline_uc %>%
    wilcox_test(MCS12 ~ Sex) %>%
    dplyr::mutate(group = "Sex"),
  
  data_baseline_uc %>%
    kruskal_test(MCS12 ~ AgeGroup) %>%
    dplyr::mutate(group = "AgeGroup"),
  
  data_baseline_uc %>%
    wilcox_test(MCS12 ~ flare_group) %>%
    dplyr::mutate(group = "flare_group"),
  
  data_baseline_uc %>%
    kruskal_test(MCS12 ~ cat) %>%
    dplyr::mutate(group = "cat"),
) %>%
  tidyr::replace_na(list(method = "Mann-Whitney")) %>%
  dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
  dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
  dplyr::mutate(
    p.signif = dplyr::case_when(
      p.adjust < 0.0001 ~ "****",
      p.adjust < 0.001 ~ "***",
      p.adjust < 0.01  ~ "**",
      p.adjust < 0.05  ~ "*",
      .default = 'ns'
    )
  ) %>%
  knitr::kable()

# Baseline plots

# Sex
data_baseline_uc %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = Sex)
  ) + theme_minimal()

# Age Groups
data_baseline_uc %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = AgeGroup)
  ) + theme_minimal()

# Flare
data_baseline_uc %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = flare_group)
  ) + theme_minimal()

# FC
data_baseline_uc %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = cat)
  ) + theme_minimal()
```

#### Crohn's Disease

```{r}
# CD patients
data_baseline_cd <- QOL %>% 
  dplyr::filter(diagnosis2 == 'CD')

# Hypothesis test
# 2 groups use wilcox (aka Mann Whitney), for 3 or more use kruskal.
# Could also use kruskal for all as it gives the same p value.

dplyr::bind_rows(
  
  data_baseline_cd %>%
    wilcox_test(MCS12 ~ Sex) %>%
    dplyr::mutate(group = "Sex"),
  
  data_baseline_cd %>%
    kruskal_test(MCS12 ~ AgeGroup) %>%
    dplyr::mutate(group = "AgeGroup"),
  
  data_baseline_cd %>%
    wilcox_test(MCS12 ~ flare_group) %>%
    dplyr::mutate(group = "flare_group"),
  
  data_baseline_cd %>%
    kruskal_test(MCS12 ~ cat) %>%
    dplyr::mutate(group = "cat"),
) %>%
  tidyr::replace_na(list(method = "Mann-Whitney")) %>%
  dplyr::mutate(p.adjust = p.adjust(p = p, method = 'holm')) %>%
  dplyr::mutate(p.adjust = signif(p.adjust, 3)) %>%
  dplyr::mutate(
    p.signif = dplyr::case_when(
      p.adjust < 0.0001 ~ "****",
      p.adjust < 0.001 ~ "***",
      p.adjust < 0.01  ~ "**",
      p.adjust < 0.05  ~ "*",
      .default = 'ns'
    )
  ) %>%
  knitr::kable()

# Baseline plots

# Sex
data_baseline_cd %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = Sex)
  ) + theme_minimal()

# Age Groups
data_baseline_cd %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = AgeGroup)
  ) + theme_minimal()

# Flare
data_baseline_cd %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = flare_group)
  ) + theme_minimal()

# FC
data_baseline_cd %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = cat)
  ) + theme_minimal()
```

### Physical Component (categorical)

#### Full Cohort

```{r}
# Rename QOL
data_baseline <- QOL

# Independent variables
independent = c('diagnosis2',
                'AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is PSC12_binary

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline, 
  dependent = 'PSC12_binary', 
  independent = independent) %>%
  print_chisq()

# Plots
baseline_plots_physical <- summon_baseline_plots(
  data = data_baseline, 
  dependent = 'PSC12_binary',
  independent = independent
  )

baseline_plots_physical$diagnosis2
baseline_plots_physical$AgeGroup
baseline_plots_physical$Sex
baseline_plots_physical$flare_group
baseline_plots_physical$cat
```

#### Ulcerative Colitis

```{r}
# UC patients only
data_baseline_uc <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is PSC12_binary

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_uc, 
  dependent = 'PSC12_binary', 
  independent = independent) %>%
  print_chisq()

# Plots
baseline_plots_physical_uc <- summon_baseline_plots(
  data = data_baseline_uc, 
  dependent = 'PSC12_binary',
  independent = independent
  )

baseline_plots_physical_uc$AgeGroup
baseline_plots_physical_uc$Sex
baseline_plots_physical_uc$flare_group
baseline_plots_physical_uc$cat
```

#### Crohn's Disease

```{r}
# CD patients only
data_baseline_cd <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'CD')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is PSC12_binary

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_cd, 
  dependent = 'PSC12_binary', 
  independent = independent)  %>%
  print_chisq()

# Plots
baseline_plots_physical_cd <- summon_baseline_plots(
  data = data_baseline_cd, 
  dependent = 'PSC12_binary',
  independent = independent
  )

baseline_plots_physical_cd$AgeGroup
baseline_plots_physical_cd$Sex
baseline_plots_physical_cd$flare_group
baseline_plots_physical_cd$cat
```

### Mental Component (categorical)

#### Full Cohort

```{r}
# Rename QOL
data_baseline <- QOL

# Independent variables
independent = c('diagnosis2',
                'AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is MCS12_binary

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline, 
  dependent = 'MCS12_binary', 
  independent = independent) %>%
  print_chisq()

# Plots
baseline_plots_mental <- summon_baseline_plots(
  data = data_baseline, 
  dependent = 'MCS12_binary',
  independent = independent
  )

baseline_plots_mental$diagnosis2
baseline_plots_mental$AgeGroup
baseline_plots_mental$Sex
baseline_plots_mental$flare_group
baseline_plots_mental$cat
```

#### Ulcerative Colitis

```{r}
# UC patients only
data_baseline_uc <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is MCS12_binary

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_uc, 
  dependent = 'MCS12_binary', 
  independent = independent)  %>%
  print_chisq()

# Plots
baseline_plots_mental_uc <- summon_baseline_plots(
  data = data_baseline_uc, 
  dependent = 'MCS12_binary',
  independent = independent
  )

baseline_plots_mental_uc$AgeGroup
baseline_plots_mental_uc$Sex
baseline_plots_mental_uc$flare_group
baseline_plots_mental_uc$cat
```

#### Crohn's Disease

```{r}
# CD patients only
data_baseline_cd <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'CD')

# Independent variables
independent = c('AgeGroup',
                'Sex',
                'flare_group',
                'cat')

# Dependent variable is MCS12_binary

# Chi-squared tests for each independent variable with score_group
# With p-values adjusted for multiple testing using holm method.
summon_chisq_test(
  data = data_baseline_cd, 
  dependent = 'MCS12_binary', 
  independent = independent)  %>%
  print_chisq()

# Plots
baseline_plots_mental_cd <- summon_baseline_plots(
  data = data_baseline_cd, 
  dependent = 'MCS12_binary',
  independent = independent
  ) 

baseline_plots_mental_cd$AgeGroup
baseline_plots_mental_cd$Sex
baseline_plots_mental_cd$flare_group
baseline_plots_mental_cd$cat
```

## Survival Analysis

### Physical Component (categorical)

#### Full Cohort

```{r}
# Survival data
# Patient reported flares
data_survival_soft <- QOL %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard <- QOL %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Soft
# Kaplan Meier
data_survival_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ PSC12_binary, data = .) %>%
  ggsurvplot(data = data_survival_soft,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_physical_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    PSC12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft
)

cox_physical_soft %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ PSC12_binary, data = .) %>%
  ggsurvplot(data = data_survival_hard,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_physical_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    PSC12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard
)

cox_physical_hard %>%
  print_survival()
```

#### Ulcerative Colitis

```{r}
# Survival data
# Patient reported flares
data_survival_soft_uc <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard_uc <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Soft
# Kaplan Meier
data_survival_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ PSC12_binary, data = .) %>%
  ggsurvplot(data = data_survival_soft_uc,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_physical_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    PSC12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft_uc
)

cox_physical_soft_uc %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ PSC12_binary, data = .) %>%
  ggsurvplot(data = data_survival_hard_uc,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_physical_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    PSC12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

cox_physical_hard_uc %>%
  print_survival()
```

#### Crohn's Disease

```{r}
# Survival data
# Patient reported flares
data_survival_soft_cd <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard_cd <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

```

Patient reported flare.

```{r}
# Soft
# Kaplan Meier
data_survival_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ PSC12_binary, data = .) %>%
  ggsurvplot(data = data_survival_soft_cd,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_physical_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    PSC12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft_cd
)

cox_physical_soft_cd %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ PSC12_binary, data = .) %>%
  ggsurvplot(data = data_survival_hard_cd,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_physical_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    PSC12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard_cd
)

cox_physical_hard_cd %>%
  print_survival()
```

### Mental Component (categorical)

#### Full Cohort

```{r}
# Survival data
# Patient reported flares
data_survival_soft <- QOL %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard <- QOL %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Soft
# Kaplan Meier
data_survival_soft %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MCS12_binary, data = .) %>%
  ggsurvplot(data = data_survival_soft,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_mental_soft <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MCS12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft
)

cox_mental_soft %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MCS12_binary, data = .) %>%
  ggsurvplot(data = data_survival_hard,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_mental_hard <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MCS12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard
)

cox_mental_hard %>%
  print_survival()
```

#### Ulcerative Colitis

```{r}
# Survival data
# Patient reported flares
data_survival_soft_uc <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard_uc <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'UC/IBDU') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

```

Patient reported flare.

```{r}
# Soft
# Kaplan Meier
data_survival_soft_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MCS12_binary, data = .) %>%
  ggsurvplot(data = data_survival_soft_uc,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_mental_soft_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MCS12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft_uc
)

cox_mental_soft_uc %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard_uc %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MCS12_binary, data = .) %>%
  ggsurvplot(data = data_survival_hard_uc,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_mental_hard_uc <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MCS12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard_uc
)

cox_mental_hard_uc %>%
  print_survival()
```

#### Crohn's Disease

```{r}
# Survival data
# Patient reported flares
data_survival_soft_cd <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

# Hard flares
data_survival_hard_cd <- data_baseline %>%
  dplyr::filter(diagnosis2 == 'CD') %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

Patient reported flare.

```{r}
# Soft
# Kaplan Meier
data_survival_soft_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MCS12_binary, data = .) %>%
  ggsurvplot(data = data_survival_soft_cd,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox model
cox_mental_soft_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MCS12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_soft_cd
)

cox_mental_soft_cd %>%
  print_survival()
```

Hard flare.

```{r}
# KM
data_survival_hard_cd %>%
  survfit(Surv(time, DiseaseFlareYN) ~ MCS12_binary, data = .) %>%
  ggsurvplot(data = data_survival_hard_cd,
             conf.int = TRUE,
             ggtheme = theme_minimal())

# Cox
cox_mental_hard_cd <- coxph(
  Surv(time, DiseaseFlareYN) ~
    MCS12_binary +
    IMD +
    Sex +
    AgeGroup +
    cat +
    frailty(SiteNo),
  data = data_survival_hard_cd
)

cox_mental_hard_cd %>%
  print_survival()
```

## Reproduction and reproducibility {.appendix}

<details class = "appendix">
  <summary> Session info </summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>

Licensed by 
<a href="https://creativecommons.org/licenses/by/4.0/">CC BY</a>
 unless otherwise stated.
