---
title: "Longitudinal analysis"
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-depth: 4
toc-expand: true
editor_options: 
  chunk_output_type: console
---

# PHQ

## Setting up in R

Load R packages required.

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(magrittr)
library(survival)
library(patchwork)

```

## Helper functions

```{r}
print_survival <- function(data) {
  data %>%
    broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    # Relocate columns
    dplyr::select(term, estimate, std.error, statistic, conf.low, conf.high, p.value) %>%
    # 3 significant figures
    dplyr::mutate(dplyr::across(.cols = where(is.numeric), .fns = ~ signif(.x, 3))) %>%
    dplyr::mutate(p.value = as.character(p.value)) %>%
    knitr::kable(col.names = c("Covariate",
                               "Hazard Ratio",
                               "Std. Error",
                               "Statistic",
                               "Lower 95% CI",
                               "Upper 95% CI",
                               "P-value"),
                 format.args = list(drop0trailing = TRUE)
                 )
}
```

## Load data

```{r}
chiara <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
alex_data <- "/Volumes/igmm/cvallejo-predicct/people/Alex/Predicct2/Data"


# Load data
phq <- readr::read_rds(
  glue::glue("{alex_data}/PHQ.rds")
)
phq_followup <- readr::read_rds(
  glue::glue("{alex_data}/phq_followup.rds")
)

flares_hard <- readRDS(paste0(chiara, "flares_hard.RDS"))
flares_soft <- readRDS(paste0(chiara, "flares_soft.RDS"))


```

## Prepare longitudinal data

```{r}
phq <- phq %>% 
  filter(!is.na(ParticipantId))

# Join longitudinal and baseline
phq %<>%
  dplyr::left_join(
    phq_followup, by = "ParticipantNo"
  )

# Survival data
data_soft <- phq %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::rename(DiseaseFlareYN = softflare, time = softflare_time)


data_hard <- phq %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::rename(DiseaseFlareYN = hardflare, time = hardflare_time)


# Create generic long-format PHQ data

phq_long <- phq %>%
  dplyr::rename(
    somatisation_0 = somatisation,
    TotalPHQ_0 = TotalPHQ
  ) %>%
  tidyr::pivot_longer(
    cols = c(somatisation_0, somatisation_12, somatisation_24, TotalPHQ_0, TotalPHQ_12, TotalPHQ_24),
    names_to = c(".value", "month"),
    names_sep = "_"
  ) %>%
  # Month numeric
  dplyr::mutate(
    month = as.numeric(month)
  )

# Add soft-flare information to the long-format PHQ data
data_soft_long <- phq_long %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(
      ParticipantNo,
      softflare,
      softflare_time
    ),
    by = "ParticipantNo"
  ) %>%
  dplyr::rename(
    DiseaseFlareYN = softflare,
    time = softflare_time
  )

# Add hard-flare information to the long-format PHQ data
data_hard_long <- phq_long %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(
      ParticipantNo,
      hardflare,
      hardflare_time
    ),
    by = "ParticipantNo"
  ) %>%
  dplyr::rename(
    DiseaseFlareYN = hardflare,
    time = hardflare_time
  )
```

## Counts

```{r}
custom_theme = theme_minimal() + 
  theme(
    title = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "top",
    plot.tag.position = c(0.55, 0.85),
    plot.tag = element_text(size = 12)
  )

palette <- c("#009E73", "#0072B2", "#FFA500", "grey")


# Complete for missing
data_soft_long_counts <- data_soft_long %>%
  tidyr::complete(ParticipantNo, month)

# Fill in the DiseaseFlare and time
data_soft_long_counts %<>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(DiseaseFlareYN, time) %>%
  dplyr::ungroup()

# Post flare flag
data_soft_long_counts %<>%
  dplyr::mutate(
    post_flare_flag = (DiseaseFlareYN == 1 & time < as.numeric(month)*365/12)
    ) %>%
  # still at risk
  dplyr::mutate(
    at_risk_flag = (time >= as.numeric(month)*365/12)
  )

# Month
data_soft_long_counts %<>%
  dplyr::mutate(month = factor(month))

data_soft_long_counts %<>%
  dplyr::mutate(
    somatisation = forcats::fct_recode(
      somatisation, "Moderate/Severe" = "ModSev"
    )
  )

# No previous flare
plot_soft_pre <- data_soft_long_counts %>%
  dplyr::filter(
    month != 0,
    at_risk_flag == TRUE) %>%
  dplyr::mutate(
    somatisation = forcats::fct_na_value_to_level(somatisation, "Missing"),
    somatisation = forcats::fct_relevel(
      somatisation, "Missing", "Moderate/Severe", "Mild", "None"
    )
  ) %>%
  dplyr::count(month, somatisation) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[somatisation != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      somatisation == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = somatisation)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, 250),
    limits = c(0, 1500)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("None", "Mild", "Moderate/Severe", "Missing")  
    ) +
  labs(
    fill = "Somatisation",
    tag = "Still at risk of patient-reported flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme


# Post flare
plot_soft_post <- data_soft_long_counts %>%
  dplyr::filter(
    month != 0,
    post_flare_flag == TRUE) %>%
  dplyr::mutate(
    somatisation = forcats::fct_na_value_to_level(somatisation, "Missing"),
    somatisation = forcats::fct_relevel(
      somatisation, "Missing", "Moderate/Severe", "Mild", "None"
    )
  ) %>%
  dplyr::count(month, somatisation) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[somatisation != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      somatisation == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = somatisation)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 700, 100),
    limits = c(0, 700)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("None", "Mild", "Moderate/Severe", "Missing")
  ) +
  labs(
    fill = "Somatisation",
    tag = "Post patient-reported flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme






# Hard
# Complete for missing
data_hard_long_counts <- data_hard_long %>%
  tidyr::complete(ParticipantNo, month)

# Fill in the DiseaseFlare and time
data_hard_long_counts %<>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(DiseaseFlareYN, time) %>%
  dplyr::ungroup()



# Post flare flag
data_hard_long_counts %<>%
  dplyr::mutate(
    post_flare_flag = (DiseaseFlareYN == 1 & time < as.numeric(month)*365/12)
    ) %>%
  # still at risk
  dplyr::mutate(
    at_risk_flag = (time >= as.numeric(month)*365/12)
  )

# Month
data_hard_long_counts %<>%
  dplyr::mutate(month = factor(month))

data_hard_long_counts %<>%
  dplyr::mutate(
    somatisation = forcats::fct_recode(
      somatisation, "Moderate/Severe" = "ModSev"
    )
  )

  

# No previous flare
plot_hard_pre <- data_hard_long_counts %>%
  dplyr::filter(
    month != 0,
    at_risk_flag == TRUE) %>%
  dplyr::mutate(
    somatisation = forcats::fct_na_value_to_level(somatisation, "Missing"),
    somatisation = forcats::fct_relevel(
      somatisation, "Missing", "Moderate/Severe", "Mild", "None"
    )
  ) %>%
  dplyr::count(month, somatisation) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[somatisation != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      somatisation == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = somatisation)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 1750, 250),
    limits = c(0, 1750)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("None", "Mild", "Moderate/Severe", "Missing")  
    ) +
  labs(
    fill = "Somatisation",
    tag = "Still at risk of objective flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme

# Post flare
plot_hard_post <- data_hard_long_counts %>%
  dplyr::filter(
    month != 0,
    post_flare_flag == TRUE) %>%
  dplyr::mutate(
    somatisation = forcats::fct_na_value_to_level(somatisation, "Missing"),
    somatisation = forcats::fct_relevel(
      somatisation, "Missing", "Moderate/Severe", "Mild", "None"
    )
  ) %>%
  dplyr::count(month, somatisation) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[somatisation != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      somatisation == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = somatisation)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 250, 50),
    limits = c(0, 250)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("None", "Mild", "Moderate/Severe", "Missing")
  ) +
  labs(
    fill = "Somatisation",
    tag = "Post objective flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme



# Together
plot_soft_pre + plot_hard_pre + plot_soft_post + plot_hard_post +
  patchwork::plot_layout(
    ncol = 2, guides = 'collect'
  ) &
  theme(legend.position = "top")

# Save 7.5x7.5

```

## Extended Cox model with time-dependent covariates

### Data preparation

```{r}
data_soft_static <- data_soft %>%
  dplyr::select(
    ParticipantNo, SiteNo, diagnosis2, Sex, age_decade, FC, Smoke, IMD,
    flare_group, OverallControl, control_8, time, DiseaseFlareYN
  ) %>%
  # Remove people with time 0
  dplyr::filter(
    time > 0
  )

data_soft_td <- data_soft_long %>%
  dplyr::select(
    ParticipantNo, month, somatisation
  ) %>%
  # Last observation carried forward
  dplyr::arrange(ParticipantNo, month) %>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(somatisation, .direction = "down") %>%
  # Convert month to days to match time
  dplyr::mutate(
    month = as.numeric(month),
    day = 365*month/12)

# Start stop
data_soft_merged <- tmerge(
  data_soft_static, 
  data_soft_static, 
  id = ParticipantNo, 
  endpoint = event(time, DiseaseFlareYN)) %>%
  tmerge(
    data2 = data_soft_td,
    id = ParticipantNo,
    somatisation = tdc(day, somatisation)
  )


## Hard ####
data_hard_static <- data_hard %>%
  dplyr::select(
    ParticipantNo, SiteNo, diagnosis2, Sex, age_decade, FC, Smoke, IMD,
    flare_group, OverallControl, control_8, time, DiseaseFlareYN
  ) %>%
  # Remove people with time 0
  dplyr::filter(
    time > 0
  )

data_hard_td <- data_hard_long %>%
  dplyr::select(
    ParticipantNo, month, somatisation
  ) %>%
  # Last observation carried forward
  dplyr::arrange(ParticipantNo, month) %>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(somatisation, .direction = "down") %>%
  # Convert month to days to match time
  dplyr::mutate(
    month = as.numeric(month),
    day = 365*month/12)

# Start stop
data_hard_merged <- tmerge(
  data_hard_static, 
  data_hard_static, 
  id = ParticipantNo, 
  endpoint = event(time, DiseaseFlareYN)) %>%
  tmerge(
    data2 = data_hard_td,
    id = ParticipantNo,
    somatisation = tdc(day, somatisation)
  )


```

### Results

#### Crohn's disease

```{r}
# Time dependent cox model
# CD
data_soft_merged_cd <- data_soft_merged %>% dplyr::filter(diagnosis2 == 'CD')
data_hard_merged_cd <- data_hard_merged %>% dplyr::filter(diagnosis2 == 'CD')
```

##### Patient-reported flare

```{r}

cox_soft_cd <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    somatisation +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_soft_merged_cd
)

cox_soft_cd %>%
  print_survival()

```

##### Objective flare

```{r}
 
cox_hard_cd <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    somatisation +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_hard_merged_cd
)

cox_hard_cd %>%
  print_survival()

```

#### Ulcerative colitis

```{r}
# Time dependent cox model
data_soft_merged_uc <- data_soft_merged %>% dplyr::filter(diagnosis2 == 'UC/IBDU')
data_hard_merged_uc <- data_hard_merged %>% dplyr::filter(diagnosis2 == 'UC/IBDU')
```

##### Patient-reported flare

```{r}

cox_soft_uc <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    somatisation +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_soft_merged_uc
)

cox_soft_uc %>%
  print_survival()

```

##### Objective flare

```{r}
 
cox_hard_uc <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    somatisation +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_hard_merged_uc
)

cox_hard_uc %>%
  print_survival()

```


### Results - Extended analysis

#### Crohn's disease

```{r}
data_soft_merged_cd_extended <- data_soft_merged %>%
  dplyr::filter(diagnosis2 == "CD")

data_hard_merged_cd_extended <- data_hard_merged %>%
  dplyr::filter(diagnosis2 == "CD")
```

##### Patient-reported flare

```{r}
cox_soft_cd_extended <- coxph(
  Surv(tstart, tstop, endpoint) ~
    somatisation +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    control_8 +
    Smoke +
    frailty(SiteNo),
  data = data_soft_merged_cd_extended
)

cox_soft_cd_extended %>%
  print_survival()
```

##### Objective flare

```{r}
cox_hard_cd_extended <- coxph(
  Surv(tstart, tstop, endpoint) ~
    somatisation +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    control_8 +
    Smoke +
    frailty(SiteNo),
  data = data_hard_merged_cd_extended
)

cox_hard_cd_extended %>%
  print_survival()
```

#### Ulcerative colitis

```{r}
data_soft_merged_uc_extended <- data_soft_merged %>%
  dplyr::filter(diagnosis2 == "UC/IBDU")

data_hard_merged_uc_extended <- data_hard_merged %>%
  dplyr::filter(diagnosis2 == "UC/IBDU")
```

##### Patient-reported flare

```{r}
cox_soft_uc_extended <- coxph(
  Surv(tstart, tstop, endpoint) ~
    somatisation +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    control_8 +
    Smoke +
    frailty(SiteNo),
  data = data_soft_merged_uc_extended
)

cox_soft_uc_extended %>%
  print_survival()
```

##### Objective flare

```{r}
cox_hard_uc_extended <- coxph(
  Surv(tstart, tstop, endpoint) ~
    somatisation +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    control_8 +
    Smoke +
    frailty(SiteNo),
  data = data_hard_merged_uc_extended
)

cox_hard_uc_extended %>%
  print_survival()
```
