---
title: "Longitudinal analysis"
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-depth: 4
toc-expand: true
editor_options: 
  chunk_output_type: console
---

# HADS

## Setting up in R

Load R packages required.

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(magrittr)
library(readxl)
library(openxlsx)
library(survival)
library(patchwork)

```

## Helper functions

```{r}
print_survival <- function(data) {
  data %>%
    broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
    # Relocate columns
    dplyr::select(term, estimate, std.error, statistic, conf.low, conf.high, p.value) %>%
    # 3 significant figures
    dplyr::mutate(dplyr::across(.cols = where(is.numeric), .fns = ~ signif(.x, 3))) %>%
    dplyr::mutate(p.value = as.character(p.value)) %>%
    knitr::kable(col.names = c("Covariate",
                               "Hazard Ratio",
                               "Std. Error",
                               "Statistic",
                               "Lower 95% CI",
                               "Upper 95% CI",
                               "P-value"),
                 format.args = list(drop0trailing = TRUE)
                 )
}
```

## Load data

```{r}
# Load data ####
data.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20221004/"
redcap.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20231030/"
prefix <- "/Volumes/igmm/cvallejo-predicct/predicct/end-of-follow-up/"
outdir <- "/Volumes/igmm/cvallejo-predicct/predicct/processed/"
chiara <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"

# Baseline hads
hads <- readxl::read_xlsx(paste0(data.path, "Baseline2022/hads.xlsx"))

# Longitudinal hads
hads_followup <- readxl::read_xlsx(paste0(data.path, "Followup/hads.xlsx"))

# Demographics
demographics <- read.xlsx(paste0(data.path, "Baseline2022/demographics2022.xlsx"))
IBD <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx"))
smoking <- readRDS(paste0(chiara, "smoking.rds"))
IMD <- readRDS(paste0(chiara, "IMD.rds"))
demo <- readRDS(paste0(outdir, "demo.RDS"))
flares <- readRDS(paste0(outdir, "flares/cutoff-flares.RDS"))
IBD_C <- readRDS(paste0(chiara, "IBD_C.RDS"))
IBD_FC <- readRDS(paste0(chiara, "IBD_FC.RDS"))
flares_soft <- readRDS(paste0(chiara, "flares_soft.RDS"))
flares_hard <- readRDS(paste0(chiara, "flares_hard.RDS"))

```

## Data cleaning

```{r}
# Data cleaning ####

# Cleaning the longitudinal hads
hads_followup %<>% 
  filter(!is.na(ParticipantId))

# Rescale
hads_followup %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = FeelTense:CanEnjoyBookTV,
      .fns = ~. - 1)
  )

#correcting inversely coded HADS
hads_followup %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = c("FeelTense", 
                "FrightenedFeelingSomethingAwful", 
                "WorryingThoughts", 
                "FeelCheerful", 
                "FeelSlowedDown",
                "LostInterestAppearance",
                "FeelRestless",
                "SuddenFeelingsPanic"),
      .fns = ~ 3 - .
    )
  )

#anxiety HADS composite score
hads_followup %<>% 
  dplyr::mutate(anxiety_hads = FeelTense + 
                  FrightenedFeelingSomethingAwful + 
                  WorryingThoughts +
                  SitAtEase +
                  FrightenedFeelingButterflies +
                  FeelRestless +
                  SuddenFeelingsPanic)

#depression HADS composite score
hads_followup %<>% 
  dplyr::mutate(depression_hads = EnjoyThings +
                  CanLaugh +
                  FeelCheerful +
                  FeelSlowedDown +
                  LostInterestAppearance +
                  LookForward +
                  CanEnjoyBookTV)

#remove breakdown hads columns
hads_followup %<>%
  dplyr::select(ParticipantNo, Q_month, anxiety_hads, depression_hads)

# Wide format
hads_followup %<>%
  tidyr::pivot_wider(names_from = Q_month, values_from = c(anxiety_hads, depression_hads))



hads <- hads %>%
  filter(!is.na(ParticipantId))

demographics <- demographics %>% 
  filter(!is.na(ParticipantId))

IBD <- IBD %>% 
  filter(!is.na(ParticipantId))

# join demographics
hads %<>%
  dplyr::left_join(demographics %>%
                     dplyr::select(ParticipantNo, Sex, age, diagnosis),
                   by = "ParticipantNo")

# Convert diagnosis column to factor with labels
hads %<>%
  dplyr::mutate(
    diagnosis2 = dplyr::case_when(
      diagnosis == 1 ~ 1,
      diagnosis == 2 ~ 2,
      diagnosis == 3 ~ 2,
      diagnosis == 4 ~ 2
    )
  ) %>%
  dplyr::mutate(
    diagnosis2 = factor(
      diagnosis2,
      levels = c("1", "2"), 
      labels = c("CD", "UC/IBDU")
    )
  )

hads %<>%
  dplyr::mutate(Sex = factor(
    Sex,
    levels = c(1, 2),
    labels = c("Male", "Female")
  )
  )

# Exclude underage
hads %<>%
  dplyr::filter(age > 17)

# Rescale
hads %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = FeelTense:CanEnjoyBookTV,
      .fns = ~. - 1)
  )

#correcting inversely coded HADS
hads %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = c("FeelTense", 
                "FrightenedFeelingSomethingAwful", 
                "WorryingThoughts", 
                "FeelCheerful", 
                "FeelSlowedDown",
                "LostInterestAppearance",
                "FeelRestless",
                "SuddenFeelingsPanic"),
      .fns = ~ 3 - .
    )
  )

#anxiety HADS composite score
hads %<>% 
  dplyr::mutate(anxiety_hads = FeelTense + 
                  FrightenedFeelingSomethingAwful + 
                  WorryingThoughts +
                  SitAtEase +
                  FrightenedFeelingButterflies +
                  FeelRestless +
                  SuddenFeelingsPanic)

#depression HADS composite score
hads %<>% 
  dplyr::mutate(depression_hads = EnjoyThings +
                  CanLaugh +
                  FeelCheerful +
                  FeelSlowedDown +
                  LostInterestAppearance +
                  LookForward +
                  CanEnjoyBookTV)

#remove breakdown hads columns
hads %<>%
  dplyr::select(-c(FeelTense:CanEnjoyBookTV))

# Add the follow up hads
hads %<>% 
  dplyr::left_join(
    hads_followup, by = "ParticipantNo"
  )
  
# FC data
hads <- hads %>% 
  dplyr::left_join(
    demo %>% dplyr::select(ParticipantNo, FC, cat), 
    by = "ParticipantNo")

# Add smoking
hads %<>% 
  dplyr::left_join(
    smoking %>% dplyr::select(ParticipantNo, Smoke), 
    by = "ParticipantNo")

# Smoke as a factor
hads %<>%
  dplyr::mutate(Smoke = forcats::as_factor(Smoke)) %>%
  dplyr::mutate(Smoke = forcats::fct_relevel(Smoke, 'Never', 'Previous', 'Current'))

# Add IMD
hads %<>% 
  dplyr::left_join(
    IMD, 
    by = "ParticipantNo") 

# IMD as a factor
hads %<>%
  dplyr::mutate(
    IMD = as.factor(IMD)
  )

# Age in decades
hads %<>%
  dplyr::mutate(
    age_decade = age/10
  )


# FC maximum detectable value is 1250
hads %<>%
  dplyr::mutate(
    FC = dplyr::case_when(
      FC > 1250 ~ 1250,
      .default = FC
    )
  )

# Log transform FC due to extreme positive skew
hads %<>%
  dplyr::mutate(
    FC = log(FC)
  )

hads %<>%
  dplyr::filter(!(is.na(anxiety_hads) & is.na(depression_hads)))


# Save hads in long format

hads_long <- hads %>%
  dplyr::rename(
    anxiety_0 = anxiety_hads,
    anxiety_12 = anxiety_hads_12,
    anxiety_24 = anxiety_hads_24,
    depression_0 = depression_hads,
    depression_12 = depression_hads_12,
    depression_24 = depression_hads_24,
  ) %>%
  tidyr::pivot_longer(
    cols = c(anxiety_0, anxiety_12, anxiety_24, depression_0, depression_12, depression_24),
    names_to = c(".value", "month"),
    names_sep = "_"
  ) %>%
  # Month numeric
  dplyr::mutate(
    month = as.numeric(month)
  )


# Survival data
data_anxiety_soft <- hads %>% 
  dplyr::select(-tidyselect::starts_with("depression")) %>%
  dplyr::inner_join(
  flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
  by = 'ParticipantNo'
) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

data_anxiety_hard <- hads %>% 
  dplyr::select(-tidyselect::starts_with("depression")) %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

# Depression
data_depression_soft <- hads %>% 
  dplyr::select(-tidyselect::starts_with("anxiety")) %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

data_depression_hard <- hads %>% 
  dplyr::select(-tidyselect::starts_with("anxiety")) %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = 'ParticipantNo'
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)


# Long format
# Anxiety
data_anxiety_soft_long <- data_anxiety_soft %>%
  dplyr::rename(anxiety_hads_0 = anxiety_hads) %>%
  tidyr::pivot_longer(
    cols = c(anxiety_hads_0, anxiety_hads_12, anxiety_hads_24),
    names_to = "month",
    names_prefix = "anxiety_hads_",
    values_to = "anxiety_hads"
  )

data_anxiety_hard_long <- data_anxiety_hard %>%
  dplyr::rename(anxiety_hads_0 = anxiety_hads) %>%
  tidyr::pivot_longer(
    cols = c(anxiety_hads_0, anxiety_hads_12, anxiety_hads_24),
    names_to = "month",
    names_prefix = "anxiety_hads_",
    values_to = "anxiety_hads"
  )

# Depression
data_depression_soft_long <- data_depression_soft %>%
  dplyr::rename(depression_hads_0 = depression_hads) %>%
  tidyr::pivot_longer(
    cols = c(depression_hads_0, depression_hads_12, depression_hads_24),
    names_to = "month",
    names_prefix = "depression_hads_",
    values_to = "depression_hads"
  )

data_depression_hard_long <- data_depression_hard %>%
  dplyr::rename(depression_hads_0 = depression_hads) %>%
  tidyr::pivot_longer(
    cols = c(depression_hads_0, depression_hads_12, depression_hads_24),
    names_to = "month",
    names_prefix = "depression_hads_",
    values_to = "depression_hads"
  )
```

## Counts

```{r}
#| message: false
#| warning: false
custom_theme = theme_minimal() + 
  theme(
    title = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "top",
    plot.tag.position = c(0.55, 0.85),
    plot.tag = element_text(size = 12)
  )

palette <- c("#FFA500", "#0072B2", "grey", "#009E73")

data_anxiety_soft_long %<>% 
  # Anxiety groups
  dplyr::mutate(
    score_group = cut(
      anxiety_hads, 
      breaks = c(0, 7, 21), 
      labels = c("0-7", "8-21"), 
      include.lowest = TRUE)
  )

data_depression_soft_long %<>% 
  # Anxiety groups
  dplyr::mutate(
    score_group = cut(
      depression_hads, 
      breaks = c(0, 7, 21), 
      labels = c("0-7", "8-21"), 
      include.lowest = TRUE)
  )


# Post flare flag
data_anxiety_soft_long %<>%
  dplyr::mutate(
    post_flare_flag = (DiseaseFlareYN == 1 & time < as.numeric(month)*365/12)
  ) %>%
  # still at risk
  dplyr::mutate(
    at_risk_flag = (time >= as.numeric(month)*365/12)
  )

data_depression_soft_long %<>%
  dplyr::mutate(
    post_flare_flag = (DiseaseFlareYN == 1 & time < as.numeric(month)*365/12)
  ) %>%
  # still at risk
  dplyr::mutate(
    at_risk_flag = (time >= as.numeric(month)*365/12)
  )

# Anxiety ####
# No previous flare
plot_anxiety_soft_pre <- data_anxiety_soft_long %>%
  dplyr::filter(
    month != 0,
    at_risk_flag == TRUE) %>%
  dplyr::mutate(
    score_group = forcats::fct_na_value_to_level(score_group, "Missing"),
    score_group = forcats::fct_relevel(score_group, "Missing")
  ) %>%
  dplyr::count(month, score_group) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[score_group != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      score_group == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = score_group)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, 250),
    limits = c(0, 1500)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("0-7", "8-21", "Missing")
  ) +
  labs(
    fill = "Anxiety HADS",
    tag = "Still at risk of patient-reported flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme


# Post flare
plot_anxiety_soft_post <- data_anxiety_soft_long %>%
  dplyr::filter(
    month != 0,
    post_flare_flag == TRUE) %>%
  dplyr::mutate(
    score_group = forcats::fct_na_value_to_level(score_group, "Missing"),
    score_group = forcats::fct_relevel(score_group, "Missing")
  ) %>%
  dplyr::count(month, score_group) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[score_group != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      score_group == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = score_group)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 700, 100),
    limits = c(0, 700)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("0-7", "8-21", "Missing")
  ) +
  labs(
    fill = "Anxiety HADS",
    tag = "Post patient-reported flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme


# Depression ####
# No previous flare
plot_depression_soft_pre <- data_depression_soft_long %>%
  dplyr::filter(
    month != 0,
    at_risk_flag == TRUE) %>%
  dplyr::mutate(
    score_group = forcats::fct_na_value_to_level(score_group, "Missing"),
    score_group = forcats::fct_relevel(score_group, "Missing")
  ) %>%
  dplyr::count(month, score_group) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[score_group != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      score_group == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = score_group)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, 250),
    limits = c(0, 1500)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("0-7", "8-21", "Missing")
  ) +
  labs(
    fill = "Depression HADS",
    tag = "Still at risk of patient-reported flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme

# Post flare
plot_depression_soft_post <- data_depression_soft_long %>%
  dplyr::filter(
    month != 0,
    post_flare_flag == TRUE) %>%
  dplyr::mutate(
    score_group = forcats::fct_na_value_to_level(score_group, "Missing"),
    score_group = forcats::fct_relevel(score_group, "Missing")
  ) %>%
  dplyr::count(month, score_group) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[score_group != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      score_group == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = score_group)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 700, 100),
    limits = c(0, 700)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("0-7", "8-21", "Missing")
  ) +
  labs(
    fill = "Depression HADS",
    tag = "Post patient-reported flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme

```

```{r}
data_anxiety_hard_long %<>% 
  # Anxiety groups
  dplyr::mutate(
    score_group = cut(
      anxiety_hads, 
      breaks = c(0, 7, 21), 
      labels = c("0-7", "8-21"), 
      include.lowest = TRUE)
  )

data_depression_hard_long %<>% 
  # Anxiety groups
  dplyr::mutate(
    score_group = cut(
      depression_hads, 
      breaks = c(0, 7, 21), 
      labels = c("0-7", "8-21"), 
      include.lowest = TRUE)
  )


# Post flare flag
data_anxiety_hard_long %<>%
  dplyr::mutate(
    post_flare_flag = (DiseaseFlareYN == 1 & time < as.numeric(month)*365/12)
  ) %>%
  # still at risk
  dplyr::mutate(
    at_risk_flag = (time >= as.numeric(month)*365/12)
  )

data_depression_hard_long %<>%
  dplyr::mutate(
    post_flare_flag = (DiseaseFlareYN == 1 & time < as.numeric(month)*365/12)
  ) %>%
  # still at risk
  dplyr::mutate(
    at_risk_flag = (time >= as.numeric(month)*365/12)
  )

# Anxiety ####
# No previous flare
plot_anxiety_hard_pre <- data_anxiety_hard_long %>%
  dplyr::filter(
    month != 0,
    at_risk_flag == TRUE) %>%
  dplyr::mutate(
    score_group = forcats::fct_na_value_to_level(score_group, "Missing"),
    score_group = forcats::fct_relevel(score_group, "Missing")
  ) %>%
  dplyr::count(month, score_group) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[score_group != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
    ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      score_group == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = score_group)) +
  geom_col(width = 0.8, alpha = 0.5) +
    geom_text(
      aes(x = month, y = mid, label = percent),
      size = 3.5
    ) +
  scale_y_continuous(
    breaks = seq(0, 1750, 250),
    limits = c(0, 1750)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("0-7", "8-21", "Missing")
  ) +
  labs(
    fill = "Anxiety HADS",
    tag = "Still at risk of objective flare"
    ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme


# Post flare
plot_anxiety_hard_post <- data_anxiety_hard_long %>%
  dplyr::filter(
    month != 0,
    post_flare_flag == TRUE) %>%
  dplyr::mutate(
    score_group = forcats::fct_na_value_to_level(score_group, "Missing"),
    score_group = forcats::fct_relevel(score_group, "Missing")
  ) %>%
  dplyr::count(month, score_group) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[score_group != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      score_group == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = score_group)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 250, 50),
    limits = c(0, 250)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("0-7", "8-21", "Missing")
  ) +
  labs(
    fill = "Anxiety HADS",
    tag = "Post objective flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme


# Depression ####
# No previous flare
plot_depression_hard_pre <- data_depression_hard_long %>%
  dplyr::filter(
    month != 0,
    at_risk_flag == TRUE) %>%
  dplyr::mutate(
    score_group = forcats::fct_na_value_to_level(score_group, "Missing"),
    score_group = forcats::fct_relevel(score_group, "Missing")
  ) %>%
  dplyr::count(month, score_group) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[score_group != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      score_group == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = score_group)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 1750, 250),
    limits = c(0, 1750)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("0-7", "8-21", "Missing")
  ) +
  labs(
    fill = "Depression HADS",
    tag = "Still at risk of objective flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme


# Post flare
plot_depression_hard_post <- data_depression_hard_long %>%
  dplyr::filter(
    month != 0,
    post_flare_flag == TRUE) %>%
  dplyr::mutate(
    score_group = forcats::fct_na_value_to_level(score_group, "Missing"),
    score_group = forcats::fct_relevel(score_group, "Missing")
  ) %>%
  dplyr::count(month, score_group) %>%
  # Calculate percentage
  dplyr::group_by(month) %>%
  dplyr::mutate(
    p = n/sum(n[score_group != "Missing"]),
    mid = rev(cumsum(rev(n))) - n/2,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(    
    percent = glue::glue("{round(p*100)}%")
  ) %>%
  dplyr::mutate(
    percent = dplyr::case_when(
      score_group == "Missing" ~ NA,
      TRUE ~ percent
    )
  ) %>%
  ggplot(aes(x = month, y = n, fill = score_group)) +
  geom_col(width = 0.8, alpha = 0.5) +
  geom_text(
    aes(x = month, y = mid, label = percent),
    size = 3.5
  ) +
  scale_y_continuous(
    breaks = seq(0, 250, 50),
    limits = c(0, 250)
  ) +
  scale_fill_manual(
    values = palette,
    breaks = c("0-7", "8-21", "Missing")
  ) +
  labs(
    fill = "Depression HADS",
    tag = "Post objective flare"
  ) +
  xlab("Month") +
  ylab("Count") +
  custom_theme

# Together
plot_anxiety_soft_pre + plot_anxiety_hard_pre +
  plot_anxiety_soft_post + plot_anxiety_hard_post +
  patchwork::plot_layout(
    ncol = 2, 
    guides = "collect"
  ) &
  theme(legend.position = "top")

plot_depression_soft_pre + plot_depression_hard_pre +
  plot_depression_soft_post + plot_depression_hard_post +
  patchwork::plot_layout(
    ncol = 2, 
    guides = "collect"
  ) &
  theme(legend.position = "top")
```

## Extended Cox model with time-dependent covariates

### Data preparation

```{r}
# Anxiety ####
## Soft ####

data_anxiety_soft_static <- data_anxiety_soft %>%
  dplyr::select(
    ParticipantNo, SiteNo, diagnosis2, Sex, age_decade, FC, Smoke, IMD, time, DiseaseFlareYN
  ) %>%
  # Remove people with time 0
  dplyr::filter(
    time > 0
  )

data_anxiety_soft_td <- data_anxiety_soft_long %>%
  dplyr::select(
    ParticipantNo, month, anxiety_hads
  ) %>%
  # Last observation carried forward
  dplyr::arrange(ParticipantNo, month) %>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(anxiety_hads, .direction = "down") %>%
  # Convert month to days to match time
  dplyr::mutate(
    month = as.numeric(month),
    day = 365*month/12) %>%
  # Anxiety as categorical
  dplyr::mutate(
    score_group = cut(
      anxiety_hads,
      breaks = c(0, 7, 21),
      labels = c("0-7", "8-21"),
      include.lowest = TRUE
    )
  )

# Start stop
data_anxiety_soft_merged <- tmerge(
  data_anxiety_soft_static, 
  data_anxiety_soft_static, 
  id = ParticipantNo, 
  endpoint = event(time, DiseaseFlareYN)) %>%
  tmerge(
    data2 = data_anxiety_soft_td,
    id = ParticipantNo,
    score_group = tdc(day, score_group)
  )




## Hard ####
data_anxiety_hard_static <- data_anxiety_hard %>%
  dplyr::select(
    ParticipantNo, SiteNo, diagnosis2, Sex, age_decade, FC, Smoke, IMD, time, DiseaseFlareYN
  ) %>%
  # Remove people with time 0
  dplyr::filter(
    time > 0
  )

data_anxiety_hard_td <- data_anxiety_hard_long %>%
  dplyr::select(
    ParticipantNo, month, anxiety_hads
  ) %>%
  # Last observation carried forward
  dplyr::arrange(ParticipantNo, month) %>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(anxiety_hads, .direction = "down") %>%
  # Convert month to days to match time
  dplyr::mutate(
    month = as.numeric(month),
    day = 365*month/12) %>%
  # Anxiety as categorical
  dplyr::mutate(
    score_group = cut(
      anxiety_hads,
      breaks = c(0, 7, 21),
      labels = c("0-7", "8-21"),
      include.lowest = TRUE
    )
  )

# Start stop
data_anxiety_hard_merged <- tmerge(
  data_anxiety_hard_static, 
  data_anxiety_hard_static, 
  id = ParticipantNo, 
  endpoint = event(time, DiseaseFlareYN)) %>%
  tmerge(
    data2 = data_anxiety_hard_td,
    id = ParticipantNo,
    score_group = tdc(day, score_group)
  )



# Depression ####
## Soft ####

data_depression_soft_static <- data_depression_soft %>%
  dplyr::select(
    ParticipantNo, SiteNo, diagnosis2, Sex, age_decade, FC, Smoke, IMD, time, DiseaseFlareYN
  ) %>%
  # Remove people with time 0
  dplyr::filter(
    time > 0
  )

data_depression_soft_td <- data_depression_soft_long %>%
  dplyr::select(
    ParticipantNo, month, depression_hads
  ) %>%
  # Last observation carried forward
  dplyr::arrange(ParticipantNo, month) %>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(depression_hads, .direction = "down") %>%
  # Convert month to days to match time
  dplyr::mutate(
    month = as.numeric(month),
    day = 365*month/12) %>%
  # depression as categorical
  dplyr::mutate(
    score_group = cut(
      depression_hads,
      breaks = c(0, 7, 21),
      labels = c("0-7", "8-21"),
      include.lowest = TRUE
    )
  )

# Start stop
data_depression_soft_merged <- tmerge(
  data_depression_soft_static, 
  data_depression_soft_static, 
  id = ParticipantNo, 
  endpoint = event(time, DiseaseFlareYN)) %>%
  tmerge(
    data2 = data_depression_soft_td,
    id = ParticipantNo,
    score_group = tdc(day, score_group)
  )



## Hard ####
data_depression_hard_static <- data_depression_hard %>%
  dplyr::select(
    ParticipantNo, SiteNo, diagnosis2, Sex, age_decade, FC, Smoke, IMD, time, DiseaseFlareYN
  ) %>%
  # Remove people with time 0
  dplyr::filter(
    time > 0
  )

data_depression_hard_td <- data_depression_hard_long %>%
  dplyr::select(
    ParticipantNo, month, depression_hads
  ) %>%
  # Last observation carried forward
  dplyr::arrange(ParticipantNo, month) %>%
  dplyr::group_by(ParticipantNo) %>%
  tidyr::fill(depression_hads, .direction = "down") %>%
  # Convert month to days to match time
  dplyr::mutate(
    month = as.numeric(month),
    day = 365*month/12) %>%
  # depression as categorical
  dplyr::mutate(
    score_group = cut(
      depression_hads,
      breaks = c(0, 7, 21),
      labels = c("0-7", "8-21"),
      include.lowest = TRUE
    )
  )

# Start stop
data_depression_hard_merged <- tmerge(
  data_depression_hard_static, 
  data_depression_hard_static, 
  id = ParticipantNo, 
  endpoint = event(time, DiseaseFlareYN)) %>%
  tmerge(
    data2 = data_depression_hard_td,
    id = ParticipantNo,
    score_group = tdc(day, score_group)
  )

```

### Results - Anxiety

#### Crohn's disease

```{r}
# Time dependent cox model
# CD
data_anxiety_soft_merged_cd <- data_anxiety_soft_merged %>% dplyr::filter(diagnosis2 == 'CD')
data_anxiety_hard_merged_cd <- data_anxiety_hard_merged %>% dplyr::filter(diagnosis2 == 'CD')
data_depression_soft_merged_cd <- data_depression_soft_merged %>% dplyr::filter(diagnosis2 == 'CD')
data_depression_hard_merged_cd <- data_depression_hard_merged %>% dplyr::filter(diagnosis2 == 'CD')
```

##### Patient-reported flare

```{r}

cox_anxiety_soft_cd <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_anxiety_soft_merged_cd
)

cox_anxiety_soft_cd %>%
  print_survival()

```

##### Objective flare

```{r}
 
cox_anxiety_hard_cd <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_anxiety_hard_merged_cd
)

cox_anxiety_hard_cd %>%
  print_survival()

```

#### Ulcerative colitis

```{r}
# Time dependent cox model
data_anxiety_soft_merged_uc <- data_anxiety_soft_merged %>% dplyr::filter(diagnosis2 == 'UC/IBDU')
data_anxiety_hard_merged_uc <- data_anxiety_hard_merged %>% dplyr::filter(diagnosis2 == 'UC/IBDU')
data_depression_soft_merged_uc <- data_depression_soft_merged %>% dplyr::filter(diagnosis2 == 'UC/IBDU')
data_depression_hard_merged_uc <- data_depression_hard_merged %>% dplyr::filter(diagnosis2 == 'UC/IBDU')
```

##### Patient-reported flare

```{r}

cox_anxiety_soft_uc <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_anxiety_soft_merged_uc
)

cox_anxiety_soft_uc %>%
  print_survival()

```

##### Objective flare

```{r}
 
cox_anxiety_hard_uc <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_anxiety_hard_merged_uc
)

cox_anxiety_hard_uc %>%
  print_survival()

```

### Results - Depression

#### Crohn's disease

##### Patient-reported flare

```{r}

cox_depression_soft_cd <- coxph(
  Surv(tstart, tstop, endpoint) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_depression_soft_merged_cd
)

cox_depression_soft_cd %>%
  print_survival()

```

##### Objective flare

```{r}
## Hard ####

cox_depression_hard_cd <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_depression_hard_merged_cd
)

cox_depression_hard_cd %>%
  print_survival()

```

#### Ulcerative colitis

##### Patient-reported flare

```{r}

cox_depression_soft_uc <- coxph(
  Surv(tstart, tstop, endpoint) ~
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_depression_soft_merged_uc
)

cox_depression_soft_uc %>%
  print_survival()

```

##### Objective flare

```{r}
## Hard ####

cox_depression_hard_uc <- coxph(
  Surv(tstart, tstop, endpoint) ~ 
    score_group +
    IMD +
    Sex +
    age_decade +
    FC +
    Smoke +
    frailty(SiteNo),
  data = data_depression_hard_merged_uc
)

cox_depression_hard_uc %>%
  print_survival()

```
