---
title: "Fatigue interaction analysis"
author:
  - name: "Alexander Rudge"
  - name: "Chiara Cotronei"
  - name: "Nathan Constantine-Cooke"
    corresponding: true
    url: https://scholar.google.com/citations?user=2emHWR0AAAAJ&hl=en&oi=ao
format:
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-depth: 4
toc-expand: true
editor_options:
  chunk_output_type: console
---

This document tests whether associations between fatigue and subsequent flares differ by diagnosis and age.

# Setting up in R

```{r}
#| message: false
#| warning: false
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = TRUE)
options(dplyr.summarise.inform = FALSE)

if (file.exists("../../functions.R")) {
  source("../../functions.R")
} else if (file.exists("../functions.R")) {
  source("../functions.R")
} else if (file.exists("Alex Rudge/functions.R")) {
  source("Alex Rudge/functions.R")
} else {
  source("src/Psychosocial/Alex Rudge/functions.R")
}

library(tidyverse)
library(survival)
library(mice)
library(splines)
```

# Load data

```{r}
if (file.exists("/.dockerenv")) {
  alex_data <- "data/people/Alex/Predicct2/Data/"
  chiara <- "data/people/chiara/"
} else {
  alex_data <- "/Volumes/igmm/cvallejo-predicct/people/Alex/Predicct2/Data/"
  chiara <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
}

fatigue <- readr::read_rds(paste0(alex_data, "Fatigue.rds"))
flares_soft <- readr::read_rds(paste0(chiara, "flares_soft.RDS"))
flares_hard <- readr::read_rds(paste0(chiara, "flares_hard.RDS"))

data_survival_soft <- fatigue %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

data_survival_hard <- fatigue %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)
```

# Interaction with diagnosis

The models include an interaction between fatigue and `diagnosis2` to test whether the association with subsequent flares differs between UC/IBDU and CD.

## Patient-reported flare

```{r}
cox_soft_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * diagnosis2 + IMD + Sex + age_decade + FC + Smoke + frailty(SiteNo),
  data = data_survival_soft
)

cox_soft_interaction %>% print_survival()

cox_soft_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * diagnosis2 + IMD + Sex + age_decade + FC + Smoke + frailty(SiteNo),
  data = data_survival_soft,
  m = 20,
  printFlag = FALSE
)

cox_soft_interaction_pool <- with(
  cox_soft_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      OftenLackEnergy * diagnosis2 + IMD + Sex + age_decade + FC + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_soft_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

## Objective flare

```{r}
cox_hard_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * diagnosis2 + IMD + Sex + age_decade + FC + Smoke + frailty(SiteNo),
  data = data_survival_hard
)

cox_hard_interaction %>% print_survival()

cox_hard_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * diagnosis2 + IMD + Sex + age_decade + FC + Smoke + frailty(SiteNo),
  data = data_survival_hard,
  m = 20,
  printFlag = FALSE
)

cox_hard_interaction_pool <- with(
  cox_hard_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      OftenLackEnergy * diagnosis2 + IMD + Sex + age_decade + FC + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_hard_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

# Interaction with age

The fatigue models are stratified by diagnosis and include an interaction between fatigue and age decade. `diagnosis2` defines the strata and is not included in these model formulas.

```{r}
data_survival_soft_uc <- data_survival_soft %>% dplyr::filter(diagnosis2 == "UC/IBDU")
data_survival_soft_cd <- data_survival_soft %>% dplyr::filter(diagnosis2 == "CD")
data_survival_hard_uc <- data_survival_hard %>% dplyr::filter(diagnosis2 == "UC/IBDU")
data_survival_hard_cd <- data_survival_hard %>% dplyr::filter(diagnosis2 == "CD")
```

## UC/IBDU — patient-reported flare

```{r}
cox_soft_uc_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo),
  data = data_survival_soft_uc
)

cox_soft_uc_age_interaction %>% print_survival()

cox_soft_uc_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo),
  data = data_survival_soft_uc,
  m = 20,
  printFlag = FALSE
)

cox_soft_uc_age_interaction_pool <- with(
  cox_soft_uc_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_soft_uc_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

## CD — patient-reported flare

```{r}
cox_soft_cd_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo),
  data = data_survival_soft_cd
)

cox_soft_cd_age_interaction %>% print_survival()

cox_soft_cd_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo),
  data = data_survival_soft_cd,
  m = 20,
  printFlag = FALSE
)

cox_soft_cd_age_interaction_pool <- with(
  cox_soft_cd_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_soft_cd_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

## UC/IBDU — objective flare

```{r}
cox_hard_uc_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo),
  data = data_survival_hard_uc
)

cox_hard_uc_age_interaction %>% print_survival()

cox_hard_uc_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo),
  data = data_survival_hard_uc,
  m = 20,
  printFlag = FALSE
)

cox_hard_uc_age_interaction_pool <- with(
  cox_hard_uc_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_hard_uc_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

## CD — objective flare

```{r}
cox_hard_cd_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo),
  data = data_survival_hard_cd
)

cox_hard_cd_age_interaction %>% print_survival()

cox_hard_cd_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo),
  data = data_survival_hard_cd,
  m = 20,
  printFlag = FALSE
)

cox_hard_cd_age_interaction_pool <- with(
  cox_hard_cd_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      OftenLackEnergy * age_decade + IMD + Sex + FC + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_hard_cd_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
