---
title: "PHQ"
author:
  - name: "Chiara Cotronei"
    url: https://www.research.ed.ac.uk/en/persons/chiara-cotronei
    affiliations:
      - ref: Lothian
---


```{r setup}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = TRUE)
```

```{r load libraries, warn.conflicts = FALSE}
set.seed(123)
library(plyr) # Used for mapping values
suppressPackageStartupMessages(library(tidyverse)) # ggplot2, dplyr, and magrittr
library(readxl) # Read in Excel files
library(datefixR) # Standardise dates  
library(openxlsx)
library(tidyr)
library(ggdist)
library(rstatix)

suppressPackageStartupMessages(library(table1))
library(knitr)
library(pander)
library(kableExtra)

# Generate flowchart of cohort derivation
library(DiagrammeR)
library(DiagrammeRsvg)

#cox-regression
library(survival)
library(survminer)
library(finalfit)

```

```{r access to data}
# paths to PREdiCCt data
if (file.exists("/.dockerenv")) {
  # If running in docker
  data.path <- "data/final/20221004/"
  redcap.path <- "data/final/20231030/"
  prefix <- "data/end-of-follow-up/"
  outdir <- "data/processed"
  save_path <- "people/chiara/"
} else {
  # Run on OS directly
  data.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20221004/"
  redcap.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20231030/"
  prefix <- "/Volumes/igmm/cvallejo-predicct/predicct/end-of-follow-up/"
  outdir <- "/Volumes/igmm/cvallejo-predicct/predicct/processed/"
  save_path <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
}
```

```{r load data}
#read-in data
#hads is question 42 of the baseline questionnaire
hads <- read.xlsx(paste0(data.path, "Baseline2022/hads.xlsx"))
health <- read.xlsx(paste0(data.path, "Baseline2022/health.xlsx"))
demographics <- read.xlsx(paste0(data.path, "Baseline2022/demographics2022.xlsx"))
IBD <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx"))
phq <- read.xlsx(paste0(data.path, "Baseline2022/phq.xlsx"))
demo <- readRDS("/Volumes/igmm/cvallejo-predicct/predicct/processed/demo.RDS")
flares <- readRDS("/Volumes/igmm/cvallejo-predicct/predicct/processed/flares/cutoff-flares.RDS")
smoking <- readRDS("/Volumes/igmm/cvallejo-predicct/people/chiara/smoking.rds")
flares_hard <- readRDS("/Volumes/igmm/cvallejo-predicct/people/chiara/flares_hard.RDS")
flares_soft <- readRDS("/Volumes/igmm/cvallejo-predicct/people/chiara/flares_soft.RDS")
```


 ## The Patient Health Questionnaire (PHQ - 15)

Q43.1 of the baseline questionnaire: "During the past 4 weeks how much have you been bothered by any of the following problems?"

13 physical symptoms + feeling tired/trouble sleeping. Scores 0 (not at all) to 2 (bothered a lot). 
Total score is between 0-30. 
Scores of ≥5, ≥10, ≥15 represent mild, moderate and severe levels of somatisation. The reliability and validity of the PHQ-15 are high in clinical and occupational health care settings.

```{r, load and tidy phq data set}
#61 missing ParticipantId (blank rows - exclude these, from 1960 to 1899)
part.missing <- phq %>% 
  filter(is.na(ParticipantId))
phq_clean <- phq %>% 
  filter(!is.na(ParticipantId))
#add age and sex data from demographics
phq_clean <- phq_clean %>% 
  inner_join(demographics %>% select(ParticipantNo, Sex, age, diagnosis), by  = "ParticipantNo")
#code diagnosis
phq_clean$diagnosis2 <- plyr::mapvalues(phq_clean$diagnosis,
  from = seq(1, 4),
  to = c(1, 2, 2, 2)
) %>%
  factor(levels = c("1", "2"), labels = c("CD", "UC/IBDU"))

#exclude <18 yo = 53 participants (1899 -> 1846)
phq_clean <- phq_clean[phq_clean$age > 17,]

#make sex into level
phq_clean$Sex <- factor(phq_clean$Sex, levels = c(1, 2), labels = c("Male", "Female"))

#find out how may columns have missing values (26, excluding menstrual cramp question which )
miss_phq <- phq_clean %>%
  filter(rowSums(is.na(select(., c(7:9, 11:21)))) > 0)
#total participants 1820
phq_clean <- phq_clean %>% 
  filter(!(rowSums(is.na(select(., c(7:9, 11:21)))) > 0))

#detract -1 from all the question colums (C7 to 21) (score 0-2 not 1-3)
if (any(phq_clean$StomachPain == 3)) {
  phq_clean <- phq_clean %>% 
    mutate_at(vars("StomachPain", "BackPain", "PainInArms", "MenstrualCramps", "Headaches", "ChestPain", "Dizziness", "FaintingSpells", "HeartPound", "ShortnessBreath", "SexualIntercourse", "ConstipationDiarrhoea", "NauseaIndigestion", "FeelingTired", "TroubleSleeping"), ~ . - 1)
}

# if value is NA (for menstrual symptoms question)
phq_clean <- phq_clean %>% 
  mutate(MenstrualCramps = ifelse(is.na(MenstrualCramps), 0, MenstrualCramps))

#add all PHQ-15 columns to make final score
phq_clean <- phq_clean %>% 
  mutate(TotalPHQ = StomachPain + BackPain + PainInArms + MenstrualCramps + Headaches + ChestPain + Dizziness + FaintingSpells + HeartPound + ShortnessBreath + SexualIntercourse + ConstipationDiarrhoea + NauseaIndigestion + FeelingTired + TroubleSleeping)

#make new variable for no, mild, moderate and severe levels of somatisation
phq_clean <- phq_clean %>% 
  mutate(
    somatisation = cut(TotalPHQ, breaks = c(0, 4, 9, 14, 30), labels = c("None", "Mild", "Moderate", "Severe"), include.lowest = TRUE)
  )

#split into age groups
phq_clean$AgeGroup <- cut(phq_clean$age, 
                         breaks = c(18, 24, 34, 44, 54, 65, Inf), 
                         labels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+"),
                         include.lowest = TRUE)

#add flare data, FC, smoking
#flare data from IBD data set
phq_clean <- phq_clean %>% 
  left_join(IBD %>% select(ParticipantNo, FlaresInPastYear))
#find missing flare data -> 4 participants, these were excluded
missing_flare <- phq_clean %>%
  filter(is.na(FlaresInPastYear))
#now 1816
phq_clean <- phq_clean %>%
  filter(!is.na(FlaresInPastYear))
#create flare groups
phq_clean <- phq_clean %>%
  mutate(flare_group = ifelse(FlaresInPastYear == 0, "No Flares", "1 or More Flares"))
#make flares into levels
phq_clean$flare_group <- factor(phq_clean$flare_group, levels = c( "No Flares","1 or More Flares"))

phq_clean <- phq_clean %>% 
  left_join(smoking %>% select(ParticipantNo, Smoke))

#204 patient smissing fc data so new dataset for this
phq_clean <- phq_clean %>% 
  left_join(demo %>% select(ParticipantNo, FC, cat, Biologic, IMD))
#tot 1608 participants
phq_clean <- phq_clean %>% 
  filter(!is.na(cat))

save_path <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
saveRDS(phq_clean, paste0(save_path, "phq_clean.RDS"))
```

Here we present somatisation prevalance across the cohort.

```{r, average somatisation score across groups, table for whole cohort PHQ}
my.render.cont <- function(x) {
  with(
    stats.apply.rounding(stats.default(x),
      digits = 3,
      round.integers = FALSE
    ),
    c("", "Median (IQR)" = sprintf("%s (%s - %s)", MEDIAN, Q1, Q3))
  )
}


table1::table1(~ diagnosis2 + Sex + AgeGroup + flare_group + cat| somatisation,
  data = phq_clean,
  render.continuous = my.render.cont,
  title = "Somatization Across Categories (Whole Cohort)"
)

```


### Baseline associations - whole cohort

Somatisation scores are: 

* Significantly higher for patients with CD than UC/IBDU (p-value = 0.043)
* Significantly higher for females (p-value < 2.2e-16).
* Significantly higher for those who reported one or more flares in the previous year (p-value = 2.879e-07)
* There was a signficant difference across somatisation score according to age (p-value = 0.017)
  * The 45-54 age group had the highest percentage of participants in the 'severe somatisation group' (12%)
  * The 18-24 age group had the highest percentage of participants in the 'moderate somatisation group' (27.2%)
  * The 65+ had the highest percentage of participants in the 'mild somatisation group' (50%) (likely that this reflects MSK pain associated with age)
  * The 25-34 had the highet percentage of participants in the 'no somatisation group' (31.7%)    
  * The percentage of participants in each somatisation category per age group are presented in table below
* FC was not significantly associated with somatisation category (p-value = 0.99)

```{r, distribution of scores for sex and disease type}
# Sex
contingency_table <- table(phq_clean$Sex, phq_clean$somatisation)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between sex and presence of somatisation")
} else {
  message("There is no significant association between sex and somatisation")
}

phq_percent <- phq_clean %>%
  group_by(TotalPHQ, Sex) %>%
  summarise(Percentage = n() / nrow(phq_clean) * 100) %>%
  ungroup()

# Plot the line graph
ggplot(phq_percent, aes(x = TotalPHQ, y = Percentage, colour = Sex)) +
  geom_line() +
  labs(title = "Distribution of baseline PSQI Scores",
       x = "PSQI score",
       y = "Percentage of Participants") +
  theme_classic()


# Disease type - stat significant for it being higher in CD (p-value = 0.018)
# need to add CI
contingency_table <- table(phq_clean$diagnosis2, phq_clean$somatisation)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between IBD type and somatization")
} else {
  message("There is no significant association between IDB type and somatization")
}

phq_percent <- phq_clean %>%
  group_by(somatisation, diagnosis2) %>%
  dplyr::summarise(Percentage = n() / nrow(phq_clean) * 100) %>%
  ungroup()

# Plot the bar graph (total counts)
ggplot(phq_percent, aes(x = somatisation, y = Percentage, fill = diagnosis2)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of somatization by diagnosis type",
       x = "Somatization Level",
       y = "Percentage of Participants") +
  theme_minimal()

#flares last year 
contingency_table <- table(phq_clean$flare_group, phq_clean$somatisation)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between somatization scores and flares in last year")
} else {
  message("There is no significant association between somatization scores and flares in last year")
}

#age - p-value = 0.02
contingency_table <- table(phq_clean$AgeGroup, phq_clean$somatisation)
contingency_table
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between somatisation scores and age")
} else {
  message("There is no significant association between somatisation scores and age")
}


percentages <- phq_clean %>%
  group_by(AgeGroup, somatisation) %>%
  summarise(count = n()) %>%
  group_by(AgeGroup) %>%
  mutate(percentage = count / sum(count) * 100)

# Create a line chart to show the percentage of respondents with a certain HADS score
ggplot(percentages, aes(x = AgeGroup, y = percentage, color = somatisation, group = somatisation)) +
  geom_line() +
  labs(title = "Sleep Distrubance (PSQI =>5) across age groups",
       subtitle = "Patients at extremes of age have higher rates of sleep disturbnace, p-value > 0.00001",
       x = "Age Groups",
       y = "Percentage of Paricipants") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme_minimal()

table1::table1(~ somatisation | AgeGroup,
  data = phq_clean,
  render.continuous = my.render.cont,
  caption = "Distribution of patients across somatisation categories (by age group)"
)

#FC - 
contingency_table <- table(phq_clean$cat, phq_clean$somatisation)
contingency_table
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between somatisation scores and FC")
} else {
  message("There is no significant association between somatisation scores and FC")
}
```

## Analysis by diagnosis type separately

Here we investigate whether somatisation scores in CD and UC/IBDU patients have different associations with sex, age, flares in last year and FC

### Baseline Associations - Crohn's Disease

* Females have significantly higher somatisation scores than men (p-value < 2.2e-16)
* Participants who reported one or more flares in previous year had higher somatisation scores (p-value = 1.8e-05)
* There was no significant association with age (p-value = 0.119)
* There was no significant association with faecal calprotectin (p-value = 0.83)

```{r, CROHNS AND SOMATISATION}

CD_phq <- phq_clean[phq_clean$diagnosis2 == "CD",]
#total 811 CD patients
```

```{r, table for somatization, CD only}

my.render.cont <- function(x) {
  with(
    stats.apply.rounding(stats.default(x),
      digits = 3,
      round.integers = FALSE
    ),
    c("", "Median (IQR)" = sprintf("%s (%s - %s)", MEDIAN, Q1, Q3))
  )
}


table1::table1(~ Sex + AgeGroup + flare_group + cat| somatisation,
  data = CD_phq,
  render.continuous = my.render.cont,
  title = "Somatization Across Categories (CD) "
)

```

```{r, associations in somatization, CD only}

# Sex for CD p-value < 2.2e-16
contingency_table <- table(CD_phq$Sex, CD_phq$somatisation)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between sex and presence of somatisation")
} else {
  message("There is no significant association between sex and somatisation")
}


#flares last year 
contingency_table <- table(CD_phq$flare_group, CD_phq$somatisation)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between sex and flares in last year")
} else {
  message("There is no significant association between sex and flares in last year")
}

#age - p-value = 0.02
contingency_table <- table(CD_phq$AgeGroup, CD_phq$somatisation)
contingency_table
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between age and presence of somatisation")
} else {
  message("There is no significant association between age and somatisation")
}


#FC - 
contingency_table <- table(CD_phq$cat, CD_phq$somatisation)
contingency_table
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between FC and somatisation")
} else {
  message("There is no significant association between FC and somatisation")
}
```

## UC/IBDU patients only 

* Significant association with sex, more females have higher somatisation scores (p-value =  3.681e-13)
* Higher scores for those reporting flare previous year (p-value = 0.0002)
* No significant association between age groups and somatic scores (p-value = 0.104)
* No significant association with FC (p-value = 0.62)

```{r, UC/IBDU AND SOMATISATION}

UC_phq <- phq_clean[phq_clean$diagnosis2 == "UC/IBDU",]
#total 904 CD patients
```

```{r, table for somatization UC/IBDU only}

my.render.cont <- function(x) {
  with(
    stats.apply.rounding(stats.default(x),
      digits = 3,
      round.integers = FALSE
    ),
    c("", "Median (IQR)" = sprintf("%s (%s - %s)", MEDIAN, Q1, Q3))
  )
}


table1::table1(~ Sex + AgeGroup + flare_group + cat| somatisation,
  data = UC_phq,
  render.continuous = my.render.cont,
  title = "Somatization Disturbance Across Categories (UC/IBDU) "
)

```

```{r, associations for somatization in UC/IBDU}
# Sex
contingency_table <- table(UC_phq$Sex, UC_phq$somatisation)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between sex and somatisation")
} else {
  message("There is no significant association between sex and somatisation")
}


#flares last year 
contingency_table <- table(UC_phq$flare_group, UC_phq$somatisation)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between sex and flares in last year")
} else {
  message("There is no significant association between sex and flares in last year")
}

#age - p-value = 0.13
contingency_table <- table(UC_phq$AgeGroup, UC_phq$somatisation)
contingency_table
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between age and somatisation level")
} else {
  message("There is no significant association between age and somatisation level")
}


#FC - 
contingency_table <- table(UC_phq$cat, UC_phq$somatisation)
contingency_table
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between FC and somatisation level")
} else {
  message("There is no significant association between FC and somatisation level")
}
```

## Survival analysis 


```{r, code for survival analysis}
cox.summary <- function(fit) {
  cat("Cox model summary:")
  fit %>%
    finalfit::fit2df(condense = FALSE) %>% 
    knitr::kable(col.names = c("Variable",
                               "HR",
                               "Lower 95%",
                               "Upper 95%",
                               "P-value"),
                 digits = 4) %>%
    print()
  cat("Proportional hazards assumption test:")
  cox.zph(fit)$table %>%
    knitr::kable(col.names = c("",
                               "Chi-squared statistic",
                               "DF",
                               "P-value"),
                 digits = 4) %>%
    print()
  return()
}



phq_soft <- phq_clean %>% 
  inner_join(flares_soft %>% select(ParticipantNo, softflare, softflare_time)) %>% 
  rename(DiseaseFlareYN = softflare, time = softflare_time)
phq_hard <- phq_clean %>% 
  inner_join(flares_hard %>% select(ParticipantNo, hardflare, hardflare_time)) %>% 
  rename(DiseaseFlareYN = hardflare, time = hardflare_time)  
  
CD_phq_soft <- CD_phq %>% 
  inner_join(flares_soft %>% select(ParticipantNo, softflare, softflare_time)) %>% 
  rename(DiseaseFlareYN = softflare, time = softflare_time)
CD_phq_hard <- CD_phq %>% 
  inner_join(flares_hard %>% select(ParticipantNo, hardflare, hardflare_time)) %>% 
  rename(DiseaseFlareYN = hardflare, time = hardflare_time)

UC_phq_soft <- UC_phq %>% 
  inner_join(flares_soft %>% select(ParticipantNo, softflare, softflare_time)) %>% 
  rename(DiseaseFlareYN = softflare, time = softflare_time)
UC_phq_hard <- UC_phq %>% 
  inner_join(flares_hard %>% select(ParticipantNo, hardflare, hardflare_time)) %>% 
  rename(DiseaseFlareYN = hardflare, time = hardflare_time)
```


### Whole cohoort

Patient-reported flare

* Significant at all levels in all analyses, here I report controlling for site, sex, smoking, IMD age and FC
  * Increase risk with mild somatisation (p-value = 8.2e-05, 95% CI 1.26-2.03)
  * Increase risk with moderate somatisation (p-value = 3.5e-07 , 95% CI 1.51-2.55)
  * Increase risk with severe somatisation (p-value = 1.4e-08, 95% CI 1.86-3.59)
  

Hard flare
* Mild somatisation significant when controlling for site only (p-value = 0.02, 95% CI 1.06-2.11), not significant for other levels or when controlling for other confounding factors

```{r, whole cohort}
#Soft flares
fit <- survfit(Surv(time, DiseaseFlareYN) ~ somatisation, data = phq_soft)

test_result <- survdiff(Surv(time, DiseaseFlareYN) ~ somatisation, data = phq_soft)
print(test_result)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + frailty(SiteNo), data = phq_soft)

fit.me <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + frailty(SiteNo),
                data = phq_soft)
cox.summary(fit.me)
summary(cox_model)

p1 <- ggsurvplot(fit,
           data = phq_soft,
           conf.int = TRUE,
           ggtheme = theme_minimal(),
           palette = c("#0B1D51", "#1C77C3", "#39A9DB", "#A0ECD0"),
           xlab = "Time from study recrutiment (days)",
           title = "All participants - Time to patient-reported flare - Somatisation (PHQ15)"
           )
p1

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke+ AgeGroup + IMD + frailty(SiteNo), data = phq_soft)
summary(cox_model)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + AgeGroup + IMD + cat + frailty(SiteNo), data = phq_soft)
summary(cox_model)



#Hard flares
fit <- survfit(Surv(time, DiseaseFlareYN) ~ somatisation, data = phq_hard)

test_result <- survdiff(Surv(time, DiseaseFlareYN) ~ somatisation, data = phq_hard)
print(test_result)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + frailty(SiteNo), data = phq_hard)
summary(cox_model)

p1 <- ggsurvplot(fit,
           data = phq_hard,
           conf.int = TRUE,
           ggtheme = theme_minimal(),
           palette = c("#3C0000", "#910707", "#FF2E00", "#FF9900"),
           xlab = "Time from study recrutiment (days)",
           title = "All participants - Time to hard flare - Somatisation (PHQ15)"
           )
p1

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + AgeGroup + IMD + frailty(SiteNo), data = phq_hard)
summary(cox_model)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + AgeGroup + IMD + Smoke + cat + frailty(SiteNo), data = phq_hard)
summary(cox_model)

```


### Survival analysis - Crohn's Disease

Patient-reported flare: 
n= 807, number of events= 273 

* controlling for site only, significant increase in risk at all levels of somatisation
* controlling for site, sex, smoking and IMD,  significant increase in risk at all levels of somatisation
* controlling for FC additionally,  significant increase in risk at all levels of somatisation
  * Mil p-value = 0.0013, 95% CI 1.29-2.86
  * Moderate p-value = 5.4e-07, 95% CI 1.93-5.25
  * Severe p-value = 1.6e-07, 95% CI 2.32 - 6.39

Hard flare: 

* Controlling for site only 
  * Significant increased risk for flare for mild  (p-value = 0.011) and severe (p-value = 0.012) somatisation
  * No significant increase in risk for moderate somatisation (p-value = 0.11)
* Controlling for sex, smoking, site, IMD
  * Significant increased risk for flare for mild  (p-value = 0.029) and severe (p-value = 0.050) somatisation
  * No significant increase in risk for moderate somatisation (p-value = 0.23)
* Controlling for FC additionally
  * Significant increased risk for flare for mild (p-value = 0.044, 95% CI 1.01-3.43) and severe (p-value = 0.046, 95% CI 1.01-5.37) somatisation
  * No significant increase in risk for moderate somatisation (p-value = 0.20, 95% CI 0.78-3.15)
  

```{r, clinical flare for CD patients and somatization}
# p-value 0.04
# testing for somatisation, controlling for frailty only: mild somatisation (p-value mild somatisation 0.015, moderate 0.07, severe 0.006). Those with no somatisation had lower than predicted flares.

#Soft flares
fit <- survfit(Surv(time, DiseaseFlareYN) ~ somatisation, data = CD_phq_soft)

test_result <- survdiff(Surv(time, DiseaseFlareYN) ~ somatisation, data = CD_phq_soft)
print(test_result)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation, data = CD_phq_soft)

fit.me <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation,
                data = CD_phq_soft)
cox.summary(fit.me)
summary(cox_model)

soma.cd <- ggsurvplot(fit,
           data = CD_phq_soft,
           conf.int = TRUE,
           ggtheme = theme_minimal(),
           palette = c("#0B1D51", "#1C77C3", "#39A9DB", "#A0ECD0"),
           xlab = "Time from study recrutiment (days)",
           title = "Crohn's - Time to patient-reported flare - Somatisation (PHQ15)",
           legend.title = "Somatisation",
           legend.labs = c("None", "Mild", "Moderate", "Severe")
           )
soma.cd

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + frailty(SiteNo), data = CD_phq_soft)
cox.summary(cox_model)


cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + IMD + frailty(SiteNo), data = CD_phq_soft)
summary(cox_model)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + IMD + cat + frailty(SiteNo), data = CD_phq_soft)
cox.summary(cox_model)
```

```{r, CD hard flare and somatization}
#Hard flares
fit <- survfit(Surv(time, DiseaseFlareYN) ~ somatisation, data = CD_phq_hard)

test_result <- survdiff(Surv(time, DiseaseFlareYN) ~ somatisation, data = CD_phq_hard)
print(test_result)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation, data = CD_phq_hard)

fit.me <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation,
                data = CD_phq_hard)
cox.summary(fit.me)
summary(cox_model)

p1 <- ggsurvplot(fit,
           data = CD_phq_hard,
           conf.int = TRUE,
           ggtheme = theme_minimal(),
           palette = c("#3C0000", "#910707", "#FF2E00", "#FF9900"),
           xlab = "Time from study recrutiment (days)",
           title = "Crohn's - Time to hard flare - Somatisation (PHQ15)"
           )
p1

fit.me <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + frailty(SiteNo),
                data = CD_phq_hard)
cox.summary(fit.me)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + IMD + frailty(SiteNo), data = CD_phq_hard)
summary(cox_model)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + IMD + cat + frailty(SiteNo), data = CD_phq_hard)
cox.summary(cox_model)

```

## UC/IBDU 


Patient-reported flares:
n= 794, number of events= 312 

* significant controlling for site only at all somatisation levels
* significant controlling for site, sex, smoking, IMD at all levels 
* signifciant when controlling for FC additionally 
  * mild p-value = 0.005, 95% CI 1.13-2.03
  * moderate p-value = 0.02, 95% CI 1.06-2.10
  * severe p-value = 0.003, 95% CI 1.24-3.06
  

Hard flares:   
n= 774, number of events= 119 

* No significant increase in risk in any analysis


```{r, clinical flare for UC/IBDU patients and PHQ}
# p-value 0.09
# testing for somatisation, controlling for frailty only: no significance
#Soft flares
fit <- survfit(Surv(time, DiseaseFlareYN) ~ somatisation, data = UC_phq_soft)

test_result <- survdiff(Surv(time, DiseaseFlareYN) ~ somatisation, data = UC_phq_soft)
print(test_result)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation, data = UC_phq_soft)

fit.me <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation,
                data = UC_phq_soft)
cox.summary(fit.me)
summary(cox_model)

soma.uc <- ggsurvplot(fit,
           data = UC_phq_soft,
           conf.int = TRUE,
           ggtheme = theme_minimal(),
           palette = c("#0B1D51", "#1C77C3", "#39A9DB", "#A0ECD0"),
           xlab = "Time from study recrutiment (days)",
           title = "UC/IBDU - Time to patient-reported flare - Somatisation (PHQ15)",
           legend.title = "Somatisation",
           legend.labs = c("None", "Mild", "Moderate", "Severe")
           )
soma.uc

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + frailty(SiteNo), data = UC_phq_soft)
cox.summary(cox_model)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + IMD + frailty(SiteNo), data = UC_phq_soft)
summary(cox_model)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + IMD + cat + frailty(SiteNo), data = UC_phq_soft)
cox.summary(cox_model)
```

```{r, hard flare, UC, somatization}
#hard flares
fit <- survfit(Surv(time, DiseaseFlareYN) ~ somatisation, data = UC_phq_hard)

test_result <- survdiff(Surv(time, DiseaseFlareYN) ~ somatisation, data = UC_phq_hard)
print(test_result)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation, data = UC_phq_hard)

fit.me <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation,
                data = UC_phq_hard)
cox.summary(fit.me)
summary(cox_model)

p1 <- ggsurvplot(fit,
           data = UC_phq_hard,
           conf.int = TRUE,
           ggtheme = theme_minimal(),
           palette = c("#3C0000", "#910707", "#FF2E00", "#FF9900"),
           xlab = "Time from study recrutiment (days)",
           title = "UC/IBDU - Time to hard flare - Somatisation (PHQ15)"
           )
p1

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + frailty(SiteNo), data = UC_phq_hard)
cox.summary(cox_model)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + IMD + frailty(SiteNo), data = UC_phq_hard)
summary(cox_model)

cox_model <- coxph(Surv(time, DiseaseFlareYN) ~ somatisation + Sex + Smoke + IMD + cat + frailty(SiteNo), data = UC_phq_hard)
cox.summary(cox_model)

```


## Reproduction and reproducibility {.appendix}

<details class = "appendix">
  <summary> Session info </summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>

Licensed by 
<a href="https://creativecommons.org/licenses/by/4.0/">CC BY</a>
 unless otherwise stated.
