---
title: "Quality of Life - SF12"
author:
  - name: "Chiara Cotronei"
    url: https://www.research.ed.ac.uk/en/persons/chiara-cotronei
    affiliations:
      - ref: Lothian
editor: 
  markdown: 
    wrap: 72
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = TRUE)
```

```{r load libraries, warn.conflicts = FALSE}
set.seed(123)
library(plyr) # Used for mapping values
suppressPackageStartupMessages(library(tidyverse)) # ggplot2, dplyr, and magrittr
library(readxl) # Read in Excel files
library(datefixR) # Standardise dates  
library(openxlsx)
library(tidyr)
library(ggdist)
library(rstatix)

suppressPackageStartupMessages(library(table1))
library(knitr)
library(pander)
library(kableExtra)

# Generate flowchart of cohort derivation
library(DiagrammeR)
library(DiagrammeRsvg)

#cox-regression
library(survival)
library(survminer)
library(finalfit)

```

```{r access to data}
# paths to PREdiCCt data
if (file.exists("/.dockerenv")) {
  # If running in docker
  data.path <- "data/final/20221004/"
  redcap.path <- "data/final/20231030/"
  prefix <- "data/end-of-follow-up/"
  outdir <- "data/processed"
  save_path <- "people/chiara/"
} else {
  # Run on OS directly
  data.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20221004/"
  redcap.path <- "/Volumes/igmm/cvallejo-predicct/predicct/final/20231030/"
  prefix <- "/Volumes/igmm/cvallejo-predicct/predicct/end-of-follow-up/"
  outdir <- "/Volumes/igmm/cvallejo-predicct/predicct/processed/"
  save_path <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
}
```

```{r load data}
#read-in data
#hads is question 42 of the baseline questionnaire
lifeevents <- read.xlsx(paste0(data.path, "Baseline2022/lifeevents.xlsx"))
demographics <- read.xlsx(paste0(data.path, "Baseline2022/demographics2022.xlsx"))
demo <- readRDS(paste0(outdir, "demo.RDS"))
IBD <- read.xlsx(paste0(data.path, "Baseline2022/IBD.xlsx"))
QOL <- read.xlsx(paste0(data.path, "Baseline2022/health.xlsx"))
```

# Quality of life

The SF12 provides two scores: the PSC12 and the MCS12 which give a score
(0-100) where higher scores equal better quality of life. The PSC12 is
calculated by domains reflecting physical function, and the MCS12
reflects mental function.

Of note, in the PREdiCCt questionnaire, 4/6 questions pertaining to
mental health were scored out of 5 options instead of 6 as per the SF12.
I have coded to reflect that option 4 ("some of the time") was never
selected, and believe this to have influenced the distribution of MCS12
scores (see distribution graph below). However, this should not affect
comparison between groups within the PREdiCCt cohort.

```{r, clean QOL as per PREdiCCt paper 2 standard}

QOL <- QOL %>% 
  filter(!is.na(ParticipantNo))

#add age and sex data from demographics
QOL <- QOL %>% 
  left_join(
    demographics %>% select(ParticipantNo, Sex, age, diagnosis), 
    by  = "ParticipantNo")

# Recode diagnosis
QOL %<>%
  dplyr::mutate(
    diagnosis2 = dplyr::case_when(
      diagnosis == 1 ~ 1,
      diagnosis == 2 ~ 2,
      diagnosis == 3 ~ 2,
      diagnosis == 4 ~ 2
    )
  ) %>%
  dplyr::mutate(diagnosis2 = factor(
    diagnosis2,
    levels = c("1", "2"),
    labels = c("CD", "UC/IBDU")
  ))

#exclude <18 yo = 53 participants (1907 -> 1854)
QOL %<>% dplyr::filter(age >= 18)

#make sex into level
QOL %<>% dplyr::mutate(Sex = factor(
  Sex,
  levels = c(1, 2),
  labels = c("Male", "Female"))
)

#make age brackets
QOL %<>%
  dplyr::mutate(AgeGroup = cut(age, 
                         breaks = c(18, 24, 34, 44, 54, 65, Inf), 
                         labels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+"),
                         include.lowest = TRUE))

QOL <- QOL %>% 
  left_join(IBD %>% select(ParticipantNo, FlaresInPastYear),
            by = "ParticipantNo")

# 6 missing flare data
#QOL %>% filter(is.na(FlaresInPastYear)) %>% nrow()

QOL <- QOL %>%
  filter(!is.na(FlaresInPastYear))

#create flare groups
QOL <- QOL %>%
  mutate(flare_group = ifelse(FlaresInPastYear == 0,
                              "No Flares", "1 or More Flares"))

#make flares into levels
QOL %<>% 
  dplyr::mutate(
    flare_group = factor(flare_group, levels = c( "No Flares","1 or More Flares"))
  )

QOL <- QOL %>% 
  left_join(demo %>% select(ParticipantNo, FC, cat, Biologic, IMD),
            by = "ParticipantNo")

#209 with missing FC
#QOL %>% filter(is.na(cat)) %>% nrow()

QOL <- QOL %>% 
  filter(!is.na(cat))

# Check missing data - 53 
# QOL %>%
#   filter(
#     dplyr::if_any(.cols = c(7:18), .fns = is.na)
#   ) %>% nrow()

# Exclude missing
QOL %<>%
  filter(
    !dplyr::if_any(.cols = c(7:18), .fns = is.na)
  )
```

Methods

Coded according to SF12 guidance which weighs each item response
differently. Some code used taken from github
https://github.com/lbraglia/lbscorer/blob/master/R/sf12.R.

Join demographics data and demo (FC data) Remove missing Participant No,
\<18y/o, missing FC Check for missing data - exclude if missing answers
(n=53) Check for data out of range (done manaully):

-   GeneralHealth - nil out of range
-   ModerateActivities - nil out of range
-   ClimbingStairs - nil out of range
-   PhysicalAccomplishedLess - nil out of range
-   PhysicalLimitedKindWork - nil out of range
-   EmotionalAccomplishedLess - nil out of range
-   EmotionalNotCarefulAsUsual - nil out of range
-   PainInterfere - nil out of range
-   CalmAndPeaceful - nil out of range
-   LotOfEnergy - nil out of range
-   DownheartedBlue - nil out of range

To account for missing option ("some of the time") recode:

-   CalmAndPeaceful 4=5 and 5=6
-   LotOfEnergy 4=5 and 5=6
-   DownheartedBlue 4=5 and 5=6
-   SocialActivities 4=5 and 5=6

Reverse code: GeneralHealth, PainInterfere, CalmAndPeaceful,
DownheartedBlue

```{r, reverse code as per protocol}

# Recoding answers to be out of 6, leaving option 4 empty.
QOL <- QOL %>%
  dplyr::mutate(
    dplyr::across(
      .cols = c(CalmAndPeaceful, LotOfEnergy, DownheartedBlue, SocialActivities),
      .fn = function(x) dplyr::case_when(
        x == 5 ~ 6,
        x == 4 ~ 5,
        .default = x
      )
    )
  )

#Reverse code GeneralHealth, PainInterfere, CalmAndPeaceful, DownheartedBlue
QOL %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = c(CalmAndPeaceful, DownheartedBlue),
      .fn = ~ 7 - .
    )
  )

QOL %<>%
  dplyr::mutate(
    dplyr::across(
      .cols = c(GeneralHealth, PainInterfere),
      .fn = ~ 6 - .
    )
  )

```

```{r, rename for ease of anaysis}
QOL <- QOL %>% 
  rename(gh1 = GeneralHealth, 
         pf02 = ModerateActivities, 
         pf04 = ClimbingStairs, 
         rp2 = PhysicalAccomplishedLess, 
         rp3 = PhysicalLimitedKindWork, 
         re2 = EmotionalAccomplishedLess, 
         re3 = EmotionalNotCarefulAsUsual, 
         bp2 = PainInterfere, 
         mh3 = CalmAndPeaceful, 
         vt2 = LotOfEnergy, 
         mh4 = DownheartedBlue, 
         sf2 = SocialActivities)
```

```{r, coding of SF12}

# Code from https://github.com/lbraglia/lbscorer/blob/master/R/sf12.R 
# Comments reworded for clarity
  
# Dummy variables for pf02 (3 as reference)
QOL$pf02_1 <- as.numeric(QOL$pf02 == 1L)
QOL$pf02_2 <- as.numeric(QOL$pf02 == 2L)

# Dummy variables for pf02 (3 as reference)
QOL$pf04_1 <- as.numeric(QOL$pf04 == 1L)
QOL$pf04_2 <- as.numeric(QOL$pf04 == 2L)

# Dummy variables for rp2 (2 as reference)
QOL$rp2_1 <- as.numeric(QOL$rp2 == 1L)

# Dummy variables for rp2 (2 as reference)
QOL$rp3_1 <- as.numeric(QOL$rp3 == 1L)

# Dummy variables for bp2 (5 as reference)
QOL$bp2_1 <- as.numeric(QOL$bp2 == 1L)
QOL$bp2_2 <- as.numeric(QOL$bp2 == 2L)
QOL$bp2_3 <- as.numeric(QOL$bp2 == 3L)
QOL$bp2_4 <- as.numeric(QOL$bp2 == 4L)

# Dummy variables for gh1 (5 as reference)
QOL$gh1_1 <- as.numeric(QOL$gh1 == 1L)
QOL$gh1_2 <- as.numeric(QOL$gh1 == 2L)
QOL$gh1_3 <- as.numeric(QOL$gh1 == 3L)
QOL$gh1_4 <- as.numeric(QOL$gh1 == 4L)

# Dummy variables for vt2 (6 as reference)
QOL$vt2_1 <- as.numeric(QOL$vt2 == 1L)
QOL$vt2_2 <- as.numeric(QOL$vt2 == 2L)
QOL$vt2_3 <- as.numeric(QOL$vt2 == 3L)
QOL$vt2_4 <- as.numeric(QOL$vt2 == 4L)
QOL$vt2_5 <- as.numeric(QOL$vt2 == 5L)

# Dummy variables for sf2 (5 as reference)
QOL$sf2_1 <- as.numeric(QOL$sf2 == 1L)
QOL$sf2_2 <- as.numeric(QOL$sf2 == 2L)
QOL$sf2_3 <- as.numeric(QOL$sf2 == 3L)
QOL$sf2_4 <- as.numeric(QOL$sf2 == 4L)

# Dummy variables for re2 (2 as reference)
QOL$re2_1 <- as.numeric(QOL$re2 == 1L)

# Dummy variables for re3 (2 as reference)
QOL$re3_1 <- as.numeric(QOL$re3 == 1L)

# Dummy variables for mh3 (6 as reference)
QOL$mh3_1 <- as.numeric(QOL$mh3 == 1L)
QOL$mh3_2 <- as.numeric(QOL$mh3 == 2L)
QOL$mh3_3 <- as.numeric(QOL$mh3 == 3L)
QOL$mh3_4 <- as.numeric(QOL$mh3 == 4L)
QOL$mh3_5 <- as.numeric(QOL$mh3 == 5L)

# Dummy variables for mh4 (6 as reference)
QOL$mh4_1 <- as.numeric(QOL$mh4 == 1L)
QOL$mh4_2 <- as.numeric(QOL$mh4 == 2L)
QOL$mh4_3 <- as.numeric(QOL$mh4 == 3L)
QOL$mh4_4 <- as.numeric(QOL$mh4 == 4L)
QOL$mh4_5 <- as.numeric(QOL$mh4 == 5L)


# Calculate scores using regression weights from the GitHub

QOL <- QOL %>% 
  mutate(RAWPCS12 = (-7.23216*pf02_1) + (-3.45555*pf02_2) +
                    (-6.24397*pf04_1) + (-2.73557*pf04_2) +
                    (-4.61617*rp2_1) + 
                    (-5.51747*rp3_1) +
                    (-11.25544*bp2_1) + (-8.38063*bp2_2) +
                    (-6.50522*bp2_3) + (-3.80130*bp2_4) + (-8.37399*gh1_1) +
                    (-5.56461*gh1_2) + (-3.02396*gh1_3) + (-1.31872*gh1_4) +
                    (-2.44706*vt2_1) + (-2.02168*vt2_2) + (-1.6185*vt2_3) +
                    (-1.14387*vt2_4) + (-0.42251*vt2_5) + (-0.33682*sf2_1) +
                    (-0.94342*sf2_2) + (-0.18043*sf2_3) + (0.11038*sf2_4) +
                    (3.04365*re2_1) + (2.32091*re3_1) + (3.46638*mh3_1) +
                    (2.90426*mh3_2) + (2.37241*mh3_3) + (1.36689*mh3_4) +
                    (0.66514*mh3_5) + (4.61446*mh4_1) + (3.41593*mh4_2) +
                    (2.34247*mh4_3) + (1.28044*mh4_4) + (0.41188*mh4_5))

QOL <- QOL %>% 
  mutate(RAWMCS12 = (3.93115*pf02_1) + (1.8684*pf02_2) +
                   (2.68282*pf04_1) + (1.43103*pf04_2) + (1.4406*rp2_1) +
                   (1.66968*rp3_1) + (1.48619*bp2_1) + (1.76691*bp2_2) +
                   (1.49384*bp2_3) + (0.90384*bp2_4) + (-1.71175*gh1_1) +
                   (-0.16891*gh1_2) + (0.03482*gh1_3) + (-0.06064*gh1_4) +
                   (-6.02409*vt2_1) + (-4.88962*vt2_2) + (-3.29805*vt2_3) +
                   (-1.65178*vt2_4) + (-0.92057*vt2_5) + (-6.29724*sf2_1) +
                   (-8.26066*sf2_2) + (-5.63286*sf2_3) + (-3.13896*sf2_4) +
                   (-6.82672*re2_1) + (-5.69921*re3_1) + (-10.19085*mh3_1) +
                   (-7.92717*mh3_2) + (-6.31121*mh3_3) + (-4.09842*mh3_4) +
                   (-1.94949*mh3_5) + (-16.15395*mh4_1) + (-10.77911*mh4_2) +
                   (-8.09914*mh4_3) + (-4.59055*mh4_4) + (-1.95934*mh4_5))


# Standardisation of scores - from GitHub
# Physical component score
QOL <- QOL %>% 
  mutate(PSC12 = RAWPCS12 + 56.57706)

# Mental component score
QOL <- QOL %>% 
  mutate(MCS12 = RAWMCS12 + 60.75781)

# How many missing score? 0
#QOL %>% filter(is.na(PSC12)) %>% nrow()
  
```

## Investigating distribution of scores.

Physical score (PSC12) and mental health scores (MCS12) were not
normally distributed therefore differences across groups were tested
using Mann-Whitney U test (comparing 2 groups: diagnosis type, sex,
flares last year) or Kruksal-Wallis (comparing \>2 groups: age groups,
faecal calprotectin category).

Note: For PSC12 and MCS12, a higher score indicates better health.

```{r, investigating distribution of scores}
plot(density(QOL$PSC12))

# Test of non-normality (significant = non normal)
shapiro.test(QOL$PSC12)
# Not normally distributed therefore cannot use t-test to compare

ggplot(QOL, aes(x = PSC12, fill = Sex)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of PSC12 by Sex",
       x = "PSC12",
       y = "Density",
       fill = "Sex") +
  theme_minimal()

plot(density(QOL$MCS12))
shapiro.test(QOL$MCS12)

ggplot(QOL, aes(x = MCS12, fill = Sex)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of MCS12 by Sex",
       x = "MCS12",
       y = "Density",
       fill = "Sex") +
  theme_minimal()
#MCS12 scores are not normally distributed therefore cannot use t-test to compare
#interesting dip in the centre - could be due to not having option 5 instead of 6 options in the test

```

## Investigating associations with baseline SF12 scores (continuous)

Here we present the summary findings of tests for association. The
results can then be seen in the sections and code below.

Summary:

Physical scores (PSC12)

-   No statistical significance by diagnosis type
-   Statistically significant difference in PSC12 scores by sex (p-value
    = 3.5e-05)
-   There is a statistically significant difference between age groups
    in PSC12 score, p-value = 4.06e-13
-   No statistically significant difference by FC category in PSC12
    scores, p-value = 0.1953
-   There is a statistically significant difference between flares group
    (reported in the previous year) in PSC12 score, p-value = 4.6e-08

Mental health scores (MCS12)

-   No statistical significance by diagnosis type
-   No statistically significant difference in MSC12 scores (p-value =
    0.17), could this be due to missing option 4 - these are specific to
    mental health domains?
-   There is a statistially significant difference between age groups in
    MCS12 score, p-value = 0.0001116
-   No statistically significant difference by FC category in PSC12
    scores, p-value = 0.5556
-   No statistically significant difference between flares group
    (reported in the previous year) in MCS12 score, p-value = 0.34

Interestingly, there were no associations with sex or flare group for
mental health domains, which is out of keeping with other data in
PREdiCCt pertaining to mental health.

These tests were re-run using chi-square test and comparing differences
between groups using cut-offs of 50 for PSC12 and 42 for MCS12 -
associations with key variables remained unchanged.

### Diagnosis type

Physical scores (PSC12) : No statistical significance by diagnosis type

Mental health scores (MCS12) : No statistical significance by diagnosis
type

```{r, create table showing SF12 and MCS12 scores, diagnosis2}
my.render.cont <- function(x) {
  with(
    stats.apply.rounding(stats.default(x),
      digits = 3,
      round.integers = FALSE
    ),
    c("", "Median (IQR)" = sprintf("%s (%s - %s)", MEDIAN, Q1, Q3),
      "Mean" = ifelse(is.numeric(MEAN), sprintf("%.3f", MEAN), MEAN),
      "Standard Deviation" = ifelse(is.numeric(SD), sprintf("%.3f", SD), SD),
      "Range" = sprintf("[%s - %s]", MIN, MAX))
  )
}


table1::table1(~ PSC12 | diagnosis2,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of PSC12 by Diagnosis Type"
)

table1::table1(~ MCS12 | diagnosis2,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of MCS12 by Diagnosis Type"
)

#Mann-Whitney-U PSC12
QOL %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = diagnosis2)
  )

pairwise_tests <- pairwise.wilcox.test(QOL$PSC12, QOL$diagnosis2, p.adjust.method = "bonferroni")
print(pairwise_tests)

#Mann-Whitney-U MCS12
QOL %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = diagnosis2)
  )

pairwise_tests <- pairwise.wilcox.test(QOL$MCS12, QOL$diagnosis2, p.adjust.method = "bonferroni")
print(pairwise_tests)


```

### Sex

Statistically significant difference in PSC12 scores by sex (p-value =
3.5e-05) No statistically significant difference in MSC12 scores
(p-value = 0.17), could this be due to missing option 4 - these are
specific to mental health domains?

```{r, PSC12 and MCS12 sex}
table1::table1(~ PSC12 | Sex,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of PSC12 by sex"
)


#mental health

table1::table1(~ MCS12 | Sex,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of MCS12 by sex"
)

#Mann-Whitney-U
QOL %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = Sex)
  )

pairwise_tests <- pairwise.wilcox.test(QOL$PSC12, QOL$Sex)
print(pairwise_tests)

#mental health
#Mann-Whitney-U
QOL %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = Sex)
  )

pairwise_tests <- pairwise.wilcox.test(QOL$MCS12, QOL$Sex)
print(pairwise_tests)

```

###Age Group

There is a statistically significant difference between age groups in
PSC12 score, p-value = 4.06e-13 There is a statistially significant
difference between age groups in MCS12 score, p-value = 0.0001116

```{r, PSC12 and MCS12 AgeGroup}
table1::table1(~ PSC12 | AgeGroup,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of PSC12 by Age Group"
)

# Violin plot
QOL %>%
  ggplot() +
  geom_violin(aes(x = AgeGroup, y = PSC12))

#mental health

table1::table1(~ MCS12 | AgeGroup,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of MCS12 by Age Group"
)

# Violin plot
QOL %>%
  ggplot() +
  geom_density(aes(x = MCS12, colour = AgeGroup))

#Kruksal Wallis test
kruskal.test(PSC12 ~ AgeGroup, data = QOL)

#mental health
#Kruksal Wallis test
kruskal.test(MCS12 ~ AgeGroup, data = QOL)

```

Flares (in last year)

There is a statistically significant difference between flares group
(reported in the previous year) in PSC12 score, p-value = 4.6e-08 There
is no statistically significant difference between flares group
(reported in the previous year) in MCS12 score, p-value = 0.34

```{r, PSC12 and MCS12 flares}
table1::table1(~ PSC12 | flare_group,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of PSC12 by flares in previous year"
)


table1::table1(~ MCS12 | flare_group,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of MCS12 by flares in previous year"
)

QOL %>% 
  ggplot() +
  geom_density(
    aes(x = PSC12, colour = flare_group)
  )

QOL %>% 
  ggplot() +
  geom_density(
    aes(x = MCS12, colour = flare_group)
  )

#Mann-Whitney-U
pairwise_tests <- pairwise.wilcox.test(QOL$PSC12, QOL$flare_group, p.adjust.method = "bonferroni")
print(pairwise_tests)

#mental health
#Mann-Whitney-U
pairwise_tests <- pairwise.wilcox.test(QOL$MCS12, QOL$flare_group, p.adjust.method = "bonferroni")
print(pairwise_tests)

```

Faecal Calprotectin

No statistically significant difference by FC category in PSC12 scores,
p-value = 0.1953 No statistically significant difference by FC category
in PSC12 scores, p-value = 0.5556

```{r, PSC12 and MCS12 FC cat}
table1::table1(~ PSC12 | cat,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of PSC12 by FC category"
)

table1::table1(~ MCS12 | cat,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Summary Statistics of MCS12 by FC category"
)

QOL %>%
  ggplot() +
  geom_violin(aes(x = cat, y = PSC12))

QOL %>%
  ggplot() +
  geom_violin(aes(x = cat, y = MCS12))


#Kruksal Wallis test
kruskal.test(PSC12 ~ cat, data = QOL)

#mental health
#Kruksal Wallis test
kruskal.test(MCS12 ~ cat, data = QOL)
```

### Investigating association with grouped SF12 scores

There are limitations in comparing mental health scores given the lack
of option "some of the time" in the baseline questionnaire. Suggested
cut off of 50 for PSC12 to determine a physical condition; while a score
of 42 or less on the MCS-12 may be indicative of ‘clinical depression’.

Analyses were re-run using chi-squared to compare grouped SF12 scores -
there was no change to associations with key variables.

```{r, cut offs an re-analyse for associations}
QOL <- QOL %>% 
  mutate(PSC_grouped = ifelse(PSC12 >= 50, "50+", "<50"))

QOL <- QOL %>% 
  mutate(MCS_grouped = ifelse(MCS12 >= 42, "42+", "<42"))
```

Tests for associations PSC12 with cut off \<50 (chi-squared)

-   diagnosis type not significantly associated
-   sex significant association (p-value = 4.631e-06)
-   age group significant (p-value = 9.029e-09)
-   flare group significant (p-value = 0.0002242)
-   fc category not significantly associated

Test for associations MCS12 with cut off \<42 (chi-squared)

-   diagnosis type not significantly associated
-   sex not significantly associated
-   age group significant (p-value = 0.03874)
-   flare group not significantly associated
-   fc category not significantly associated

```{r, PSC12 and MCS12 grouped tables}
table1::table1(~ diagnosis2 + Sex + AgeGroup + flare_group + cat  | PSC_grouped,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Distribution by variable into PSC12 score groupe"
)

table1::table1(~ diagnosis2 + Sex + AgeGroup + flare_group + cat  | MCS_grouped,
  data = QOL,
  render.continuous = my.render.cont,
  caption = "Distribution by variable into MCS12 score groupe"
)

```

```{r, PSC12 and grouped associations}
contingency_table <- table(QOL$diagnosis2, QOL$PSC_grouped)
print(contingency_table)

#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between diagnosis type and PSC12 scores (cut off of 50)")
} else {
  message("There is no significant association between diagnosis type and PSC12 scores (cut off of 50)")
}

contingency_table <- table(QOL$Sex, QOL$PSC_grouped)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between sex and PSC12 scores (cut off of 50)")
} else {
  message("There is no significant association between sex and PSC12 scores (cut off of 50)")
}


contingency_table <- table(QOL$AgeGroup, QOL$PSC_grouped)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between age group and PSC12 scores (cut off of 50)")
} else {
  message("There is no significant association between age group and PSC12 scores (cut off of 50)")
}

contingency_table <- table(QOL$flare_group, QOL$PSC_grouped)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between flare group and PSC12 scores (cut off of 50)")
} else {
  message("There is no significant association between flare group and PSC12 scores (cut off of 50)")
}


contingency_table <- table(QOL$cat, QOL$PSC_grouped)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between FC categories and PSC12 scores (cut off of 50)")
} else {
  message("There is no significant association between FC categories and PSC12 scores (cut off of 50)")
}

```

```{r, MCS12 grouped associations}

contingency_table <- table(QOL$diagnosis2, QOL$MCS_grouped)
print(contingency_table)

#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between diagnosis type and MCS12 scores (cut off of 50)")
} else {
  message("There is no significant association between diagnosis type and MCS12 scores (cut off of 50)")
}

contingency_table <- table(QOL$Sex, QOL$MCS_grouped)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between sex and MCS12 scores (cut off of 50)")
} else {
  message("There is no significant association between sex and MCS12 scores (cut off of 50)")
}


contingency_table <- table(QOL$AgeGroup, QOL$MCS_grouped)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between age group and MCS12 scores (cut off of 50)")
} else {
  message("There is no significant association between age group and MCS12 scores (cut off of 50)")
}

contingency_table <- table(QOL$flare_group, QOL$MCS_grouped)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between flare group and MCS12 scores (cut off of 50)")
} else {
  message("There is no significant association between flare group and MCS12 scores (cut off of 50)")
}


contingency_table <- table(QOL$cat, QOL$MCS_grouped)
print(contingency_table)
#chi.sq
chi_squared_result <- chisq.test(contingency_table)
print(chi_squared_result)
# Check if the result is significant at a certain alpha level (e.g., 0.05)
alpha <- 0.05
if (chi_squared_result$p.value < alpha) {
  message("There is a significant association between FC categories and MCS12 scores (cut off of 50)")
} else {
  message("There is no significant association between FC categories and MCS12 scores (cut off of 50)")
}

```

## Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{R Session info}
#| echo: false
pander::pander(sessionInfo())
```

</details>

Licensed by <a href="https://creativecommons.org/licenses/by/4.0/">CC
BY</a> unless otherwise stated.
