---
title: "Diet"
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-expand: true
---

# Setting up in R

## Load the data

```{r}
library(splines)
library(patchwork)

source("Survival/utils.R")

# Setup analysis environment
analysis_setup <- setup_analysis()
paths <- analysis_setup$paths
demo <- analysis_setup$demo

flare.df <- readRDS(paste0(paths$outdir, "flares-biochem.RDS"))
flare.cd.df <- readRDS(paste0(paths$outdir, "flares-biochem-cd.RDS"))
flare.uc.df <- readRDS(paste0(paths$outdir, "flares-biochem-uc.RDS"))
```

## Data cleaning

```{r}
# FC has been logged twice - reverse
flare.df %<>%
  dplyr::mutate(FC = exp(FC))

flare.cd.df %<>%
  dplyr::mutate(FC = exp(FC))

flare.uc.df %<>%
  dplyr::mutate(FC = exp(FC))
# So now FC is the log of the FC measurement

# Transform age variable to decades
flare.df %<>%
  dplyr::mutate(
    age_decade = Age/10
  )

flare.cd.df %<>%
  dplyr::mutate(
    age_decade = Age/10
  )

flare.uc.df %<>%
  dplyr::mutate(
    age_decade = Age/10
  )

# Categorised continuous variables

# Categorize meat protein by quantiles
flare.cd.df <- categorize_by_quantiles(flare.cd.df, "Meat_sum", reference_data = flare.df)
flare.uc.df <- categorize_by_quantiles(flare.uc.df, "Meat_sum", reference_data = flare.df)


# Categorize dietary fibre by quantiles
flare.cd.df <- categorize_by_quantiles(flare.cd.df, "fibre", reference_data = flare.df)
flare.uc.df <- categorize_by_quantiles(flare.uc.df, "fibre", reference_data = flare.df)


# Categorize PUFA by quantiles
flare.cd.df <- categorize_by_quantiles(flare.cd.df, "PUFA_percEng", reference_data = flare.df)
flare.uc.df <- categorize_by_quantiles(flare.uc.df, "PUFA_percEng", reference_data = flare.df)


# Categorize UPF percentage by quantiles
flare.cd.df <- categorize_by_quantiles(flare.cd.df, "UPF_perc", reference_data = flare.df)
flare.uc.df <- categorize_by_quantiles(flare.uc.df, "UPF_perc", reference_data = flare.df)

# Smoking
# Reorder Smoke factor levels
flare.cd.df %<>%
  dplyr::mutate(
    Smoke = forcats::fct_relevel(Smoke, "Never", "Previous", "Current")
  )

flare.uc.df %<>%
  dplyr::mutate(
    Smoke = forcats::fct_relevel(Smoke, "Never", "Previous", "Current")
  )

# Missingness in smoking
# CD
flare.cd.df %>%
  dplyr::pull(Smoke) %>%
  forcats::fct_count(prop = TRUE)
# 28.5% missing

# UC
flare.uc.df %>%
  dplyr::pull(Smoke) %>%
  forcats::fct_count(prop = TRUE)
# 25.7%

```

## Helper functions

```{r}

summon_reference_values <- function(data, variables){
  # Function that extracts the reference value for each variable in a data frame
  # Reference is either the first factor level or the median
  # Used for producing HR plots
  
  data %>%
    dplyr::select(tidyselect::all_of(variables)) %>%
    dplyr::summarise(across(
      .cols = everything(),
      .fns = function(x) {
        if (is.factor(x)) {
          levels(x) %>% dplyr::first()
        } else if (is.numeric(x)) {
          median(x, na.rm = TRUE)
        } else {
          NULL
        }
      }
    ))
  
}


plot_continuous_hr <- function(data, model, variable){
  # Plotting continuous Hazard Ratio curves
  
  # model is a Cox model
  # variable is the variable we are plotting
  
  # Range of the variable to plot
  variable_range <- data %>%
    dplyr::pull(variable) %>%
    quantile(., probs = seq(0, 0.9, 0.01), na.rm = TRUE) %>%
    unname() %>%
    unique()
  
  # Reference variables
  # Extract variables used in the Cox model
  # Remove the variable we want to vary and the SiteNo
  ref_variables <- all.vars(delete.response(terms(model))) %>%
    tibble(x = .) %>%
    dplyr::filter(
      x != variable,
      x != 'SiteNo'
    ) %>%
    dplyr::pull(x)
  
  # Create dummy data for plotting
  # Reference levels
  data %>%
    summon_reference_values(
      data = .,
      variables = ref_variables
    ) %>% {
      # Creating dummy data
      tibble(
        .,
        !!sym(variable) := variable_range
      )
    } %>% {
      # Risk calculate HR relative to the default chosen by cox
      pred = predict(model, newdata = ., type = "risk", se = TRUE)
      
      tibble(., p = pred$fit, se = pred$se)
    } %>%
    ggplot(aes(x = !!rlang::sym(variable), y = p, ymin = p - 1.96*se, ymax = p + 1.96*se)) +
    geom_point() +
    geom_line() +
    geom_ribbon(alpha = 0.2) +
    ylab("HR")
  
}

# Function to perform an LRT
summon_lrt <- function(model, remove = NULL, add = NULL) {
  # Function that removes a term and performs a
  # Likelihood ratio test to determine its significance
  
  new_formula = "~ ."

  if (!is.null(remove)){
    new_formula = paste0(new_formula, " - ", remove)
  } else {remove = NA}
  
  if (!is.null(add)){
    new_formula = paste0(new_formula, " + ", add)
  } else {add = NA}
  
  update(model, new_formula) %>%
    anova(model, ., test = 'LRT') %>%
    broom::tidy() %>%
    dplyr::filter(!is.na(p.value)) %>%
    dplyr::select(-term) %>%
    dplyr::mutate(
      removed = remove,
      added = add) %>%
    dplyr::mutate(test = 'LRT') %>%
    dplyr::select(test, removed, added, tidyselect::everything())
  
}

```

## Total meat protein

### Crohns' disease

#### Patient reported flare

```{r}
# Only select relevant columns for the imputation model
# Only imputing smoking so remove missing others
# Calculate cumulative hazard

# Soft flare
data_impute_meat_cd_soft <- flare.cd.df %>%
  dplyr::select(softflare_time,
                softflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                Meat_sum_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
    data = .,
    timevar = softflare_time,
    statusvar = softflare
  ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_meat_cd_soft)

pred_matrix[, 'softflare_time'] <- 0

# MICE with 10 imputations
mice_meat_cd_soft <- mice::mice(
  data = data_impute_meat_cd_soft,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_meat_cd_soft %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_meat_cd_soft,
  coxph(
    Surv(softflare_time, softflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      Meat_sum_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```



#### Objective flare

```{r}
# Hard flare
data_impute_meat_cd_hard <- flare.cd.df %>%
  dplyr::select(hardflare_time,
                hardflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                Meat_sum_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
    data = .,
    timevar = hardflare_time,
    statusvar = hardflare
  ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_meat_cd_hard)

pred_matrix[, 'hardflare_time'] <- 0

# MICE with 10 imputations
mice_meat_cd_hard <- mice::mice(
  data = data_impute_meat_cd_hard,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_meat_cd_hard %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_meat_cd_hard,
  coxph(
    Surv(hardflare_time, hardflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      Meat_sum_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```


### Ulcerative colitis

#### Patient reported flare

```{r}
# Only select relevant columns for the imputation model
# Only imputing smoking so remove missing others
# Calculate cumulative hazard

# Soft flare
data_impute_meat_uc_soft <- flare.uc.df %>%
  dplyr::select(softflare_time,
                softflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                Meat_sum_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = softflare_time,
      statusvar = softflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_meat_uc_soft)

pred_matrix[, 'softflare_time'] <- 0

# MICE with 10 imputations
mice_meat_uc_soft <- mice::mice(
  data = data_impute_meat_uc_soft,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73, 
  print = FALSE
)

# Check convergence
mice_meat_uc_soft %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_meat_uc_soft,
  coxph(
    Surv(softflare_time, softflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      Meat_sum_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```



#### Objective flare

```{r}
# Hard flare
data_impute_meat_uc_hard <- flare.uc.df %>%
  dplyr::select(hardflare_time,
                hardflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                Meat_sum_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = hardflare_time,
      statusvar = hardflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_meat_uc_hard)

pred_matrix[, 'hardflare_time'] <- 0

# MICE with 10 imputations
mice_meat_uc_hard <- mice::mice(
  data = data_impute_meat_uc_hard,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73, 
  print = FALSE
)

# Check convergence
mice_meat_uc_hard %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_meat_uc_hard,
  coxph(
    Surv(hardflare_time, hardflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      Meat_sum_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```


## Dietary Fibre

### Crohn's disease

#### Patient reported flare

```{r}
# Only select relevant columns for the imputation model
# Only imputing smoking so remove missing others
# Calculate cumulative hazard

# Soft flare
data_impute_fibre_cd_soft <- flare.cd.df %>%
  dplyr::select(softflare_time,
                softflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                fibre_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = softflare_time,
      statusvar = softflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_fibre_cd_soft)

pred_matrix[, 'softflare_time'] <- 0

# MICE with 10 imputations
mice_fibre_cd_soft <- mice::mice(
  data = data_impute_fibre_cd_soft,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_fibre_cd_soft %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_fibre_cd_soft,
  coxph(
    Surv(softflare_time, softflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      fibre_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```



#### Objective flare

```{r}
# Hard flare
data_impute_fibre_cd_hard <- flare.cd.df %>%
  dplyr::select(hardflare_time,
                hardflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                fibre_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = hardflare_time,
      statusvar = hardflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_fibre_cd_hard)

pred_matrix[, 'hardflare_time'] <- 0

# MICE with 10 imputations
mice_fibre_cd_hard <- mice::mice(
  data = data_impute_fibre_cd_hard,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_fibre_cd_hard %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_fibre_cd_hard,
  coxph(
    Surv(hardflare_time, hardflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      fibre_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```


### Ulcerative colitis

#### Patient reported flare

```{r}
# Only select relevant columns for the imputation model
# Only imputing smoking so remove missing others
# Calculate cumulative hazard

# Soft flare
data_impute_fibre_uc_soft <- flare.uc.df %>%
  dplyr::select(softflare_time,
                softflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                fibre_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = softflare_time,
      statusvar = softflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_fibre_uc_soft)

pred_matrix[, 'softflare_time'] <- 0

# MICE with 10 imputations
mice_fibre_uc_soft <- mice::mice(
  data = data_impute_fibre_uc_soft,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_fibre_uc_soft %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_fibre_uc_soft,
  coxph(
    Surv(softflare_time, softflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      fibre_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```



#### Objective flare

```{r}
# Hard flare
data_impute_fibre_uc_hard <- flare.uc.df %>%
  dplyr::select(hardflare_time,
                hardflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                fibre_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = hardflare_time,
      statusvar = hardflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_fibre_uc_hard)

pred_matrix[, 'hardflare_time'] <- 0

# MICE with 10 imputations
mice_fibre_uc_hard <- mice::mice(
  data = data_impute_fibre_uc_hard,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_fibre_uc_hard %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_fibre_uc_hard,
  coxph(
    Surv(hardflare_time, hardflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      fibre_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```


## Polyunsaturated fatty acids

### Crohn's disease

#### Patient reported flare

```{r}
# Only select relevant columns for the imputation model
# Only imputing smoking so remove missing others
# Calculate cumulative hazard

# Soft flare
data_impute_pufa_cd_soft <- flare.cd.df %>%
  dplyr::select(softflare_time,
                softflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                PUFA_percEng_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = softflare_time,
      statusvar = softflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_pufa_cd_soft)

pred_matrix[, 'softflare_time'] <- 0

# MICE with 10 imputations
mice_pufa_cd_soft <- mice::mice(
  data = data_impute_pufa_cd_soft,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_pufa_cd_soft %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_pufa_cd_soft,
  coxph(
    Surv(softflare_time, softflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      PUFA_percEng_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```



#### Objective flare

```{r}
# Hard flare
data_impute_pufa_cd_hard <- flare.cd.df %>%
  dplyr::select(hardflare_time,
                hardflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                PUFA_percEng_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = hardflare_time,
      statusvar = hardflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_pufa_cd_hard)

pred_matrix[, 'hardflare_time'] <- 0

# MICE with 10 imputations
mice_pufa_cd_hard <- mice::mice(
  data = data_impute_pufa_cd_hard,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_pufa_cd_hard %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_pufa_cd_hard,
  coxph(
    Surv(hardflare_time, hardflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      PUFA_percEng_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```


### Ulcerative colitis

#### Patient reported flare

```{r}
# Only select relevant columns for the imputation model
# Only imputing smoking so remove missing others
# Calculate cumulative hazard

# Soft flare
data_impute_pufa_uc_soft <- flare.uc.df %>%
  dplyr::select(softflare_time,
                softflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                PUFA_percEng_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = softflare_time,
      statusvar = softflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_pufa_uc_soft)

pred_matrix[, 'softflare_time'] <- 0

# MICE with 10 imputations
mice_pufa_uc_soft <- mice::mice(
  data = data_impute_pufa_uc_soft,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_pufa_uc_soft %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_pufa_uc_soft,
  coxph(
    Surv(softflare_time, softflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      PUFA_percEng_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```



#### Objective flare

```{r}
# Hard flare
data_impute_pufa_uc_hard <- flare.uc.df %>%
  dplyr::select(hardflare_time,
                hardflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                PUFA_percEng_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = hardflare_time,
      statusvar = hardflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_pufa_uc_hard)

pred_matrix[, 'hardflare_time'] <- 0

# MICE with 10 imputations
mice_pufa_uc_hard <- mice::mice(
  data = data_impute_pufa_uc_hard,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_pufa_uc_hard %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_pufa_uc_hard,
  coxph(
    Surv(hardflare_time, hardflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      PUFA_percEng_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```


## Ultra-processed food

### Crohn's disease

#### Patient reported flare

```{r}
# Only select relevant columns for the imputation model
# Only imputing smoking so remove missing others
# Calculate cumulative hazard

# Soft flare
data_impute_upf_cd_soft <- flare.cd.df %>%
  dplyr::select(softflare_time,
                softflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                UPF_perc_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = softflare_time,
      statusvar = softflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_upf_cd_soft)

pred_matrix[, 'softflare_time'] <- 0

# MICE with 10 imputations
mice_upf_cd_soft <- mice::mice(
  data = data_impute_upf_cd_soft,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_upf_cd_soft %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_upf_cd_soft,
  coxph(
    Surv(softflare_time, softflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      UPF_perc_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```



#### Objective flare

```{r}
# Hard flare
data_impute_upf_cd_hard <- flare.cd.df %>%
  dplyr::select(hardflare_time,
                hardflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                UPF_perc_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = hardflare_time,
      statusvar = hardflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_upf_cd_hard)

pred_matrix[, 'hardflare_time'] <- 0

# MICE with 10 imputations
mice_upf_cd_hard <- mice::mice(
  data = data_impute_upf_cd_hard,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_upf_cd_hard %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_upf_cd_hard,
  coxph(
    Surv(hardflare_time, hardflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      UPF_perc_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```


### Ulcerative colitis

#### Patient reported flare

```{r}
# Only select relevant columns for the imputation model
# Only imputing smoking so remove missing others
# Calculate cumulative hazard

# Soft flare
data_impute_upf_uc_soft <- flare.uc.df %>%
  dplyr::select(softflare_time,
                softflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                UPF_perc_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = softflare_time,
      statusvar = softflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_upf_uc_soft)

pred_matrix[, 'softflare_time'] <- 0

# MICE with 10 imputations
mice_upf_uc_soft <- mice::mice(
  data = data_impute_upf_uc_soft,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_upf_uc_soft %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_upf_uc_soft,
  coxph(
    Surv(softflare_time, softflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      UPF_perc_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```



#### Objective flare

```{r}
# Hard flare
data_impute_upf_uc_hard <- flare.uc.df %>%
  dplyr::select(hardflare_time,
                hardflare,
                Sex,
                cat,
                IMD,
                dqi_tot,
                UPF_perc_cat,
                Smoke,
                SiteNo) %>%
  # Remove missing on all columns except smoke
  dplyr::filter(!dplyr::if_any(
    .cols = -Smoke,
    .fns = is.na
  )) %>%
  # Calculate Cumulative hazard
  dplyr::mutate(
    cumhaz = mice::nelsonaalen(
      data = .,
      timevar = hardflare_time,
      statusvar = hardflare
    ))

# Predictor matrix - need to exclude time from the model
pred_matrix <- mice::make.predictorMatrix(data_impute_upf_uc_hard)

pred_matrix[, 'hardflare_time'] <- 0

# MICE with 10 imputations
mice_upf_uc_hard <- mice::mice(
  data = data_impute_upf_uc_hard,
  predictorMatrix = pred_matrix,
  m = 10,
  maxit = 20,
  seed = 73,
  print = FALSE
)

# Check convergence
mice_upf_uc_hard %>% plot()

```


```{r}

# Fit pooled Cox model
with(
  mice_upf_uc_hard,
  coxph(
    Surv(hardflare_time, hardflare) ~
      Sex +
      cat +
      IMD +
      dqi_tot +
      UPF_perc_cat +
      Smoke +
      frailty(SiteNo)
  )
) %>%
  mice::pool() %>%
  summary(conf.int = TRUE,
          conf.level = 0.95,
          exponentiate = TRUE)

```


