---
title: "Hospital Anxiety and Depression Scale (HADS) interactions — extended analysis"
author:
  - name: "Chiara Cotronei"
    url: https://www.research.ed.ac.uk/en/persons/chiara-cotronei
    affiliations:
      - ref: Lothian
  - name: "Alexander Rudge"
    affiliations:
      - ref: CGEM
  - name: "Nathan Constantine-Cooke"
    corresponding: true
    url: https://scholar.google.com/citations?user=2emHWR0AAAAJ&hl=en&oi=ao
    affiliations:
      - ref: CGEM
      - ref: HGU
format:
  html:
    code-fold: true
    code-tools: true
editor: visual
toc: true
toc-depth: 4
toc-expand: true
editor_options:
  chunk_output_type: console
---

This document contains the extended-analysis interaction models for HADS scores and subsequent patient-reported and objective flares.

# Setting up in R

```{r}
#| message: false
#| warning: false
knitr::opts_chunk$set(echo = TRUE, error = FALSE, warning = TRUE)
options(dplyr.summarise.inform = FALSE)

if (file.exists("../../functions.R")) {
  source("../../functions.R")
} else if (file.exists("../functions.R")) {
  source("../functions.R")
} else if (file.exists("Alex Rudge/functions.R")) {
  source("Alex Rudge/functions.R")
} else {
  source("src/Psychosocial/Alex Rudge/functions.R")
}

library(tidyverse)
library(magrittr)
library(survival)
library(mice)
library(splines)
```

# Load and prepare data

```{r}
if (file.exists("/.dockerenv")) {
  chiara <- "data/people/chiara/"
} else {
  chiara <- "/Volumes/igmm/cvallejo-predicct/people/chiara/"
}

hads <- readRDS(
  "/Volumes/igmm/cvallejo-predicct/people/Alex/Predicct2/Data/HADS.rds"
)
flares_soft <- readRDS(paste0(chiara, "flares_soft.RDS"))
flares_hard <- readRDS(paste0(chiara, "flares_hard.RDS"))

data_survival_soft <- hads %>%
  dplyr::inner_join(
    flares_soft %>% dplyr::select(ParticipantNo, softflare, softflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = softflare, time = softflare_time)

data_survival_anxiety_soft <- data_survival_soft %>%
  dplyr::filter(hads_type == "anxiety_hads")

data_survival_depression_soft <- data_survival_soft %>%
  dplyr::filter(hads_type == "depression_hads")

data_survival_hard <- hads %>%
  dplyr::inner_join(
    flares_hard %>% dplyr::select(ParticipantNo, hardflare, hardflare_time),
    by = "ParticipantNo"
  ) %>%
  dplyr::mutate(DiseaseFlareYN = hardflare, time = hardflare_time)

data_survival_anxiety_hard <- data_survival_hard %>%
  dplyr::filter(hads_type == "anxiety_hads")

data_survival_depression_hard <- data_survival_hard %>%
  dplyr::filter(hads_type == "depression_hads")
```

# Survival - interaction term

Testing the heterogeneity between ulcerative colitis and Crohn's disease by modelling the cohort together with an interaction term between ibd type and anxiety/depression.

## Anxiety - categorical

### Patient-reported flare

##### Cox proportional hazards model

###### Complete case

```{r}
# Cox model
cox_anxiety_soft_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group*diagnosis2 +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    control_8 +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_soft
)

cox_anxiety_soft_interaction %>%
  print_survival()

```

###### MICE

```{r}
# Cox model with MICE
# Run the mice
cox_anxiety_soft_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group*diagnosis2 +
    IMD +
    Sex +
    age_decade +
    FC +
    flare_group +
    OverallControl +
    control_8 +
    Smoke +
    frailty(SiteNo),
  data = data_survival_anxiety_soft,
  m = 20,
  printFlag = FALSE
)

# Fit the Cox
cox_anxiety_soft_interaction_pool <- with(cox_anxiety_soft_interaction_mice,
     coxph(
       Surv(time, DiseaseFlareYN) ~
         score_group*diagnosis2 +
         IMD +
         Sex +
         age_decade +
         FC +
         flare_group +
         OverallControl +
         control_8 +
         Smoke +
         frailty(SiteNo)
     )) %>%
  mice::pool()

cox_anxiety_soft_interaction_pool %>%
  broom::tidy(conf.int = TRUE,
              conf.level = 0.95,
              exponentiate = TRUE) %>%
  print_survival_mice()

```

### Objective flare

###### Complete case

```{r}
cox_anxiety_hard_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_hard
)

cox_anxiety_hard_interaction %>% print_survival()
```

###### MICE

```{r}
cox_anxiety_hard_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_hard,
  m = 20,
  printFlag = FALSE
)

cox_anxiety_hard_interaction_pool <- with(
  cox_anxiety_hard_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_anxiety_hard_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

## Depression - categorical

### Patient-reported flare

###### Complete case

```{r}
cox_depression_soft_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_soft
)

cox_depression_soft_interaction %>% print_survival()
```

###### MICE

```{r}
cox_depression_soft_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_soft,
  m = 20,
  printFlag = FALSE
)

cox_depression_soft_interaction_pool <- with(
  cox_depression_soft_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_depression_soft_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

### Objective flare

###### Complete case

```{r}
cox_depression_hard_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_hard
)

cox_depression_hard_interaction %>% print_survival()
```

###### MICE

```{r}
cox_depression_hard_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_hard,
  m = 20,
  printFlag = FALSE
)

cox_depression_hard_interaction_pool <- with(
  cox_depression_hard_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_depression_hard_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

## Anxiety - continuous

### Patient-reported flare

###### Complete case

```{r}
cox_anxiety_cts_soft_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_soft
)

cox_anxiety_cts_soft_interaction %>% print_survival()
```

###### MICE

```{r}
cox_anxiety_cts_soft_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_soft,
  m = 20,
  printFlag = FALSE
)

cox_anxiety_cts_soft_interaction_pool <- with(
  cox_anxiety_cts_soft_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_anxiety_cts_soft_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

### Objective flare

###### Complete case

```{r}
cox_anxiety_cts_hard_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_hard
)

cox_anxiety_cts_hard_interaction %>% print_survival()
```

###### MICE

```{r}
cox_anxiety_cts_hard_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_hard,
  m = 20,
  printFlag = FALSE
)

cox_anxiety_cts_hard_interaction_pool <- with(
  cox_anxiety_cts_hard_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_anxiety_cts_hard_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

## Depression - continuous

### Patient-reported flare

###### Complete case

```{r}
cox_depression_cts_soft_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_soft
)

cox_depression_cts_soft_interaction %>% print_survival()
```

###### MICE

```{r}
cox_depression_cts_soft_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_soft,
  m = 20,
  printFlag = FALSE
)

cox_depression_cts_soft_interaction_pool <- with(
  cox_depression_cts_soft_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_depression_cts_soft_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

### Objective flare

###### Complete case

```{r}
cox_depression_cts_hard_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_hard
)

cox_depression_cts_hard_interaction %>% print_survival()
```

###### MICE

```{r}
cox_depression_cts_hard_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_hard,
  m = 20,
  printFlag = FALSE
)

cox_depression_cts_hard_interaction_pool <- with(
  cox_depression_cts_hard_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      hads_score * diagnosis2 + IMD + Sex + age_decade + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_depression_cts_hard_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```

# Interaction - Age

The categorical HADS models are stratified by diagnosis and include an interaction between HADS score group and age decade. The extended-analysis covariates are retained in each model.

## Stratified survival data

```{r}
data_survival_anxiety_soft_uc <- data_survival_anxiety_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU")

data_survival_anxiety_soft_cd <- data_survival_anxiety_soft %>%
  dplyr::filter(diagnosis2 == "CD")

data_survival_anxiety_hard_uc <- data_survival_anxiety_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU")

data_survival_anxiety_hard_cd <- data_survival_anxiety_hard %>%
  dplyr::filter(diagnosis2 == "CD")

data_survival_depression_soft_uc <- data_survival_depression_soft %>%
  dplyr::filter(diagnosis2 == "UC/IBDU")

data_survival_depression_soft_cd <- data_survival_depression_soft %>%
  dplyr::filter(diagnosis2 == "CD")

data_survival_depression_hard_uc <- data_survival_depression_hard %>%
  dplyr::filter(diagnosis2 == "UC/IBDU")

data_survival_depression_hard_cd <- data_survival_depression_hard %>%
  dplyr::filter(diagnosis2 == "CD")
```

## Anxiety - categorical

### Patient-reported flare

#### Ulcerative colitis

::: panel-tabset
##### Complete case

```{r}
cox_anxiety_soft_uc_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_soft_uc
)

cox_anxiety_soft_uc_age_interaction %>% print_survival()
```

##### MICE

```{r}
cox_anxiety_soft_uc_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_soft_uc,
  m = 20,
  printFlag = FALSE
)

cox_anxiety_soft_uc_age_interaction_pool <- with(
  cox_anxiety_soft_uc_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_anxiety_soft_uc_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
:::

### Patient-reported flare

#### Crohn's disease

::: panel-tabset
##### Complete case

```{r}
cox_anxiety_soft_cd_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_soft_cd
)

cox_anxiety_soft_cd_age_interaction %>% print_survival()
```

##### MICE

```{r}
cox_anxiety_soft_cd_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_soft_cd,
  m = 20,
  printFlag = FALSE
)

cox_anxiety_soft_cd_age_interaction_pool <- with(
  cox_anxiety_soft_cd_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_anxiety_soft_cd_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
:::

### Objective flare

#### Ulcerative colitis

::: panel-tabset
##### Complete case

```{r}
cox_anxiety_hard_uc_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_hard_uc
)

cox_anxiety_hard_uc_age_interaction %>% print_survival()
```

##### MICE

```{r}
cox_anxiety_hard_uc_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_hard_uc,
  m = 20,
  printFlag = FALSE
)

cox_anxiety_hard_uc_age_interaction_pool <- with(
  cox_anxiety_hard_uc_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_anxiety_hard_uc_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
:::

### Objective flare

#### Crohn's disease

::: panel-tabset
##### Complete case

```{r}
cox_anxiety_hard_cd_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_hard_cd
)

cox_anxiety_hard_cd_age_interaction %>% print_survival()
```

##### MICE

```{r}
cox_anxiety_hard_cd_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_anxiety_hard_cd,
  m = 20,
  printFlag = FALSE
)

cox_anxiety_hard_cd_age_interaction_pool <- with(
  cox_anxiety_hard_cd_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_anxiety_hard_cd_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
:::

## Depression - categorical

### Patient-reported flare

#### Ulcerative colitis

::: panel-tabset
##### Complete case

```{r}
cox_depression_soft_uc_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_soft_uc
)

cox_depression_soft_uc_age_interaction %>% print_survival()
```

##### MICE

```{r}
cox_depression_soft_uc_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_soft_uc,
  m = 20,
  printFlag = FALSE
)

cox_depression_soft_uc_age_interaction_pool <- with(
  cox_depression_soft_uc_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_depression_soft_uc_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
:::

### Patient-reported flare

#### Crohn's disease

::: panel-tabset
##### Complete case

```{r}
cox_depression_soft_cd_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_soft_cd
)

cox_depression_soft_cd_age_interaction %>% print_survival()
```

##### MICE

```{r}
cox_depression_soft_cd_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_soft_cd,
  m = 20,
  printFlag = FALSE
)

cox_depression_soft_cd_age_interaction_pool <- with(
  cox_depression_soft_cd_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_depression_soft_cd_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
:::

### Objective flare

#### Ulcerative colitis

::: panel-tabset
##### Complete case

```{r}
cox_depression_hard_uc_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_hard_uc
)

cox_depression_hard_uc_age_interaction %>% print_survival()
```

##### MICE

```{r}
cox_depression_hard_uc_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_hard_uc,
  m = 20,
  printFlag = FALSE
)

cox_depression_hard_uc_age_interaction_pool <- with(
  cox_depression_hard_uc_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_depression_hard_uc_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
:::

### Objective flare

#### Crohn's disease

::: panel-tabset
##### Complete case

```{r}
cox_depression_hard_cd_age_interaction <- coxph(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_hard_cd
)

cox_depression_hard_cd_age_interaction %>% print_survival()
```

##### MICE

```{r}
cox_depression_hard_cd_age_interaction_mice <- coxph_mice(
  Surv(time, DiseaseFlareYN) ~
    score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo),
  data = data_survival_depression_hard_cd,
  m = 20,
  printFlag = FALSE
)

cox_depression_hard_cd_age_interaction_pool <- with(
  cox_depression_hard_cd_age_interaction_mice,
  coxph(
    Surv(time, DiseaseFlareYN) ~
      score_group * age_decade + IMD + Sex + FC + flare_group + OverallControl + control_8 + Smoke + frailty(SiteNo)
  )
) %>% mice::pool()

cox_depression_hard_cd_age_interaction_pool %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  print_survival_mice()
```
:::

# Reproduction and reproducibility {.appendix}

<details class="appendix">

<summary>Session info</summary>

```{r}
#| echo: false
pander::pander(sessionInfo())
```

</details>

Licensed by <a href="https://creativecommons.org/licenses/by/4.0/">CC BY</a> unless otherwise stated.
